### 2017-02-15 Paper_NA.tex
first commit

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{bm}
\usepackage[english]{babel} 
\usepackage{xcolor}
\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{natbib} % Powerful citation management
\bibpunct[; ]{(}{)}{,}{a}{}{;} % Custom minimal bibliography styling
\renewcommand{\baselinestretch}{2}\normalsize

\title{Paper NA}
\date{\today}

\begin{document}

\maketitle

\section{Introduction}

The North American continent consists of an ancient cratonic core, formed during the Archean that has not been subjected to orogeny for at least 1.8 Ga, bordered to the southeast, by progressively younger, stable, Proterozoic provinces \citep{hoffman1988united}. In the west, the Rocky Mountain Front (RMF) separates the Proterozoic and Archean shield from younger, tectonically active provinces (Fig. 1). This relatively regular configuration makes North America an ideal target for seismic tomography, to investigate the relationship of lithospheric and asthenospheric structure to the geological features observed at the surface.

Differences in the seismic velocity structure between the craton and the western US extending down to at least 200 km were established already in the 1960's and 70's from the first studies of teleseismic P and S wave travel time anomalies  \citep[e.g.][]{cleary1966analysis,herrin1968regional,poupinet1979relation}, the first tomographic images based on P and S travel times \citep{romanowicz1979seismic} and long period surface waveforms \citep{woodhouse1984mapping}, as well as contrasted velocity-depth profiles obtained from the modeling of shear body waveforms at relatively short distances \citep[e.g.][]{grand1984upper}.

More recent shear wave tomography, at scales from global \citep[e.g.][]{su1994degree,megnin2000three,shapiro2002monte,panning2006three,kustowski2008anisotropic,ritsema2010s40rts,lekic2011inferring,debayle2012global,schaeffer2013global} to continental \citep{lee2005surface,yuan20113,yuan2014lithospheric,schaeffer2014imaging}, confirm the presence of strong lateral variations in the thickness of the lithosphere across the RMF, with lithospheric roots extending down to 200-250 km under the craton, thinning abruptly to the west to less than 80-100 km. This is also found from azimuthal anisotropy tomography, where it is manifested by a change in the anisotropy fast axis direction across the lithosphere-asthenosphere boundary (LAB) \citep{marone2007depth,yuan2010lithospheric}.

While the mechanism of craton formation is still widely debated \citep[e.g.][]{s2005archean,lee2011building}, the presence of finer scale lateral and depth variations in seismic structure suggest a complex history. Studies of azimuthal anisotropy have shown the presence of laterally varying layering within the cratonic lithosphere \citep[e.g.][]{levin1999shear,deschamps2008stratified,yuan2010lithospheric}, which may indicate different modes and/or times of formation of the top c.a. 100 km of this lithosphere. This is also suggested by fine scale studies of converted and reflected phases, which indicate the presence of a sharp mid-lithospheric discontinuity (MLD) within the craton, marking the top of a mid-lithospheric low velocity zone \citep{thybo1997seismic,bostock1998mantle,abt2010north,fischer2010lithosphere,rader2015characterization,ford2016midlithospheric} and even possibly several mid-lithospheric discontinuities \citep{calo2016layered}. In addition, there is evidence from shear wave radially anisotropic tomography for separation between blocks of different crustal ages extending to depths of at least 150 km \citep[e.g.][]{yuan2014lithospheric}.

Much of this finer scale structure has been highlighted owing to the availability of data from the dense broadband USArray TA deployment. With the completion of the coverage of the conterminous US, as well as availability of data from Canada and Greenland, it is now possible to further refine shear wave tomographic images of North America, and in particular of the stable Proterozoic and Archean provinces, to try and improve our understanding of its formation. It is also an unprecedented opportunity to experiment with improved waveform-based tomographic techniques.

In a previous study, we presented a radially anisotropic shear velocity model of the North American upper mantle based on a combination of long period teleseismic and regional waveform data \citep{yuan2014lithospheric} The regional waveform data (down to 40 s period) were, for the first time at this scale, compared to 3D synthetics computed using RegSEM \citep{cupillard2012regsem}, a Spectral Element Method (SEM) code appropriate for continental scale wavefield computations. Due to the extremely lengthy computations that would have been necessary to compute it numerically at each iteration of the inversion, the predicted teleseismic wavefield was, instead, computed using Non-Linear Assymptotic Coupling Theory \citep[NACT,][]{li1995comparison}, a methodology based on normal mode perturbation theory, that is approximate, but computationally more efficient, and has been used in the development of several generations of global and continental scale shear velocity models \citep{li1996global,megnin2000three,gung2003global,panning2006three,yuan20113}

The resulting 2014 model of North America presents some interesting features, in particular a correlation of radial anisotropy structure with lithospheric blocks corresponding to different orogenies in the eastern US and continental shelf. 
While more rigorous than inversions practiced by most groups and based on the path-average surface wave approximation, this mixed methodology nevertheless presents some inconsistencies from the theoretical point of view, since the predicted wavefield through the target model space is computed with different theories for teleseismic and regional distance data.

 To better integrate teleseismic data into our regional tomographic inversions, we developed a general framework called "Box Tomography"  (see \cite{masson2013numerical}, Masson and Romanowicz, 2016a,b) that allow us to consistently model and invert both teleseismic (i.e. associated with sources or receivers outside of the imaged region) and regional (i.e. associated with sources and receivers within the imaged region) waveform data using accurate numerical methods such as SEM. In Box tomography, prior to the inversion, the seismic wavefield generated by teleseismic sources is first modeled numerically at the global scale for a given reference model and recorded at the surface of the region to be imaged.  This reference wavefield is then used to construct virtual sources lying at the boundary of the regional modeling domain and reproducing the original wavefield as illustrated in Figure~\ref{injection}. Once the telesismic sources have been moved (i.e. replaced by virtual sources) within the regional modeling domain, the tomographic inversion can be performed efficiently in a classic manner (i.e. using regional modeling only) as the teleseismic data are accounted for seamlessly thanks to the virtual sources. Related concepts have been proposed to account for teleseismic data into regional imaging (e.g. Wang et al., 2016 Monteiller et al., 2015), however, in these studies the global modeling of the reference wavefield is performed using a faster method that do not account for the effects induced by the 3D structure of the Earth outside the regional modeling domain. In this paper, we present the first application of Box tomography, where the 3D structure is accounted for outside the imaging box, to continental scale waveform tomography in North America.

\section{Methodology}

Many studies have inverted for continental scale structure, in particular in North America, using fundamental mode and, in some cases, overtone surface wave dispersion data or waveforms observed at teleseismic distances. 
The usual practice is to first consider the best possible global scale tomographic shear wave velocity model, compute predictions of observables in this model using an approximate theory, generally the surface wave path average approximation \citep[e.g.][]{nettles2008radially,bedle2009svelocity,schaeffer2014imaging}, or, in our group, NACT \citep{marone2007three,yuan20113}. 
In the case when secondary observables such as dispersion data are considered, the contribution from outside of the target region is calculated once and for all in the background global model, and subtracted from the observed dispersion data. 
The resulting residual is attributed to structure in the target region and the tomographic inversion proceeds within this target model volume. 
When inverting waveforms, and in particular when a more accurate theory than the PAVA is considered, such a simple procedure is not possible, and it is necessary to recompute the synthetic teleseismic waveforms, at each iteration, in a 3D model which is fixed outside of the target region, and updated only within it. 
This leads to substantial computations, even in the case of an approximate theory such as NACT, and becomes prohibitively expensive, if the 3D teleseismic wavefield is computed using SEM or another accurate numerical method, as one aims to include increasingly shorter periods.

{\color{red}To overcome this issue, we take advantage of the "Box Tomography" theory which allows to combine teleseismic and regional waveforms in a consistent manner to image regional targets at arbitrary locations. This general framework accomodates for arbitrary acquisition setups (i.e. with sources and receivers located outside the regional imaging box), is compatible with most popular numerical methods such as SEM or FD, and can produce exact seismograms and senitivity kernels (i.e. similar to those obtained when solving the problem globally). In this study, we limit ourselves to the situation where both regional and teleseismic events are employed but where all the seismic stations lie within the regional computational domain. Furthermore, we neglect some higher order scattering effects as proposed by Masson and Romanowicz (2016a). In this situation, Box Tomography is particularly efficient as the computational effort  to account for teleseismic data in regional inversions is limited to a few global scale simulations that are done once and for all prior to the inversion. Masson and Romanowicz (2016b) showed that this approach can produce accurate regional tomographic images even though the elastic structure outside the imaged region is neither fully known prior to the inversion nor updated during the inversion.}

Here, we apply this methodology to the case of teleseismic three component waveform data observed at stations within North America, combined with "regional" waveform data for which both earthquakes and stations are located within the target region.  This is the first time this methodology is applied in practice, and this study represents a proof of concept for this approach. Adding teleseismic data allows better azimuthal coverage of the target region than can be achieved using only the regional dataset and eventually provides for the inclusion of constraints from additional phases such as, for example, teleseismic SS phases reflected inside the target region.

\subsection{Dataset and model parametrization}.

The dataset includes three component acceleration waveforms from more than 2400 permanent and temporary broadband stations located in the target region, a $90^o \times 90^o $ area encompassing most of north America (Fig. 3). We considered two different datasets:
\begin{itemize}
	\item a regional dataset consisting of 155 events ($4.5 < M_w < 6.0$) for which sources are located within the target region.
	\item a teleseismic dataset, consisting of 122 events ($5.5 <M_w <6.9$) for which sources are located outside of the target region.
\end{itemize}

The waveforms are filtered between 40 and 400 s, with corner frequencies of 53 and 250 s, and windowed into wavepackets, according to the procedure of \cite{li1996global} allowing different weights to be applied according to relative amplitude of individual wavepackets, redundancy of paths and signal to noise ratio. This weighting step is crucial, as it homogeneizes the data coverage within the region and provides a way to construct the data covariance matrix $C_D$, which we approximate as a diagonal matrix. 
At each iteration of the inversion, only those wavepackets are considered that satisfy predefined goodness of fit criteria compared to synthetics calculated in the current 3D model, which avoids cycle slipping in our time domain waveform inversion procedure. As the model improves, at each iteration, additional wavepackets are included. The total number of wavepackets of fundamental and overtone Love and Rayleigh waves is given in Table 1, and provides a good azimuthal coverage within the NA continent (Fig. 4).

In addition to long period waveforms, we also consider a group velocity dispersion dataset \citep{shapiro2002monte} provided in the form of $1^o \times 1^o $ maps between 25 and 100 s. The shorter periods (below 60 s) are used to constrain our homogeneized crustal structure at each iteration, as described below, while the entire period range is included during the inversion for mantle structure, providing additional constraints for structure in the shallow parts of the mantle that are consistent with the treatment of the crust.

As in our previous work at the global scale \citep{french2013waveform,french2014whole}, we do not rely on an existing layered crustal model. There are several drawbacks to considering such models. First, they are constructed using data for limited regions that are then extrapolated to other regions based on a tectonic regionalization, and may not always fit real waveform data very well. Second, the inclusion of thin low velocity layers slows down the SEM computation significantly. 
Instead, we compute a radially anisotropic smooth crustal model on a $2^o \times 2^o $ grid through Monte Carlo Markov Chain (MCMC) simulation constrained by the group velocity dispersion data. This process is a sort of "homogeneization" (e.g. Backus, 1961; Capdeville and Marigo, 2008), which yields a crustal model equivalent to a real crust within the period range of the data used to constrain it (i.e. for periods longer than 20s). 
The thickness of the crust is a priori chosen as that of model Crust2.0 \citep{basscrust2} where it is larger or equal to 30 km, and it is fixed at 30 km otherwise. 
This results in a smooth, but realistic crust within most of the continent, and a crustal model that is not interpretable in the border regions of our model space (i.e. in the oceans and borders of the continent). 
We have shown that this does not bias the mantle structure obtained below 50 km \citep{french2014whole}. 
At each iteration of the inversion, we only invert for structure below 30 km depth, and subsequently recompute the crustal model for the next iteration by applying our MCMC approach with mantle structure fixed to be that of the current iteration model. The procedure is described and discussed in detail in \cite{french2014whole}

Assuming that our azimuthal coverage is everywhere sufficient within our target region (Fig. 4), we here solve only for radially anisotropic structure and specifically for two of the 5 parameters that describe such a structure, the two parameters to which long period surface waveforms are the most sensitive: isotropic velocity $V_{S}$ and the anisotropic parameter $\xi = {\frac{V_{SH}}{V_{SV}}}^2$, where $V_{SH}$ is the velocity of waves polarized horizontally and $V_{SV}$ that of waves polarized vertically . 
The other three parameters and density are constrained through empirical scaling relationships, following \cite{montagner1989petrological}, and are based on laboratory measurements for upper mantle rocks. 
If (A,C,F,L and N) are the 5 elastic parameters describing a radially anisotropic (or VTI) medium (Love, 1927), then:

\begin{equation}
 N = \rho V^{2}_{SH}, \quad L = \rho V^{2}_{SV}, \quad A = \rho V^{2}_{PH} , \quad C = \rho V^{2}_{PV} 
 \end{equation}
then the medium can equivalently be described by the 5 parameters ($V_{S}, V_{P}, \xi, \phi, \varepsilon$) where:

\begin{equation}
V_{S} = \sqrt{\frac{2V^{2}_{SV}+V^{2}_{SH}} {3}} \\
V_{P} = \sqrt{\frac{V^{2}_{PV} + 4V^{2}_{PH}}{5}}
 \xi = \frac{N}{L}, \quad \varphi = \frac{C}{A}, \quad \eta = \frac{F}{A-2L} 
 \end{equation}

 The scaling parameters considered are \citep{montagner1989petrological}

 \begin{equation}
\frac{\delta(V_{P})}{\delta(V_{S})} = 0.5, \quad
\frac{\delta(\rho)}{\delta(V_{S})} = 0.33, \quad
\frac{\delta(\eta)}{\delta(\xi)} = -2.5,   \quad
\frac{\delta(\varphi)}{\delta(\xi)} = -1.5,   \quad
\end{equation}

We prefer parametrization in terms of $V_S$ and $\xi$, rather than $V_{SH}$ and $V_{SV}$, as this allows us to consider different spatial resolution and apply higher damping in the inversion to the less well resolved anisotropic parameter $\xi$, rather than having to reconstruct this parameter from differences in perturbations in the two quantities $V_{SH}$ and $V_{SV}$, which would limit the resolution allowed in isotropic velocity.

In previous studies \citep{marone2007depth,yuan2010lithospheric}, after inverting for radial anisotropic structure \citep{marone2007three,yuan20113}, we also inverted for azimuthal anisotropy, by adding constraints from SKS splitting measurements. This step will be the topic of a separate publication.

The target model space is geographically defined as shown in Figure 3, and is limited in depth down to 800 km. As in our previous tomographic studies, the model space is parametrized in terms of 26 cubic splines $\nu(r)_{q}$ vertically \citep{megnin2000three} from the core-mantle boundary to the Moho, although we invert for structure only in the top 16 splines, corresponding to the top 700-800 km of the mantle. The spline nodes are spaced more closely at shallow depths, and located at the following radii: 5690, 5810, 5910, 5900, 6061, 6101, 6131, 6161, 6191, 6221, 6241, 6261, 6281, 6301, 6321, 6341 km.  
Laterally, we parametrize our model in terms of spherical splines $\beta(\theta,\phi)_{p}$ \citep{wang1995spherical}. 
The combination of vertical and spherical splines constitutes a local basis for the description of smooth functions within the model volume. 
Thus, the value of a model parameter $m(r,\theta,\phi)$ can be computed at any point in space given the set of spline coefficients $m_{pq}$:

\begin{equation}
	m(\theta,\varphi,r) = \sum_{p}\sum_{q}m_{pq}\beta_{p}(\theta,\varphi)\nu_{q}(r)
\end{equation}

The spherical spline parametrization has the advantage of allowing for variable grid parametrization, which can be adjusted according to data coverage \citep[e.g.][]{marone2007three}. 
In our case, outside of the target region, we adopt a "level 6" spherical grid for $V_{S}$ ($2^{o}$ knot spacing) and a "level 4" spherical grid for $\xi$ ($8^{o}$ knot spacing), consistent with the parametrization of SEMUCB\_wm1. Inside the well-sampled target region, we define a level 7 spherical grid ($1^{o}$ knot spacing) for $V_{S}$ and level 6 ($2^{o}$ knot spacing) for $\xi$.

 \subsection{Forward modelling}

\begin{figure}
\centering
\begin{tabular}{ccc}
(a)&&(b)\\
  \includegraphics[width=.45\textwidth]{snap_glob.pdf}  &&
  \includegraphics[width=.45\textwidth]{snap_reg.pdf} \\ 
  \includegraphics[width=.45\textwidth]{seismo.pdf}      &&
  \includegraphics[width=.45\textwidth]{seismo.pdf}
\end{tabular}
\caption{\label{injection}Comparison between a global scale simulation and a regional scale simulation of wave propagation following an earthquake in south America. The seismograms corresponds to the recordings at 37 station located in north America (red spheres). The pink sphere shows the epicenter of the earthquake. In (a), the wavefield is modeled globally using the spectral element method Specfem3D\_globe \citep{komatitsch2002spectrala}. In (b), the wavefield is modeled regionally using the RegSEM  
During the inversion, all the synthetic seismograms are computed using the regional SEM code RegSEM \citep{cupillard2012regsem} code and regenerated thanks to virtual sources located around the computational domain \citep[see][]{masson2013numerical}. The sismograms in (b) are exactly similar to those in (a), however, the computational effort is greatly reduced in the regional simulation (b).}
\end{figure}

During the inversion, all the synthetic seismograms are computed using the regional SEM code RegSEM \citep{cupillard2012regsem} that takes into account effects of oceans, topography/bathymetry, ellipticity, and anelasticity, and where the computational domain is truncated using absorbing boundaries. 
The synthetic seismograms associated with regional data (i.e. where both the seismic stations and the earthquake are located within the regional modeling domain) do not require any specific treatment and are modeled as in our previous studies \citep[e.g.][]{yuan2014lithospheric}. 
The synthetic seismograms associated with teleseismic data (i.e. where the earthquake and the seismic stations are located outside and inside the regional modeling domain, respectively) are obtained using a two step procedure as proposed by Masson and Romanowicz (2016). 
Prior to the inversion, the wavefields generated by teleseismic earthquakes are computed globally within our starting model ($SEMucb\_wm1$) and using an adapted version of SPECFEM3D\_globe \citep{komatitsch2002spectrala}. 
During these global simulations, the 3 component displacement wavefield is recorded at a set of points with locations prescribed by the RegSEM code.
Within the regional solver RegSEM, these points are the coallocation or Gauss-Lobatto-Legendre (GLL) points belonging to the one element thick surface surrounding the regional modeling domain \citep[see][]{masson2013numerical}. 
Notice that our procedure do not require to store the stress nor the strain wavefield which are computed naturally by the regional solver, this makes it easy to swap between different codes for modeling global wave propagation (i.e. any code that outputs displacement seismograms can be used as such).
Because the Courant criteria which ensure the stability in SEM often leads to oversampling, we compress the recordings versus time using a least square B-spline transform (Unser et al., 1993ab), we found this  approach more efficient and practical than the more classic decimation/interpolation scheme.
We typically achieve data compression ratio lying between 10 and 100 with no significant loss of accuracy. 
During the inversion, the global recordings are transformed to virtual sources that regenerate the global wavefield regionally, this operation is done on the fly by the RegSEM code and illustrated in Figure ~\ref{injection}b.

Overall, the additional computational effort to account for teleseismic event consists of one global simulation per teleseismic event that are done once and for all before the inversion starts.

 \subsection{Inverse modelling}

In continuity of our previous work \citep[e.g][]{marone2007three,yuan20113,yuan2014multiscale}, we use a hybrid iterative inversion scheme where, at each iteration, the forward wavefield is computed accurately using SEM, but the inverse step is solved using the formalism of \cite{tarantola1982generalized}  with sensitivity kernels calculated approximately using normal mode perturbation theory. This allows us to apply a fast converging Gauss-Newton quadratic optimization scheme. 
As shown in \cite{tarantola2005inverse}, and in Appendix A of \cite{lekic2011inferring}, it is far more important to use an accurate forward modelling scheme to compute the misfit function, while inaccuracies in the theoretical treatment of kernels result in smoothing errors that can be compensated for in subsequent iterations. Currently popular "adjoint tomographic" approaches \citep{zhu2015seismic,bozdag2016global} do compute the numerically exact gradient, however, they also need to apply smoothing operators and regularization, which degrades the accuracy of their kernels. Importantly, they rely on a linear optimization scheme (conjugate gradient method), which is characterized by very slow convergence.

Our misfit function is defined in the time domain from the point by point differences between observed and synthetic waveforms by:

\begin{equation}
2\Phi(m_{k}) = [d - g(m_{k})]^{T} C^{-1}_{d} [d-g(m_{k})] + [m_{p}-m_{k}]^{T} C_{m}^{-1} [m_{p}-m_{k}]
\end{equation}

where $m_{k}$ represents the model estimate at the k-th iteration, $d$ is the data vector (waveform discretized in time or group velocity as a function of period) and $g(m_k)$ is the corresponding discretized wavefield computed using SEM, or the predicted group velocity dispersion. $m_p$ is the model prior (i.e. the 3D starting model) and $C_m$ and $C_d$ represent a priori model and data covariance matrices, respectively. Minimizing $\Phi$ in the sense of the L2 norm leads to the equation for the k+1 model update:

\begin{equation}
m_{k+1} = m_{k} + (C_{m}G^{T}_{k}C^{-1}_{d}G_{k} + I)^{-1} (C_{m}G_{k}^{T}C^{-1}[d-g(m_{k})]+ m_{p}-m_{k})
\end{equation}

where $G$ is the matrix of Frechet derivatives of $g(m)$ calculated at $m_k$. We compute G using NACT or the Path Average Approximation (PAVA,\cite{woodhouse1984mapping}), depending on the distance range of the corresponding source-station path.
NACT and PAVA are both asymptotic (high frequency) approximations to normal mode perturbation theory. 
The PAVA includes along branch mode coupling only and is the standard surface wave approximation \citep[e.g][]{mochizuki1986free,romanowicz1987multiplet}, as used for example in \cite{woodhouse1984mapping}, in which a frequency shift and distance shift are introduced for each mode to account for the effects of heterogeneous structure. 
The corresponding kernels are "1D", i.e. they only depend on the average structure between the source and the receiver. This approximation is valid for single-mode seismograms, such as fundamental mode surface waves \citep[e.g][]{romanowicz2008computation}.
NACT includes across branch-coupling, in addition to PAVA, which brings out 2D sensitivity of waveforms in the vertical plane containing the source and the receiver. NACT breaks down when the distance between the source and station is short, so we compute kernels using NACT for epicentral distances larger than $15^o$, and PAVA for shorter distances.
Neither PAVA nor NACT consider off-great circle plane sensitivity (i.e. focusing effects, e.g. \cite{zhou2005finite}). These effects become important for accurate amplitude fitting. With our choice of misfit function, we are first and foremost fitting the phase, and for that, the 2D effects in the vertical plane are dominant, and important especially for overtones, as illustrated in \cite{romanowicz2008computation}.

% \subsection{Tectonic back ground}

% Cratons are large areas of continental lithosphere that have remained coherent and relatively rigid since Precambrian \citep{hoffman1988united}. Through more than $3Ga$, cratons have survived compared to the $<200Ma$ life span of ocean basins.
% In accordance with the absence of geoid anomalies over continents, Jordan proposed the concept of Continental Tectosphere \citep{jordan1975continental} where continental mantle is made of low-density, melt-depleted peridotite residua, whose low densities exactly offset the effect of thermal contraction associated with a growing thermal boundary layer \citep{lee2011building}.
% The fact that the continental lithosphere manages to avoid the fate of recycling suffered by oceanic lithosphere has been attributed to high viscosity \citep{sleep2003survival, doin1997mantle}, neutral to positive buoyancy, low heat flow compared to surronding area \citep{jaupart2014building}, dehydrated \citep[i.e dryer than oceanic mantle until 250km deep according to electrical conductivity tomography,][]{yang2015three, hirth2000comparison}, and chemically depleted until ca. 175km deep \citep{kopylova1999petrology} while still fertile below until ca. 300km \citep{s2005archean}.

% Since \cite{grand1984upper}, one well established feature within NA lithosphere is the contrast between the high velocity craton and the low velocity western active part. This feature is observed in many global models (e.g., \citep{megnin2000three, shapiro2002monte, panning2006three, houser2008shear, nettles2008radially, kustowski2008anisotropic, ritsema2010s40rts, simmons2010gypsum, lekic2011inferring, debayle2012global, schaeffer2013global, french2014whole,bozdag2016global} where the cratonic lithosphere is caracterized by higher than average (ca. $4.8 km.s^{-1}$) velocities down to at least $200km$ deep.

% % Such values, at upper mantle depths with Standart Temperature Condition, are very important when interpreting shear wave velocities. For example, a relatively small velocity increase of $1 \%$ in the uppermost mantle can be attributed to a $5$ unit increase in the Mg\# relative to Fe, or an $220°C$ decrease in temperature. \citep{lee2003compositional, bedle2009svelocity}

% % The combination of low temperature and depleted lithosphere induces observed high velocity anomalies.

% Continental scale studies showed that, fastest anomalies are beneath archean and proterozoic terranes around 150km deep \citep{bedle2009svelocity,lee2005surface,schaeffer2014imaging,marone2007three,yuan20113}, there are correlations between surface tectonics provinces and shear waves velocities heterogeneities \citep{nettles2008radially,yuan2014lithospheric,darbyshire2013seismic}, upper mantle cratonic keels extends down to ca. 250km with lateral thin out downward. \cite{yuan2010lithospheric} infered from azimuthal anisotropy study, that cratonic keels are vertically stratified and comprise a cold core of highly depleted material bounded deeper by a fertile (less depleted) thermal layer. While the sharp Mid-Lithospheric Discontinuity (MLD) between the chemical layer and thermal layer is observed (e.g \cite{thybo1997seismic,abt2010north,calo2016layered}), the Lithosphere-Asthenosphere Boundary (LAB) is more diffuse and is identified as flat in topography through a negative gradient in shear wave velocities associated with an alignement of anisotropic fast axes direction with Absolute Plate Motion. \citep{yuan2010lithospheric,eaton2009elusive}\\

% At local scale, it has been observed that the stagnant shield suffered deformations from the convective mantle \citep{kaban2015cratonic,eaton2007seismic,frederiksen2013modification,boyce2016subduction}

% !!! Still a lot of local studies to include,  Greenland, deep slab imaged, Wyoming, talk more about Divot, Hotspots, Appalachian, Gulf of Mexico, JdF. !!!
% % This structure expands over both Archean and Proterozoic terranes in the canadian shield, which is not necessarly the case for all archean terranes within the plate (e.g the Wyoming block in the US \citep{yuan2014lithospheric}).

% Such high shear wave velocity values, at upper mantle depths with Standart Temperature Condition, are very important when interpreting shear wave velocities. For example, a relatively small velocity increase of $1 \%$ in the uppermost mantle can be attributed to a $5$ unit increase in the Mg\# relative to Fe, or an $220°C$ decrease in temperature. \citep{lee2003compositional, bedle2009svelocity}

% Mechanism for craton formation is debated (e.g., \cite{lee2011building} \cite{s2005archean}), and caracterizing cratons through high resolution imaging gives clues to elucidate this puzzle.
% We propose interpretations on the geologic and tectonic evolution of the continent since both parameter variabilities are related to compositional, mineralogical and thermal states of the mantle.

% The North American (NA) continent is fully part of the North American plate. This tectonic plate is moving at ca. $2cm$ per year in a west-south westerly direction around a rotation pole located in north western Ecuador \citep{bleeker2005north} Along its western margin, we can find:
%  \begin{itemize}
%  	\item Subduction of the Pacific plate under NA plate south to Alaska, the Cascadian subduction zone of the Juan de Fuca plate and the Coco plate subducting under central America, both generating arc volcanism and being part of the ancient Farallon plate which was subducting under NA plate since Jurassic period.
%  	\item Strike-slip along the Queen Charlotte-Fairweather system in Canada and along the San Andreas fault in California.
%  \end{itemize}

%  The evolution of western North America has been controlled by this Mesozoic-Cenozoic ocean-continent collisional system. 
%  Subduction events of oceanic lithosphere are thought to be responsible for the Laramide orogeny (Late Cretaceous), characterized by minor horizontal shortening in the crust, but considerable crustal thickening generating the rocky mountains (Bird 1988).
%  The Rocky-Mountain Front (RMF) almost perfectly coincides with the boundary between this tectonically active area and the more stable cratonic part.

%  Two third of the Laurentia craton consist of distinct archean provinces: 
% \begin{itemize}
% 	\item undeformed Slave, Superior and Wyoming cratons
% 	\item highly reworked Rae and Hearne cratons (Churchill structural Province)
% 	\item Nain and North Atlantic Craton (NAC) underlying southern Greenland
% 	\item Sask which underlies the Trans-Hudson orogen Proterozoic belts (THO)
% \end{itemize}
%   All of these archean blocks, comprising most of the Canadian shield, were mainly welded by Proterozoic belts like the THO, but it is still under debates whether archean blocks originated from related ancestral landmasses (i.e a unique super craton, a couple of super cratons, or many super cratons). Greenland, despite its continental part was separated from NA continent ca. $80Ma$ ago, is fully part of the NA plates. While the NAC on its southern part has being studied over the last decades, central (Apparently a continuity of the THO \cite{bleeker2005north} and northern (apparently Rae-like archaean blocks) part are still poorly documented.

%  In the periphery of this NA archean core, proterozoic belts like Wopmay, Penokean and Makkovik-Ketilian orogens from 2.0 to 1.8 Ga and Labradorian, Central Plains and Yavapai-Mazatzal orogens from 1.8 to 1.6 Ga represent zones of accretion for Proterozoic material and mark the amalgamation of the archean cores. \citep{hoffman1988united} 
%  The Grenvile orogen (ca. $1.1 Ga$) at the southeast of the craton core completed to shape the Laurentia craton as we can observe it today, surviving both Rodinia ($825-550Ma$) and Pangaea ($335-175Ma$) break-ups.

\section{Results}

\subsection{Continental model}
To link with the 1D profile!
As observed in many other studies, average profiles of $V_S$ within North America (NA) in the shallow upper mantle is higher compared to Global average profiles biased by ocean structures. 
The difference is clearly distinct until ca. $270km$ and then fade away deeper, though, NA remains slightly higher than the global average. 
The global negative gradient of $V_S$ near $150km$ is here less pronounced beneath the continent by the influence of the positive gradient beneath the cratonic part. 
NA average of $\xi$ is lower than the global average, the latter influenced by the strong $\xi > 1$ oceanic signature where shearing is more important. 
Both trends are similar, while maxima and minima are always slightly deeper than globally and the change from $V_{SH} > V_{SV}$ to $V_{SH} < V_{SV}$ is reached at ca. $420km$.

\subsection{Beneath the craton}
\begin{figure}[h]
\centering
\label{cratoncross}
% \begin{tabular}{ccc}
\includegraphics[width=.45\textwidth]{/Users/Pierre/Documents/MATLAB/figures/iter2/profiles_NASEM_vs_dxi_A.pdf}  &&
% \end{tabular}
\caption{\bm{Cross Section of the lithospheric core}}
\end{figure}

To link with slices and a set of profiles
% Gung 2003 Positive anomalies of dlnxi down to 400km in the craton.

Our study confirms the presence of a thick cratonic root with faster than average $V_S$ down to $250km$ deep, with highest velocities, ca. $4.8km.s^{-1}$, located at depths shalower than at most $150km$. 
Laterally, the cratonic root in terms of $V_S$, coincides with precambrian aged crust. 
The western and eastern/south eastern edges follow the Rocky Mountain Front (RMF) and the Llano-Grenville Front respectively. 
The northern part is delimited by the coastine of the Arctic ocean including the arctic archipelago and Greenland. 
These lateral edges are well observed until $150km$, below, the cratonic root retracts toward the center of the archaean continent beneath Slave, Hearne/Rae and western Superior cratons. 
Everywhere within the lithosphere, a negative gradient of $V_S$ is observed with a minimum of ca. $4.6km.s^{-1}$ reached at $300km$. 
Deeper, this gradient switches to a positive trend. While the beginning of the negative gradient shows a topography, the bottom of the gradient located beneath Slave, Hearne/Rae and western Superior cratons; appears very flat.
As observed in Gung et al., 2003; down to 250km negative anomalies of $dln(\xi)$ are observed in the whole craton. 
Deeper, positive anomalies are emerging beneath archaean Slave/Rae blocks and proterozoic Yavapai/Mazatzal/Great Rhyolites provinces, however $dln\xi$ anomalies remain positives beaneath the archaean Superior block. 
In the whole lithosphere, $\xi$ is higher than $1$ which means that $V_{SH} > V_{SV}$.

\subsubsection{Wyoming}
The Archaean aged Wyoming craton does not show as high anomalies as within the craton nucleus. 
It is split into 2 distinctive part. A western part showing low velocities influenced by the active tectonic with a radially positive gradient. 
In contrast with the eastern part showing higher velocities with a radially negative then positive gradient. 
The depth of negative gradient's minima is variable, from 70 (western part) to 250km (eastern part) deep. 
The separation between the two areas coincides with the RMF.

\subsubsection{Slave}
The Archaean aged Slave craton shows typical lithospheric craton signatures with higher than average velocities. 
It shows in the entire region a positive velocity gradient with a maximum of ca. $4.8km.s^{-1}$ associated with a maximum of $\xi$ at ca. $150km$ deep.
Deeper, a negative gradient is observed with a minimum of ca. $4.6km.s^{-1}$ at ca. $250km$ deep. 
The topography of the extrema appears flat. (Est ce que c'est une characteristique d'un craton stable non deforme?)
Although the Slave lithosphere shows highest velocities at $150km$, above $100km$ velocities are slightly higher than the continental average ($dln(V_S)< 4\%$).

\subsubsection{Superior Craton}
In our model, the Superior craton can be split in two parts: The western part showing typical cratonic structures while the eastern part shows slower velocities at all depths. Below $150km$, velocity anomalies are below $2\%$ (almost $0\%$).

\paragraph{Western part}
The Archaean aged Superior craton shows typical lithospheric craton signatures with higher than average velocities. 
It shows in the entire region a positive velocity gradient with a maximum of ca. $4.8km.s^{-1}$  associated with a maximum of $\xi$ at ca. $150km$ deep.
Deeper a negative gradient is observed with a minimum of ca. $4.6km.s^{-1}$ at ca. $250km$ deep. 
Opposed to the Slave craton, the topography of the extrema is variable. 
From SW to NE, the core of maximum velocity can reach beween 70 and 150km of thickness. 
Beneath the maximum velocity region of ca. $150km$ the negative gradient appears stipper than beneath the $70km$ one, while the topography of the deeper positive gradient appears flat. 
At the south-eastern part of the western Superior, close the Mid-Lithospheric rift, we can observe the thinning of the cratonic lithosphere when crossing the Grenville front. 
The band of minimum velocity within the cratonic root has similar amplitudes, in terms of velocity, than beneath the younger lithosphere but at shalower depths between $100$ and $150km$. 
Folowing the Great Meteor track built by \cite{heaman2000timing}, which crosses the Superior craton, it is interseting to note that the low velocity channel beneath the base of the lithosphere shows lower velocties than usualy observed beneath the other Archaean cratons. 
This slow anomaly corresponds to the V-shape dent, or divot, observed by \cite{lee1997upper} who explained the feature by the presence of water in the mantle. 
We observe at $250km$ that the slow anomaly go further inland and crosses the Superior Craton northward until the Hudson bay following the GMT.

\paragraph{Eastern part}
Compared to its western part, the structure here shows differences. 
Fast velocities are present down to $150km$deep but are slower. 
The negative gradient is steeper and the minimum is radially thicker. 
However at the center, the minimum channel is not continuous and shows higher (still slower than the center of the western Superior) velocities.
% the gradual trend in velocity is replaced by an almost constant slow anomaly from $150$ to $300km$ deep.

\subsubsection{Trans Hudson Orogen}
The Trans Hudson Orogen (THO) is a belt of Paleo-Proterozoic aged region. It refers to a major tectonic event (between $2.0$ to $1.8Ga$) that welded Archaean blocks to form the Laurentia Craton.
% At shallow depths ($ < 60km$) shows lower velocities than archaean surroundings. 
It is interesting to notice that at shallower depths than $100km$, the proterozoic THO (including the Sask craton) which welded the Superior and Rae/Hearne cratons, shows lighter (i.e, $+4\%$ compared to $+8\%$) positive anomalies than its archaean surrounding. While most of the proterozoic root start to thin out at $150km$, the THO at the vicinity of the hudson bay shows as high anomalies as Rae/Hearne and Superior areas.

\subsubsection{Yavapai Mazatsal region}
Yavapai-Mazatzal orogens from 1.8 to 1.6 Ga represent zones of accretion for Proterozoic material and mark the amalgamation of the archean cores. \citep{hoffman1988united} \\
West to the RMF, the region shows low velocities influenced by the active tectonic with a radially positive gradient. 
The north eastern part shows velocity anomalies slightly lower than the Archaean nucleus. Velocity maxima of $4.8km.s^{-1}$ are located shalower than $100km$. 
Below, the radially negative gradient is observed and shows the same characteristic than the eastern Superior craton. It is interesting to note that, beneath Nebraska and Kansas, a high velocity anomaly of about $4.7km.s^{-1}$ is present until ca. $170km$ (while it is $120km$ in most of the region) where cretaceous aged kimberlites are sampled. 
These kimberlites indicate a lithosphere thickness of about $160km$ and have been affected by asthenospheric upwelling related to the Mid-Continent Rift System. \citep{griffin2004lithosphere}

% Basal drag from Kaban 2015 cannot be visible from temperature sensitive shear waves model, we dont see it but it is normal.

\subsection{Surrounding oceanic}
Our inversion is focusing on the continental part, as shown in our spatial parameterization (i.e. see figure...) where the region is sampled by data.
Ecxept the oceanic borders of NA, further away, the structure has not been updated from SEMucb\_wm1 \citep{french2014whole}.

\subsubsection{Eastern Passive margin}
As in \cite{schaeffer2014imaging,yuan2014lithospheric,james2014rayleigh,silveira2002anisotropic,mocquet1990three}, we observe along the negative anomaly structure following the Appalachian continental shelf, a stretched positive anomaly structure from the Suwanne terrane in Florida to the Avalonian terranes of New-Scotland. 
At $60km$ it is made of several blocks of at most $+3\%$ anomalies emdedded in a $+2\%$ structure. At $150km$, the structure is split into 2 blocks by a slow anomaly comprising both the Great Meteor (GMT) and Bermuda tracks, and further southeast, the Low Velocities Fingers (LVF) observed in \cite{french2013waveform}.

Such structure would correspond to late proterozoic gondwanian continental terranes that formed during the Pan African orogenesis \citep{kennedy1964structural} and remained attached to the North American margin during the opening of the Atlantic ocean. \citep[e.g.][]{thomas2006tectonic}
On the continental shelf, authors \citep[e.g.][]{o1983avalon,nance2002cordilleran} identified the west Avalonian terranes of pan-african affinity as in Florida \citep[see][]{smith1982review} that might have been underlayed by a thin (ca. $100km$) lithosphere before the Pangaea assembling. \citep{mckenzie2015lithospheric}

Fast anomalies are observed beneath the whole gulf of Mexico down to $150km$ deep. Despite the proximity with pan-african lithosphere, such high velocity anomalies represent older than $ca. 150Ma$ oceanic lithosphere. \citep{muller2008age,pindell2009tectonic}
-- No Magnetic anomalies recorded as in Pan African, Oceanic seafloor according to Pindel --

% \subsubsection{Bermuda low velocity channel} 
% A low velocity anomaly is observed within the Atlantic ocean plate beneath the bermuda hotspot at $60km$ deep. Deeper, the slow anomaly broadens and is elongated along the absolute plate motion from the atlantic ridge to the Mid-Continental rift. This feature, parrallel the Absolute Plate Motion (APM) corresponds to the Low Velocity Fingers (LVF) introduced by \cite{french2013waveform} and it shows highest negative anomalies between $100km$ and $200km$ and seems to vanish below $250km$.

% While in \cite{french2013waveform}, the North Atlantic LVF was stopped at the shelf of the US, our model shows that the slow velocity anomaly goes further through the continental lithosphere with a contrast of $-2\%$ to $-3\%$ compared to the surrounding cratonic root.

% En regardant mieux les cartes, je ne suis pas sur a 100\% Le LVF semble se diriger vers Bermuda. Il y a une faible anomalie qui fait le lien vers GMT mais latendance general de l'anomalie GMT Bermuda LVF ne semble pas etre la meme chose. Je prefere finir la partie GMT et comprendre les mecanismes, lire plus de biblio lies aux LVF et faire les liens. (Pour aller plus loin que Huaiyu2014 et Scaeffer2014)

\subsubsection{The Great Meteor track}
Our model images a slow velocity anomaly in agreement with the reconstruction of the Great Meteor Track proposed by \cite{heaman2000timing}.
Below $100km$ the pan african lithosphere is cut in two by a slow velocity anomaly beneath the New England seamounts (youngest part of the GMT). Further East, the slow anomaly follows the GMT crossing the continental shelf, Appalachian and Grenville orogens down to $200km$. 
This slow anomaly corresponds to the V-shape dent, or divot, observed by \cite{lee1997upper} who explained the feature by the presence of water in the mantle. 
We observe at $250km$ that the slow anomaly go further inland and crosses the Superior Craton northward until the Hudson bay following the GMT.

\cite{heaman2000timing}, based on $U-Pb$ perovskite age determinations of kimberlites, proposed a progressive southeastward younging of kimberlite magmatism throughout much of eastern North America. Considering it as a strong support for small volume mantle melting that have occurred along the continental portion of the Mesozoic Great Meteor mantle plume hotspot track initiated during the opening of the North Atlantic Ocean at about $200Ma$. \citep[See figure 4. in][]{heaman2000timing}

\cite{davies1994thermomechanical} described a process of thermo-mechanical erosion which involves heating-softening of the lithospheric base by a plume tail, followed by mechanical removal of the material due to convective transfer. \citep{rondenay2000lithospheric}
% This hypothesis is based on different points:
% \begin{itemize}
% \item The lithosphere is moving above a fixed mantle where plume originate \citep[e.g.][]{morgan1972deep}
% \item Age progression along seamounts \citep[see for example][]{crough1980kimberlites,morgan1983hotspot,duncan1984age,sleep1990monteregian}
% \item Concordance between kimberlite ages and the continental drifting rate of North America. \citep{crough1980kimberlites,duncan1984age}
% \item Evidence of hotspot related uplift along both continental and oceanic structures. \citep{crough1981mesozoic,sleep1990monteregian}
% \item Similarities in initial Sr and Nd isotopic ratios between land and sea portions. (Foland et al., 1988)
% \end{itemize}

% Continuity: 
% Chercher anomalie de flux thermique (Jaupart 2000 2002 2014) ne montrent rien
% Parler de l'hypothese rifting of the continental plate qui est en opposition avec Mantle plume.
% -- Lien avec Vs (effet thermique/chimique/fabrics) ou il faudrait prendre en compte les effets anharmonicity/anelasticity. Voir Ebinger et Sleep 1998
% -- Haydar voit une anomalie negative de Q dans la region Great Meteor + Finger!! 
% -- Trouver model de conductivite, anomalie de Gravi? Pour appuyer 
% -- Faire le lien avec Kaban 2015 Byriol 2016 Poritt 2014
% -- Fingers From SEMum2 se dirigent plus vers le Bermuda hotspot que GMT, Meme si il y a un lien entre les 2 a 150km.
% -- Aucune update du Model de GMT depuis Haeman 2000...
% This feature, parrallel the Absolute Plate Motion (APM) corresponds to the Low Velocity Fingers (LVF) introduced by \cite{french2013waveform} and it shows highest negative anomalies between $100km$ and $200km$ and seems to vanish below $250km$. While in \cite{french2013waveform}, the North Atlantic LVF was stopped at the shelf of the US, our model shows that the slow velocity anomaly goes further through the continental lithosphere with a contrast of $-2\%$ to $-3\%$ compared to the surrounding cratonic root. This slow anomaly corresponds to the V-shape dent, or divot, observed by \cite{lee1997upper}.

% Some authors [Sykes,1978; Croughet al., 1980] have suggested a genetic relation between
% kimberlites and hotspot tracks, here defined as lin-
% ear, anorogenicvolcanic chains mapped over oceanic
% and young continental lithospheres. The termination
% of these features over thickening continental roots has been cited as evidence supporting the origin of hotspot tracks in the interaction of fixed mantle plumes with an overriding plate [Morgan,1972;Burke and Wilson,
% 1976; Crough et al., 1980; Sleep, 1990a; Duncan and
% Richards, 1991], although compelling arguments may also be made for association of hotspot tracks with an-
% cient rifts and subvertical lithospheric discontinuities [Curtie,1976;Sykes,1978;Anderson1,998

% Hotspot tracks preserve records of past plate motion and provide markers with which the relative motion between a plate’s surface and underlying mantle regions may be examined

% Quand tu vas parler de l'implication thermale et chimqieu il faut citer le boulot de Haydar avec son anomalue negative de Q.

% At least two Mesozoic hotspot tracks have been postulated for eastern North America; the Juras- sic Verde and Cretaceous Great Meteor tracks [38^40,42,43]. According to Morgan [39], these two parallel and closely spaced hotspot tracks passed through Ontario and the New England region at slightly di¡erent times (c. 40 Myr apart). The location of the continental portion of these tracks is based on several lines of evidence includ- ing: (1) the current position of Atlantic Ocean hotspots; (2) the record of ocean £oor magnetic anomalies; (3) the assumption that the position of hotspots remains relatively ¢xed on Earth with time; and (4) the relative motions of North Amer- ica, South America, and Africa during the various stages recorded in opening of the Atlantic Ocean since about 200 Myr [50]. In the case of these Atlantic hotspots, the initial portion of the track typically begins on the continent and eventually moves toward the current oceanic position of the corresponding mantle plume.

\subsubsection{Gulf of Mexico}
Fast anomalies are observed beneath the whole gulf of Mexico down to $150km$ deep. Despite the proximity with pan-african lithosphere, such high velocity anomalies represent older than $150Ma$ oceanic lithosphere.

 % \subsection{Previous models}
 % Since \cite{grand1984upper}, one well established feature within NA lithosphere is the contrast between the high velocity craton and the low velocity western active part. This feature is observed in many global models (e.g., \cite{megnin2000three}, \cite{shapiro2002monte},  \cite{panning2006three}, \cite{houser2008shear}  \cite{nettles2008radially}, \cite{kustowski2008anisotropic}, \cite{ritsema2010s40rts}, \cite{simmons2010gypsum}, \cite{lekic2011inferring}, \cite{debayle2012global}, \cite{schaeffer2013global}, \cite{french2014whole}) where the cratonic lithosphere is caracterized by higher than average (ca. $4.8 km.s^{-1}$) velocities down to at least $200km$ deep.

 % Measurements of heat flux and heat production over North America ( e.g \cite{jaupart2014building} ) showed low heat flux (i.e $30-50 mW.m^{-2}$) compared to surrounding area, with a contribution at the base of the archean lithosphere brought by convection and a small internal contribution from heat producing elements (e.g \cite{michaut2007transient}). This indicates that the cratonic lithosphere that have cooled down by conduction through time can reach a $400km$ thickness and is cold (\cite{rudnick1998thermal} \cite{s2005archean}).

% Preservation of the keel through time:
% - depth 400-500km incompatible with sismo and thermobarometric models.
% - Diamond inclusions in kimberlites and Re–Os isotope analyses of peridotite xenoliths indicate that xenoliths with Late-Archean ages erupted from depths up to 200km.
% 	-  If we assume that these xenoliths represent samples of the cratonic keel, this suggests that cratonic keels formed soon after (or along with) the craton and that subcratonic lithosphere was cold as early as the Late-Archean
% 		- Thus, cratonic keels would have appeared to have formed soon after the craton formed and have moved with the craton, relatively undeformed, through time. 
% 		- However, recent evidence suggests that Re–Os ages may over-estimate the age of xenoliths because zircon grown, and presumably diamond, does not reset the Re–Os system. It is difficult, at best, to preserve the diamond stability conditions for 3 Gyr, even in a stable continental keel. Perhaps the apparent ages are over-estimates; this would eliminate the need to maintain stable, low temperature keels for time periods up to 3 Gyr in the dynamical modeling.

% Jordan's tectosphere
% root of the cratonic lithosphere have remained unmixed with the rest of the upper mantle.
% 	Problem: No positive geoid anomaly observed so no mass excess
% Roots composition must be different in term of composition and iron depleted
% Chemistry of the tectosphere is mechanically strong and buoyant as the submantle but chemically distinct
% The effect of composition offsets the increase in density that results from cooling, leading to the isopycnic hypothesis, cratons and keels are neutrally buoyant.
% 	Problem: Neutral buyancy is not sufficiant to stabilyse cratonic keels from the shearing of mantle flow.
% 			Thermal conduction is a slow process (for the cooling), it is unlikely that compositional process can balance thermal cooling through time that would lead to neutral buoyancy.
% 			Geodynamical models cant maintain a stable cratonic root through time.

% geochemistry measures
% 	Cold and dehydrated keels appear to be residual from varying degrees of partial melting over a range of pressure {}with some post depletion metasomatic activity (King 2005).

% 	Thermobarometric studies indicate that until ca. $175km$ the mantle is chemically depleted. \cite{kopylova1999petrology}

% 	Beneath 175km peridotite more fertile: Top of the convective mantle or archean oceanic crust undeplating craton or metasomatic activities... Re-Os datattion on eclogites showed that they are older than some peridotites -> keels are subducted archean ocean crust. \cite{shirey2001archean}

% 	A number of petrological studies confirm the relative buoyancy of xenoliths thought to represent the composition of the keel [4,24,51–53]. Reconciling the petrological models of root composition with the seismic velocities remains elusive (e.g., [20,24]). Dehydration can lead to several order of magnitude increase in viscosity, relative to mantle at the same pressure and temperature conditions [54,55]. Hence if the cratonic keels represent the depleted residual from a large melting event, this would be consistent with the buoyant and viscous properties that are needed to maintain stable keels over a long time period.

% Electrical conductovity:
% The bootom of upper mantle is wetter than above beneath craton and dryer than ocean until 250km \cite{yang2015three} \cite{hirth2000comparison}
% Jordan [23] hypothesized that subcratonic keels or
% roots are attached to the cratonic lithosphere and that
% these roots or keels have remained isolated from and
% unmixed with the rest of the mantle since the
% formation of the craton. If the cratonic root is a thick
% thermal boundary layer and the subcratonic mantle is
% average upper mantle composition, there should be a
% 10-m positive geoid anomaly over cratons [29].
% Because there is no correlation between continents
% and the geoid [30–32], the subcratonic mantle must be
% compositionally distinct from the surrounding upper
% mantle. Jordan called this subcratonic material tectosphere
% to distinguish it from normal upper mantle. In
% the tectosphere hypothesis these keels are chemically
% distinct from the convective mantle and are both
% compositionally buoyant and mechanically strong.
% The effect of composition offsets the increase in
% density that results from cooling, leading to the
% disopycnicT hypothesis [33]—cratons and keels are
% neutrally buoyant. Numerical studies have clearly
% demonstrated that neutral buoyancy is not sufficient to
% stabilize cratonic keels from the shearing of mantle
% flow. Thermal conduction is a slow and continuous
% process and it seems highly improbable that a
% compositional process would exactly balance thermal
% cooling through time, leading to neutral buoyancy at
% all depths envisioned in the isopycnic hypothesis.
% While both strength and buoyancy act to stabilize
% keels in theory, it has been quite challenging to
% produce geodynamical models that maintain stable
% cratonic keels in a convecting mantle over several
% billion years of mantle evolution (e.g., [3] and see
% later section on craton stability).

% In the steady state thermal conditions that prevail in old continents, the man- tle heat flux, i.e., the heat flux at the Moho, is the sum of the heat flux at the base of the lithosphere, which is brought by mantle convection, and heat generated by heat producing elements in the lithospheric mantle. The first contribution originates deeper than 200 km and potential lateral variations would get smoothed out by horizontal diffusion. The contribution of heat producing elements in the lithospheric mantle is very small, as shown by xenolith studies [Michaut et al., 2007], and horizontal anomalies would also be smeared out. Joint analysis of surface heat flux data and seismic wave velocity variations confirms that Moho heat flux variations are constrained to be less than about ± 3 mW m−2 , so that the total variation is 6 mW m−2 across the province [Perry et al., 2006b; Lévy and Jaupart, 2011].

 % \subsection{Craton caracteristic}

 % Through times, Lauretia craton remained undeformed while being part of the earliest supercontinent Nuna-Columbia \citep{meert2012whats}, then Rodinia and finally Pangaea.

% Archean part more buoyant than proterozoic one
% The archean province may reflect a relatively depleted underlying mantle lithosphere.
% The presume secular decline in mean temperature of the asthenosphere impliesa greater depth and volume of melting accompanying mantle upwelling in the Archean, and consequently thicker, more depleted mantle lithosphere.

% Most importantly, the 2.7 Ga age of the Greenland eclogites supports the observation by Shirey and Richardson (2011) that mantle eclogite materials older than 3.0 Ga appear absent from the cratonic record. They suggest that the appearance of cratonic eclogites in most cratons after 3.0 Ga marks the onset of the Wilson Cycle of plate tectonics, when more frequent subduction and collision events became the most prominent continental growth mechanism. (Tappe et al., 2011)

% Previous studies
% Link with upper mantle tomography
% layering and radial intro to aximuthal

% In the early 1960s, theoretical work (e.g. Anderson 1961) on the dispersive properties of transversely isotropic media suggested that radial anisotropy in the upper mantle could have a pronounced ef- fect on the shape of surface wave dispersion curves and could lead to an apparent discrepancy between Love and Rayleigh wave data.

% For  more  than  thirty  years,  seismologists  have  imaged  the  earth's  interior  using  seismic
% waves generated by earthquakes,  and traveling through di erent structures of the planet.
% A  remaining  challenge  in  seismology  is  to  interpret  the  recovered  Earth  models  in  terms
% of physical properties (e.g.  temperature, density, mineral composition) that are needed for
% understanding  mantle  dynamics  and  plate-tectonics.   For  example,  a  region  of  slow  wave
% speed can be either interpreted as anomalously warm, or rich in water, or iron.
% Although seismic waves are sensitive to a large number of visco-elastic parameters as well
% as density, the mantle models constructed from seismic tomography are only parameterized
% with a few physical parameters, for example average isotropic shear wave velocity and radial
% anisotropy (e.g. French et al., 2013). This is because given the available information observed
% at the surface, there is not enough resolution to entirely describe the local elastic tensor.  In
% addition to the limitednumber of resolvable elastic (and anelastic) parameters, there is also
% the question of spatial resolution, namely the smallest spatial scale at which heterogeneities
% can be imaged.
% The  number  of  independent  elastic  parameters  that  can

% \section{Methodology}
% % \subsection{Kernels}
% In the continuity of previous work (e.g Marone et al.,2007; Yuan et al.,2011; Yuan et al., 2014) we invert 3-component fundamental and overtone surface waves in the time domain to image North American upper mantle 3D structure using an hybrid inversion scheme. \\
% Inferring the elastic structure of the earth from the analysis of propagating waveforms is a well-studied inverse problem in geophysics. 
% % A non linear inverse problem that can be solved iteratively, where the accuracy of the different solution's terms may come along with the expensiveness of their calculation. 
% Our inverse problem is solved iteratively following the generalized least-squares formalism of Tarantola and Valette, 1982. We define a discrete L2 misfit functional:

% \begin{equation}
% 2\Phi(m_{k}) = [d - g(m_{k})]^{T} C^{-1}_{d} [d-g(m_{k})] + [m_{p}-m_{k}]^{T} C_{m}^{-1} [m_{p}-m_{k}]
% \end{equation}

% where \ensuremath{m_{k}} represents the \ensuremath{k^{th}} iterative model estimate, \ensuremath{d} and \ensuremath{g(m_{k})} are the observed data and data predictions (time-discretized waveforms and period-discretized group velocities), \ensuremath{m_{p}} is the model prior (1D earth or 3D starting model), and \ensuremath{C_{m}} and \ensuremath{C_{d}} represent a priori model and data covariance operators respectively.

% We solve for successive updates to the current model estimate \ensuremath{m_{k} \to m_{k+1}} following the update equation that minimizes $\Phi$, namely:
% \begin{equation}
% m_{k+1} = m_{k} + (C_{m}G^{T}_{k}C^{-1}_{d}G_{k} + I)^{-1} (C_{m}G_{k}^{T}C^{-1}[d-g(m_{k})]+ m_{p}-m_{k})
% \end{equation}
% We employ a hybrid waveform inversion scheme, in which “exact” forward modeling of the wavefield (i.e. the waveform elements of $g(m_{k})$ is achieved using the spectral element method (SEM: e.g. Komatitsch and Vilotte, 1998; Komatitsch and Tromp, 2002a,b), while approximate sensitivity kernels (i.e. elements of G) are calculated using non-linear asymptotic coupling theory (NACT: Li and Romanowicz, 1995).

% NACT is a normal-mode perturbation approach, which takes into account coupling between modes both along and across dispersion branches, allowing the computation of 2-D broad-band sensitivity kernels which more rigorously reproduce the sensitivity of body waveforms to structure along the ray geometrical path in the vertical plane containing the source and the receiver.
% However, when the distance between a source and a station is below $15^{o}$, NACT breaks down and we instead use the more standard Path Average Approximation (PAVA). Developed by Woodhouse \& Dziewonski, 1984; PAVA kernels are sensitive to the average structure between the source and the receiver and along the great circle path, that is, the corresponding sensitivity kernels are ‘1-D’ in the vertical plane containing the source and the receiver. This approximation is best for single-mode seismograms, such as fundamental mode surface waves (Romanowicz, 2008).

% PAVA/NACT do not consider off great circle path sensitivity in the horizontal direction, which gives rise 
% to focusing and defocusing effects (e.g. Zhou et al. 2005). These effects are important for amplitude 
% fitting. Here, we are, however, primarily concerned with fitting the phase and for that, the 2-D effects in
% the vertical plane are dominant, especially for overtones. The dominance of 2-D effects in the
% vertical plane can be assessed by considering an asymptotic expansion of normal mode first order
% perturbation theory, which shows that 2-D coupling effects in the vertical plane are of zeroth 
% order, while focusing effects are of higher order (e.g. Romanowicz, 1987; Li \& Romanowicz, 1995; Romanowicz, 2008).

% On the other hand, we use the spectral element method to accurately compute the forward seismic wavefield $g(m_{k})$. 
% Lekić \& Romanowicz., 2011 showed, that using approximate methods, such as normal mode perturbations theory, to calculate synthetic seismograms can introduce significant errors in the modeling of the wavefield in a heterogeneous 3-D earth. These errors can be seen as additional noise in the inversion and may introduce some model dependent errors. In that situation, convergence of the inverse problem can be compromised. We use a modified version of the spectral element code RegSEM (Cupillard et al., 2012) which initially was designed to calculate synthetic seismograms in a $90^{o} \times 90^{o}$ area within which both receivers and event sources are located. RegSEM takes into account effects of the oceans, topography / bathymetry, ellipticity, gravity, rotation and anelasticity. It also includes Perfectly Matched Layers (PML) absorbing boundary conditions, which allow us to restrict the computation laterally and radially to the region of interest with the vanishing of spurious reflections at the borders. (Patera \& Vilotte, 2005)

% Using only regional events is an opportunity to image the North American upper mantle without the influence on the wavefield of the global 3-D structure. However the current data coverage from a pure regional data set is not sufficient to constrain our region. To address this problem, we chose to include teleseismic events and developed a new method in order to calculate teleseismic synthetics where, through iterations, the wavefield is calculated only in the region of interest by the use of secondary sources.
% Developed by Masson et al., 2013, the idea is to construct an equivalent of time-reversal mirrors (Fink et al., 1989) at the boundary of our region of interest where the wavefield is then re-generated forwardly. 
% Time reversal is a two step process. First, waves propagating in a medium from a source are recorded thanks to an array of transducers (also named time-reversal mirrors). Second, after having reversed in time our records, transducers emit back into the medium toward the source where the ernergy is refocused in time and space. The feasability of such procedure relies in the reciprocity in space and reversibility in time of the wave equation. (Robertson \& Chapman, 2000)

% % As in Robertson \& Chapman, 2000 we use a two step method. The first step performs wave propagations in the global medium, then as a second step the wavefield is used as an input (time-reversal mirrors) to obtain the wavefield into the medium under study.

% In the elastic case, the ideal response of a time-reversal mirror can be expressed in the frequency domain with the use of the Helmotz-Kirchhoff representation theorem:
% \begin{equation}\label{eq:repres}
% \begin{split}
% 	u_{i}^{*}(\bm x) =&  \int_{V} G_{in}(\bm x, \bm x')f_{n}^{*}(\bm x')dV' \\
% 	 +& \oint_{S} G_{in}(\bm x, \bm x') n_{j} C_{njkl} \partial_{k}^{'} u_{l}^{*}(\bm x') dS' \\
% 	 -& \oint_{S} u_{n}^{*}(\bm x') n_{j} C_{njkl} \partial_{k}^{'} G_{il}(\bm x, \bm x')dS'
% \end{split}
% \end{equation}

% where $u$ is the displacement vector, $G_{in}(\bm x, \bm x')$ is the Green's tensor denoting the displacement at location $\bm x$ in the $i$ direction due to a unit point source at $\bm x'$ in the $n$ direction, $C_{ijkl}(\bm x)$ is the stiffness tensor at the $\bm x$ location, $\partial_{k}^{'}$ stands the derivative with respect to the $x_{k}^{'}$ coordinate and the star symbol (*) stands for the complex conjugate. Eq.\ref{eq:repres} shows that, if the exciting force $f_{n}(\bm x')$ is known throughout the volume $V$ and that the wavefield $u_{i}(\bm x)$, the associated traction $n_{j} C_{njkl} \partial_{k}^{'} u_{l}(\bm x')$ are known on the surface $S$; we are then able to regenerate the wavefield anywhere within the volume $V$. (Masson et al., 2013)

% Masson et al., 2013 proposed to introduce a spatial window function into the wave equation (original wavefield) in order to determine the exciting forces, that applied to the surface $S$, will generate the response $u_{i}(\bm x)$ defined in equation \ref{eq:repres}.

% Starting with the wave equation in the frequency domain:

% \begin{equation}\label{eq:wavefreq}
% 	\rho \omega^{2} \bm u_n + \partial_j (C_{njkl}\partial_k \bm u_l) = -f_n,
% \end{equation}

% the generated wavefield in $M \in V$, is defined as:

% \begin{equation}\label{eq:um}
% 	\bm u^M (\bm x ,t) = \bm u(\bm x ,t)\omega(\bm x).
% \end{equation}

% with the use of a bi-valued function $\omega$:

% \begin{equation}\label{eq:window}
% 	 \omega(\bm x) = \left\{\begin{array}{ll}
% 						1 	&\mbox{ for all \quad $\bm x \in M$} \\
% 						0   &\mbox{ for all \quad $\bm x \notin M$}
% 						\end{array}
% 						\right.
% \end{equation}

% such as:

% \begin{equation}\label{eq:um_in_u}
%  \bm u^M (\bm x ,t) = \left\{\begin{array}{cl}
% 						\bm u (\bm x ,t) &\mbox{ for all \quad $\bm x \in M$} \\
% 						0                &\mbox{ for all \quad $\bm x \notin M$}
% 						\end{array}
% 						\right.
% \end{equation}

% where $M$ is the regional subvolume bounded by $\delta M$ within the global volume $V$.

% The mirror's excitation force that generates $\bm u^M$ is introduced as:
% \begin{equation}\label{eq:exc_force_reg}
%  	f_n^M = -\rho \omega^2 \bm u_n^M - \partial_j (C_{njkl}\partial_k\bm u_l^M)
%  \end{equation}

%  Using eq. \ref{eq:wavefreq} - \ref{eq:exc_force_reg} and after developing:
%  \begin{equation}
%    	f_n^M = \omega f_n - C_{njkl} \partial_{j}\omega \partial_k u_l - \partial_j(u_l C_{njkl} \partial_k \omega).
%  \end{equation}

%  Finally, convolving eq. \ref{eq:exc_force_reg} with the Green's function of the medium and then integrate over the volume (see Masson et al., 2013 for a detailed derivation) gives :

%  \begin{equation}\label{eq:umfinal}
%  \begin{split}
%   	u_{i}^{M}(\bm x) =& \int_{M} G_{in}(\bm x , \bm x{'})f_{n} (\bm x') dV' \\
%   					 +& \oint_{\delta M} G_{in} (\bm x, \bm x') n_{j} C_{njkl} \partial_{k}^{'} u_l(\bm x')dS' \\
%   					 -& \oint_{\delta M} u_{n}  (\bm x') n_{j}   C_{njkl} \partial_{k}^{'} G_{il}(\bm x, \bm x')dS'
%  \end{split}
%  \end{equation}

% In practice, one global simulation per event is needed to generate such mirrors, they are then used as secondary sources to generate the wavefield in our region of interest. These secondary sources are calculated once and for all. The main advantage of using such a method lies in the fact that we update the elastic structure of the region iteratively, we only need to calculate synthetics in the region. Since we're not updating the global structure (following the zeroth iteration ) of the earth but only North America, this allow us to save significant computation time and therefore attain shorter periods.

% The hybrid inversion scheme represents a compromise between speed and accuracy. The use of exact numerical methods to calculate the wavefield remains expensive, while waveform partial derivatives computed using NACT are approximate and inexpensive. Despite the approximate nature of NACT, we point out two important properties: sensitivity kernels, that are recalculated at each iteration in the current 3-D model) accommodate non-linearity due to multiple forward scattering (Romanowicz et al., 2008); and it includes finite-frequency effects in the 2D great-circle plane, critical to modeling surface-wave overtones and long-period body waves (Li and Romanowicz, 1995).

% % \subsection{Effect of the crust}
% % From Scott's thesis 
% % Because SEM simulations are so precise, we require a model of crustal structure that accurately predicts the seismic response of the earth’s crust – particularly for highly sensitive fundamental- mode surface waves. Indeed, fidelity to crustal effects is especially critical for accurate retrieval of anisotropic mantle structure, as often noted when discussing the shortcomings of standard linear crustal corrections (Bozdag and Trampert, 2008; Lekic ́ et al., 2010; Ferreira et al., 2010). Here, we adopt the smooth crustal mod- eling technique proposed in Lekic ́ and Romanowicz (2011) and later refined by French et al. (2013). The thin layers of the earth’s crust slow down SEM computations considerably, particularly in the oceans. To mitigate this effect, we construct models of the crust that contain no interior discontinuities and are potentially thicker than the real crust in some regions. These smooth mod- els are calibrated to mimic the seismic response of the earth’s crust, as seen through observations of surface-wave group-velocity dispersion, by introducing crustal radial anisotropy (Backus, 1962; Capdeville and Marigo, 2007).
% % In the continents, our smooth crust matches the Moho depth of Crust2.0 (Bassin et al., 2000), with slight smoothing to avoid aliasing in the SEM mesh, while Moho depth saturates at 30 km in oceanic regions. A similar technique, but calibrated to mimic the dispersion of an a priori crustal model and not restricted to match continental Moho topography, has previously been shown to be effective for regional-scale waveform inversions (Fichtner et al., 2009). The disadvantages of such a technique are: (a) that the crustal model is only valid over the period range it is calibrated, in this case periods at or above 25 s, and (b) susceptibility to noise in the dispersion data upon which it is calibrated. The first concern should not be an issue for the 40 s and 60 s SEM modeling em- ployed here, confirmed in the global and North-American region validation simulations reported by French et al. (2013). To address the second concern, these authors produced ensembles of crustal models calibrated on dispersion data with realistic-amplitude syn- thetic noise added, and examined variation in the resulting SEM synthetics, concluding that such errors could not significantly af- fect long-period waveforms. Thus, we are confident that the crustal modeling approach is able to attain SEM speedup, while maintain- ing fidelity to the observed surface-wave dispersion characterizing crustal structure, and without negatively affecting the resulting long-period fundamental and higher mode waveforms and the re- sulting upper mantle shear velocity and radial anisotropy structure.

% \subsection{Parameterization}
% Propagation of seismic waves through an arbitrary Hookean medium depends on 21 parameters of the stiffness tensor,
% and inferring the values of all these parameters at all locations within the mantle is not feasible with available seismic data. 
% However, by approximating the Earth as a transversely isotropic medium with a vertical axis of symmetry ( VTI ), we can reduce the number of free parameters,
% while capturing the first order observation that horizontally polarized surface waves travel, on average, faster than vertically polarized ones (e.g. Anderson 1961; McEvilly 1964).

% A VTI medium is described by density \ensuremath{\rho} and the five Love coefficients (\ensuremath{A}, \ensuremath{C}, \ensuremath{F},
%  \ensuremath{L} and \ensuremath{N} ) (Love 1927). 
% We parametrize our model in terms of \ensuremath{\rho}, the Voigt average isotropic shear ($V_S$) and compressional ($V_P$) velocities: \\
% \begin{align}
% V_{S} =& \sqrt{\frac{2V^{2}_{SV}+V^{2}_{SH}} {3}} \\
% V_{P} =& \sqrt{\frac{V^{2}_{PV} + 4V^{2}_{PH}}{5}}
% \end{align}

%  and the anisotropic parameters:

%  \begin{equation}
%  \xi = \frac{N}{L}, \quad \varphi = \frac{C}{A}, \quad \eta = \frac{F}{A-2L} 
%  \end{equation}

% Love and Rayleigh waves are primarily sensitive to shear wave structure, whereas their sensitivities are significantly weaker for other elastic parameters (Dahlen \& Tromp 1998). 
% Thus, we decrease the number of parameters of interest by choosing not to invert for lateral variations in the poorly constrained $V_{P}$ $\varphi$, $\rho$ and $\eta$ parameters. 
% Instead, we parametrize the elastic structure of the mantle in terms of $V_{S}$ and $\xi$.
% The parameter $\eta$ governs the variation of wave speed at directions intermediate to the horizontal and vertical. 
% When $\eta$ and $\varphi$ are approximately equal to one, which is very likely the case in the mantle, we can approximately relate Voigt average velocities to those of vertically and horizontally polarized waves.
% Following Montagner \& Anderson, 1989 we impose an a priori information on other parameters perturbations by scaling them with inverted parameters:

% \begin{equation}
% \frac{\delta(V_{P})}{\delta(V_{S})} = 0.5, \quad
% \frac{\delta(\rho)}{\delta(V_{S})} = 0.33, \quad
% \frac{\delta(\eta)}{\delta(\xi)} = -2.5,   \quad
% \frac{\delta(\varphi)}{\delta(\xi)} = -1.5,   \quad
% \end{equation}

% We opt for $V_{S}$ and $\xi$ parameterization instead of the more traditional $V_{SV}$ and $V_{SH}$, to avoid additional uncertainties on the amplitude and possibly even sign of the anisotropic parameter $\xi$ introduced by damping in the inversion process. (Marone \& Romanowicz, 2007)
% Indeed, if $\xi$ is computed after the inversion from the derived $V_{SV}$ and $V_{SH}$, its value can be biased by the different effect of damping on the two velocity parameters. ( Panning \& Romanowicz; 2006).

% In theory, the retrieved radial anisotropic model could be biased by the presence of azimuthal anisotropy, neglected at the present stage of this study. In practice, with sufficient azimuthal coverage, this possible effect is reduced because the azimuthal variations will be averaged out at each point of the model. This study, limited to radial anisotropy, should be regarded as a first step towards a more complex model of anisotropy, where the assumption of a vertical symmetry axis will be dropped and which will, therefore, still be characterized by radial anisotropy but with a symmetry axis with an arbitrary orientation (e.g. Montagner \& Nataf 1988).\\

% From the Core Mantle Boundary up to the Moho, the model is expressed on 26 cubic splines $\nu_{q}(r)$ defined in Megnin \& Romanowicz (2000), though we invert for structure only in the top 16 splines; deeper structure is fixed to that of our starting model $SEMucb_wm1$ (French \& Romanowicz, 2014). 
% The knot locations are at radii: 5690, 5810, 5910, 5990, 6061, 6101, 6131, 6161, 6191, 6221, 6241, 6261, 6281, 6301, 6321, 6341 km. 
% Laterally, we parametrize our model spatially in terms of spherical splines $\beta_{p}(\theta, \phi))$ (Wang \& Dahlen 1995). 
% Thus, the value of a given model parameter $m$ at any location in the Earth $(\theta, \phi, r)$ can then be calculated from a set of spline coefficients $m_{pq}$:

% \begin{equation}
% m(\theta,\varphi,r) = \sum_{p}\sum_{q}m_{pq}\beta_{p}(\theta,\varphi)\nu_{q}(r)
% \end{equation}

% The splines are a local basis, and thus help minimize the mapping of structure in one region into structure in distant regions, which can be an undesirable effect of global parametrization such as spherical harmonics (Lekic et al., 2011) .
% % In general when parameterizing our model, we put strict a priori constraints on the minimum length scale of structure allowed in our model. This truncation results in spectral leakage (aliasing) of short scale heterogeneity into longer length scales (Trampert \& Snieder 1996), though the use of splines reduces this aliasing when compared to spherical harmonics or spherical pixels (Chiao \& Kuo 2001). To further reduce the aliasing of retrieved structure, we allow structure to vary at shorter length-scales than those that we can reasonably expect to image and interpret (Spetzler \& Trampert 2003).

% In our regional tomographic studies, the mesh for the target region can be refined by adding knots halfway between two neighboring grid points, until the desired knot spacing has been reached, while maintaining a coarser mesh throughout the rest of the globe (Marone \& Romanowicz 2007a). This approach allows flexibility: the grid spacing can, for instance, be chosen as a function of the data coverage. Therefore, we maximize the resolution within the ideally sampled region, without need for over parameterizing the areas with insufficient data density. For the $V_{S}$ parameter we use a regional level 7 spherical grid (~1 degree knots spacing) and level 6 ( ~2 degree ) for $\xi$.

% \subsection{A priori information}
% When solving an inverse problem, the importance of the a priori information depends greatly on the use one plans for the estimated model parameters.
% If one simply wants one example of a solution to the problem, the choice of a priori information is unimportant. However, if one wants to develop arguments that depend on the uniqueness of the estimates, the validity of the a priori assumptions is of paramount importance. (Menke, 2012)

% As a starting model, we opted for the global 3D radially anisotropic shear wave velocity model $SEMUCBwm1$ (French \& Romanowicz, 2014). This global whole mantle model was built in the continuity of SEMum generation models (Lekić \& Romanowicz, 2011) with an extended data set of :
% \begin{itemize}
% 	\item 3 components 60-400s periods waveforms
% 	\item 3 components 30-400s periods body-waves 
% 	\item group-velocity dispersion maps 25-150s 
% \end{itemize}

% Concerning the isotropic structure, the a priori information on which SEMum is based on is a VTI 1D earth model. The idea was to fit long-period ($T > 60 s$) waveforms (Cammarano \& Romanowicz 2007) starting from one of the physical reference models of Cammarano et al. (2005), which are calculated from a fixed composition (dry pyrolite) and a thermal profile using the elastic and anelastic properties of principal mantle minerals.

% As for the the anisotropic part, the a priori information of $\xi$ was obtained by testing several values against a set of observed frequencies of spheroidal and toroidal modes, keeping fixed the elastic structure. 
% In this process, $\xi$ was allowed to be stronger or equal to 1 (isotropic case) at depth shallower than 320 km as verified in numerous previous seismic studies (Dziewonski \& Anderson 1981).

% When solving the inverse problem, we also include a priori information on the model in a covariance matrix $C_{M}$. It specifies the expected deviation of true mantle structure from that specified by our starting model. This matrix is defined by the variance $\sigma^{2}_{0}$ (the diagonal entries) and the horizontal and vertical correlation lengths, $h_{0}$ and $v_{0}$, associated with each spline knot.

%  These correlation lengths are spatially tuned according to data-coverage density and quality. As in \cite{lekic2011inferring}, the element of $C_{M}$ jointly describing the state of a priori information on model coefficients i and j is given by:

% \begin{equation}
% C_{m} = \sigma_{ij} exp(\frac{cos\Delta_{ij}-1}{h^{2}_{ij}}) exp(\frac{-2d^{2}_{ij}}{v^{2}_{ij}})
% \end{equation}

% where $\Delta_{ij}$ is the minor-arc distance separating model coefficients i and j, $d_{ij}$ their radial separation, and $v_{ij}$ and $h_{ij}$ the averages of their associated radial and lateral correlation lengths.

% In order to increase image resolution, we allowed the horizontal length of correlations to vary between $100 km$ and $300 km$ for $V_{S}$; $300 km$ to $600 km$ for $\xi$. Thus, within the model, $h_{ij}$ can vary based upon an estimator for relative sensitivity. The idea is to take the diagonal of the approximated Gauss-Newton Hessian $G^{T}C^{-1}G$ associated with the data misfit term, to represent aggregate sensitivity of the data to each model coefficient. Contributions from different data type are weighted accordingly to the quality and uncertainty estimates appearing in the data covariance matrix $C_{D}$. These values are, in turn, used to select local correlation-length estimates by scaling to the interval between prescribed minimum and maximum values for a given model parameter.

% \subsection{data}

% In this study, long-period waveforms and group velocity dispersion maps are used to constrain shear wave velocities and radial anisotropy in both crust and upper mantle of North America.

% The group-velocity dispersion data are from Ritzwoller et al., 2002 and are in the form of $1^o \times 1^o$ maps between 25 and 150s. The use of short period dispersion curves ( below 80s ) constrained our homogenized crustal structure, while the entire period range was included during the inversion of the mantle structure.

% The long-period waveform data set comprises Rayleigh and Love accelerograms including both fundamental and higher modes recorded on 3 components broad-band seismographs. 
% Fundamental modes are primarly sensitive to the first 300km while overtones have sensitivity into the transition zone. We used 2 different data sets:

% \begin{itemize}
% 	\item a regional data set consisiting of 155 events $(4.5 < M_{w} < 6.0)$ where sources are located within the $90^o \times 90^o $RegSEM mesh.
% 	\item a teleseismic data set consisting of 122 events $(5.5 < M_{w} < 6.9)$ for which secondary sources are calculating around the $90^o \times 90^o $ RegSEM mesh using the procedure of \cite{masson2013numerical}.
% \end{itemize}

% Such data set, recorded at more than 2400 stations, provides an exellent azimuthal coverage of NA continent, with a total of 19.991 Love wavepackets, 55.176 Rayleigh wavepackets for fundamental modes, 82.809 overtones with 20 per cent from the T component and 63.179 wavepackets mixing fundamental and overtone energies.

% We filter accelerograms between 40 and 400s with corner frequencies at 53 and 250s, and separate them into wave-packets. Before inversio, the wavepackets are weighted according to their relative amplitudess, in order to avoid dominance by the larger amplitude fundamental modes in the inversion. In our weighting scheme, we also account for background noise, and path redundancy, as described in \cite{li1996global} This weighting step is crucial as it homogenizes the data coverage within the region and provides a way to construct the data covariance matrix $C_D$, which we approximate as a diagonal matrix.

\bibliographystyle{apalike}
\bibliography{/Users/Pierre/Documents/PHD/Thesis/biblio.bib}
%
\end{document}

%

### 2017-02-17 Paper_NA.tex
second commit

@@
 \renewcommand{\baselinestretch}{2}\normalsize
-
+\author{Pierre}
 \title{Paper NA}
@@
 
-{\color{red}To overcome this issue, we take advantage of the "Box Tomography" theory which allows to combine teleseismic and regional waveforms in a consistent manner to image regional targets at arbitrary locations.
+To overcome this issue, we take advantage of the "Box Tomography" theory which allows to combine teleseismic and regional waveforms in a consistent manner to image regional targets at arbitrary locations.
 This general framework accomodates for arbitrary acquisition setups (i.e.
@@
 In this situation, Box Tomography is particularly efficient as the computational effort  to account for teleseismic data in regional inversions is limited to a few global scale simulations that are done once and for all prior to the inversion.
-Masson and Romanowicz (2016b) showed that this approach can produce accurate regional tomographic images even though the elastic structure outside the imaged region is neither fully known prior to the inversion nor updated during the inversion.}
+Masson and Romanowicz (2016b) showed that this approach can produce accurate regional tomographic images even though the elastic structure outside the imaged region is neither fully known prior to the inversion nor updated during the inversion.
 
@@
 
-The dataset includes three component acceleration waveforms from more than 2400 permanent and temporary broadband stations located in the target region, a $90^o \times 90^o $ area encompassing most of north America (Fig.
-3).
+\begin{figure}[ht]
+	\centering
+	\begin{tabular}{ccc}
+		\includegraphics[width=1\textwidth]{figures/map_events_stn_global.png}
+	\end{tabular}
+
+	\caption{\baselineskip 18pt 
+	Source and station distribution for the new North American inversion.
+Black triangles show the seismic stations.
+Stars are 155 regional events, in addition to 122 teleseismic events.
+Orange stars indiscate events shallower than $30km$ deep, Yellow stars for deeper than $300km$ and red one are in between.
+Thick red line indicates the boundaries used for the RegSEM forward simulation, which extends 89◦ by 89◦ horizontally and down to 1600 km and contains all stations and regional events.}
+
+	\label{map_evt_stn}
+\end{figure}
+
+The dataset includes three component acceleration waveforms from 2860 permanent and temporary broadband stations located in the target region, a $89^o \times 89^o $ area encompassing most of north America (Figure ~\ref{map_evt_stn}).
 We considered two different datasets:
@@
 As the model improves, at each iteration, additional wavepackets are included.
-The total number of wavepackets of fundamental and overtone Love and Rayleigh waves is given in Table 1, and provides a good azimuthal coverage within the NA continent (Fig.
-4).
+The total number of wavepackets of fundamental and overtone Love and Rayleigh waves is given in Table 1, and provides a good azimuthal coverage within the NA continent (Figure ~\ref{density_coverage}).
+
+\begin{figure}[p]
+	\begin{minipage}{0.47\linewidth}
+		\centerline{\includegraphics[bb=120 10 470 380,clip=,width=\linewidth]{figures/Coverage_L_iter3.png}}
+	\end{minipage}
+	\hfill
+	\begin{minipage}{0.47\linewidth}
+		\centerline{\includegraphics[bb=120 10 470 380,clip=,width=\linewidth]{figures/Coverage_T_iter3.png}}
+	\end{minipage}
+
+	\begin{minipage}{0.47\linewidth}
+		\centerline{\includegraphics[bb=120 10 470 380,clip=,width=\linewidth]{figures/Coverage_Z_iter3.png}}
+	\end{minipage}
+	\hfill
+	\begin{minipage}{0.47\linewidth}
+		\centerline{\includegraphics[bb=120 10 470 380,clip=,width=\linewidth]{figures/Coverage_all_iter3.png}}
+	\end{minipage}
+
+\caption{Density coverage for each component and all gathered.} 
+\label{density_coverage}
+
+\end{figure}
 
@@
 
-Assuming that our azimuthal coverage is everywhere sufficient within our target region (Fig.
-4), we here solve only for radially anisotropic structure and specifically for two of the 5 parameters that describe such a structure, the two parameters to which long period surface waveforms are the most sensitive: isotropic velocity $V_{S}$ and the anisotropic parameter $\xi = {\frac{V_{SH}}{V_{SV}}}^2$, where $V_{SH}$ is the velocity of waves polarized horizontally and $V_{SV}$ that of waves polarized vertically .
+Assuming that our azimuthal coverage is everywhere sufficient within our target region (Figure ~\ref{density_coverage}), we here solve only for radially anisotropic structure and specifically for two of the 5 parameters that describe such a structure, the two parameters to which long period surface waveforms are the most sensitive: isotropic velocity $V_{S}$ and the anisotropic parameter $\xi = {\frac{V_{SH}}{V_{SV}}}^2$, where $V_{SH}$ is the velocity of waves polarized horizontally and $V_{SV}$ that of waves polarized vertically .
 The other three parameters and density are constrained through empirical scaling relationships, following \cite{montagner1989petrological}, and are based on laboratory measurements for upper mantle rocks.
@@
 
-The target model space is geographically defined as shown in Figure 3, and is limited in depth down to 800 km.
+The target model space is geographically defined as shown in Figure ~\ref{map_evt_stn}, and is limited in depth down to 800 km.
 As in our previous tomographic studies, the model space is parametrized in terms of 26 cubic splines $\nu(r)_{q}$ vertically \citep{megnin2000three} from the core-mantle boundary to the Moho, although we invert for structure only in the top 16 splines, corresponding to the top 700-800 km of the mantle.
@@
 
-\begin{figure}
-\centering
+\begin{figure}[p]
+% \centering
 	\begin{tabular}{ccc}
 		(a)&&(b)\\
-  \includegraphics[width=.45\textwidth]{snap_glob.pdf}  &&
-  \includegraphics[width=.45\textwidth]{snap_reg.pdf} \\ 
-  \includegraphics[width=.45\textwidth]{seismo.pdf}      &&
-  \includegraphics[width=.45\textwidth]{seismo.pdf}
+		\includegraphics[width=.4\textwidth]{figures/snap_glob.pdf} && \includegraphics[width=.4\textwidth]{figures/snap_reg.pdf} \\ 
+		\includegraphics[width=.4\textwidth]{figures/seismo.pdf}    && \includegraphics[width=.4\textwidth]{figures/seismo.pdf}
 	\end{tabular}
-\caption{\label{injection}Comparison between a global scale simulation and a regional scale simulation of wave propagation following an earthquake in south America.
+
+	\caption{\baselineskip 18pt 
+		Comparison between a global scale simulation and a regional scale simulation of wave propagation following an earthquake in south America.
 The seismograms corresponds to the recordings at 37 station located in north America (red spheres).
@@
 In (a), the wavefield is modeled globally using the spectral element method Specfem3D\_globe \citep{komatitsch2002spectrala}.
-In (b), the wavefield is modeled regionally using the RegSEM
+In (b), the wavefield is modeled regionally using the RegSEM.
 During the inversion, all the synthetic seismograms are computed using the regional SEM code RegSEM \citep{cupillard2012regsem} code and regenerated thanks to virtual sources located around the computational domain \citep[see][]{masson2013numerical}.
 The sismograms in (b) are exactly similar to those in (a), however, the computational effort is greatly reduced in the regional simulation (b).}
+
+	\label{injection}
+
 \end{figure}
@@
 
-% \subsection{Tectonic back ground}
-
-% Cratons are large areas of continental lithosphere that have remained coherent and relatively rigid since Precambrian \citep{hoffman1988united}.
-Through more than $3Ga$, cratons have survived compared to the $<200Ma$ life span of ocean basins.
-% In accordance with the absence of geoid anomalies over continents, Jordan proposed the concept of Continental Tectosphere \citep{jordan1975continental} where continental mantle is made of low-density, melt-depleted peridotite residua, whose low densities exactly offset the effect of thermal contraction associated with a growing thermal boundary layer \citep{lee2011building}.
-% The fact that the continental lithosphere manages to avoid the fate of recycling suffered by oceanic lithosphere has been attributed to high viscosity \citep{sleep2003survival, doin1997mantle}, neutral to positive buoyancy, low heat flow compared to surronding area \citep{jaupart2014building}, dehydrated \citep[i.e dryer than oceanic mantle until 250km deep according to electrical conductivity tomography,][]{yang2015three, hirth2000comparison}, and chemically depleted until ca.
-175km deep \citep{kopylova1999petrology} while still fertile below until ca.
-300km \citep{s2005archean}.
-
-% Since \cite{grand1984upper}, one well established feature within NA lithosphere is the contrast between the high velocity craton and the low velocity western active part.
-This feature is observed in many global models (e.g., \citep{megnin2000three, shapiro2002monte, panning2006three, houser2008shear, nettles2008radially, kustowski2008anisotropic, ritsema2010s40rts, simmons2010gypsum, lekic2011inferring, debayle2012global, schaeffer2013global, french2014whole,bozdag2016global} where the cratonic lithosphere is caracterized by higher than average (ca.
-$4.8 km.s^{-1}$) velocities down to at least $200km$ deep.
-
-% % Such values, at upper mantle depths with Standart Temperature Condition, are very important when interpreting shear wave velocities.
-For example, a relatively small velocity increase of $1 \%$ in the uppermost mantle can be attributed to a $5$ unit increase in the Mg\# relative to Fe, or an $220°C$ decrease in temperature.
-\citep{lee2003compositional, bedle2009svelocity}
-
-% % The combination of low temperature and depleted lithosphere induces observed high velocity anomalies.
-
-% Continental scale studies showed that, fastest anomalies are beneath archean and proterozoic terranes around 150km deep \citep{bedle2009svelocity,lee2005surface,schaeffer2014imaging,marone2007three,yuan20113}, there are correlations between surface tectonics provinces and shear waves velocities heterogeneities \citep{nettles2008radially,yuan2014lithospheric,darbyshire2013seismic}, upper mantle cratonic keels extends down to ca.
-250km with lateral thin out downward.
-\cite{yuan2010lithospheric} infered from azimuthal anisotropy study, that cratonic keels are vertically stratified and comprise a cold core of highly depleted material bounded deeper by a fertile (less depleted) thermal layer.
-While the sharp Mid-Lithospheric Discontinuity (MLD) between the chemical layer and thermal layer is observed (e.g \cite{thybo1997seismic,abt2010north,calo2016layered}), the Lithosphere-Asthenosphere Boundary (LAB) is more diffuse and is identified as flat in topography through a negative gradient in shear wave velocities associated with an alignement of anisotropic fast axes direction with Absolute Plate Motion.
-\citep{yuan2010lithospheric,eaton2009elusive}\\
-
-% At local scale, it has been observed that the stagnant shield suffered deformations from the convective mantle \citep{kaban2015cratonic,eaton2007seismic,frederiksen2013modification,boyce2016subduction}
-
-% !!! Still a lot of local studies to include,  Greenland, deep slab imaged, Wyoming, talk more about Divot, Hotspots, Appalachian, Gulf of Mexico, JdF. !!!
-% % This structure expands over both Archean and Proterozoic terranes in the canadian shield, which is not necessarly the case for all archean terranes within the plate (e.g the Wyoming block in the US \citep{yuan2014lithospheric}).
-
-% Such high shear wave velocity values, at upper mantle depths with Standart Temperature Condition, are very important when interpreting shear wave velocities.
-For example, a relatively small velocity increase of $1 \%$ in the uppermost mantle can be attributed to a $5$ unit increase in the Mg\# relative to Fe, or an $220°C$ decrease in temperature.
-\citep{lee2003compositional, bedle2009svelocity}
-
-% Mechanism for craton formation is debated (e.g., \cite{lee2011building} \cite{s2005archean}), and caracterizing cratons through high resolution imaging gives clues to elucidate this puzzle.
-% We propose interpretations on the geologic and tectonic evolution of the continent since both parameter variabilities are related to compositional, mineralogical and thermal states of the mantle.
-
-% The North American (NA) continent is fully part of the North American plate.
-This tectonic plate is moving at ca.
-$2cm$ per year in a west-south westerly direction around a rotation pole located in north western Ecuador \citep{bleeker2005north} Along its western margin, we can find:
-%  \begin{itemize}
-%  	\item Subduction of the Pacific plate under NA plate south to Alaska, the Cascadian subduction zone of the Juan de Fuca plate and the Coco plate subducting under central America, both generating arc volcanism and being part of the ancient Farallon plate which was subducting under NA plate since Jurassic period.
-%  	\item Strike-slip along the Queen Charlotte-Fairweather system in Canada and along the San Andreas fault in California.
-%  \end{itemize}
-
-%  The evolution of western North America has been controlled by this Mesozoic-Cenozoic ocean-continent collisional system. 
-%  Subduction events of oceanic lithosphere are thought to be responsible for the Laramide orogeny (Late Cretaceous), characterized by minor horizontal shortening in the crust, but considerable crustal thickening generating the rocky mountains (Bird 1988).
-%  The Rocky-Mountain Front (RMF) almost perfectly coincides with the boundary between this tectonically active area and the more stable cratonic part.
-
-%  Two third of the Laurentia craton consist of distinct archean provinces: 
-% \begin{itemize}
-% 	\item undeformed Slave, Superior and Wyoming cratons
-% 	\item highly reworked Rae and Hearne cratons (Churchill structural Province)
-% 	\item Nain and North Atlantic Craton (NAC) underlying southern Greenland
-% 	\item Sask which underlies the Trans-Hudson orogen Proterozoic belts (THO)
-% \end{itemize}
-%   All of these archean blocks, comprising most of the Canadian shield, were mainly welded by Proterozoic belts like the THO, but it is still under debates whether archean blocks originated from related ancestral landmasses (i.e a unique super craton, a couple of super cratons, or many super cratons).
-Greenland, despite its continental part was separated from NA continent ca.
-$80Ma$ ago, is fully part of the NA plates.
-While the NAC on its southern part has being studied over the last decades, central (Apparently a continuity of the THO \cite{bleeker2005north} and northern (apparently Rae-like archaean blocks) part are still poorly documented.
-
-%  In the periphery of this NA archean core, proterozoic belts like Wopmay, Penokean and Makkovik-Ketilian orogens from 2.0 to 1.8 Ga and Labradorian, Central Plains and Yavapai-Mazatzal orogens from 1.8 to 1.6 Ga represent zones of accretion for Proterozoic material and mark the amalgamation of the archean cores.
-\citep{hoffman1988united}
-%  The Grenvile orogen (ca.
-$1.1 Ga$) at the southeast of the craton core completed to shape the Laurentia craton as we can observe it today, surviving both Rodinia ($825-550Ma$) and Pangaea ($335-175Ma$) break-ups.
-
 \section{Results}
 
+\begin{figure}[p]
+	\centering
+	\includegraphics[width=1\textwidth]{figures/Vs-60-300-for-paper.png}
+
+	\caption{\baselineskip 18pt
+	3D isotropic shear wave velocity structure of the continent.
+Map views are shown from 60 km down to 300 km, shown as variations with respect to the global mean (thick black line in Figure ~\ref{1daverage}).
+	% Labels are: BP, Blake Plateau; NQ, New Quebec Orogen; THO, Trans-Hudson Orogen. State abbreviations are: CO, Colorado; MT, Montana; ND, North Dakota; SD, South Dakota; and WY, Wyoming. 
+	% The geological limits of the cratonic crust are indicated as follows: (to be defined when using final figures... 
+	}
+
+	\label{3d-VS}
+
+\end{figure}
+
+The tomographic inversion results for $V_S$ and $\xi$ are presented in figures (to be defined).
+We divide the model description and interpretation into two parts, focusing first on the continent-wide structure and then, in more detail, on eastern different regions.
+
 \subsection{Continental model}
-To link with the 1D profile!
-As observed in many other studies, average profiles of $V_S$ within North America (NA) in the shallow upper mantle is higher compared to Global average profiles biased by ocean structures. 
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=1\textwidth]{figures/1D_profile_withmean.png}
+
+	\caption{1D profiles of the global mean and the continental mean}
+	\label{1daverage}
+\end{figure}
+
+As observed in many other studies, average profiles of $V_S$ within North America (NA) in the shallow upper mantle is higher compared to Global average profiles biased by ocean structures as shown in Figure ~\ref{1daverage}.
 The difference is clearly distinct until ca. $270km$ and then fade away deeper, though, NA remains slightly higher than the global average. 
@@
 \subsection{Beneath the craton}
-\begin{figure}[h]
-\centering
+
+\begin{figure}
+	\begin{minipage}{0.5\linewidth}
+		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_A.pdf}}
+	\end{minipage}
+	\hfill
+	\begin{minipage}{0.5\linewidth}
+		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_B.pdf}}
+	\end{minipage}
+
+	\caption{Cross Section of the lithospheric core}
 	\label{cratoncross}
-% \begin{tabular}{ccc}
-\includegraphics[width=.45\textwidth]{/Users/Pierre/Documents/MATLAB/figures/iter2/profiles_NASEM_vs_dxi_A.pdf}  &&
-% \end{tabular}
-\caption{\bm{Cross Section of the lithospheric core}}
+
 \end{figure}
 
-To link with slices and a set of profiles
 % Gung 2003 Positive anomalies of dlnxi down to 400km in the craton.
@@
 While the beginning of the negative gradient shows a topography, the bottom of the gradient located beneath Slave, Hearne/Rae and western Superior cratons; appears very flat.
-As observed in Gung et al., 2003; down to 250km negative anomalies of $dln(\xi)$ are observed in the whole craton. 
+As observed in \cite{gung2003global}; down to 250km negative anomalies of $dln(\xi)$ are observed in the whole craton in contrast with the oceanic surrounding. 
 Deeper, positive anomalies are emerging beneath archaean Slave/Rae blocks and proterozoic Yavapai/Mazatzal/Great Rhyolites provinces, however $dln\xi$ anomalies remain positives beaneath the archaean Superior block.
@@
 
-\subsubsection{Wyoming}
-The Archaean aged Wyoming craton does not show as high anomalies as within the craton nucleus. 
+\subsubsection{Wyoming}{h}
+\begin{figure}
+	\begin{minipage}{0.5\linewidth}
+		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_wyoming_A.pdf}}
+	\end{minipage}
+	\hfill
+	\begin{minipage}{0.5\linewidth}
+		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_wyoming_B.pdf}}
+	\end{minipage}
+
+	\caption{Cross Section of the Wyoming lithosphere}
+	\label{wyomingcross}
+
+\end{figure}
+
+The Archaean aged Wyoming craton does not show as high anomalies as within the craton nucleus as shown in figure ~\ref{wyomingcross}. 
 It is split into 2 distinctive part. A western part showing low velocities influenced by the active tectonic with a radially positive gradient. 
@@
 \subsubsection{Slave}
-The Archaean aged Slave craton shows typical lithospheric craton signatures with higher than average velocities. 
+
+	\begin{figure}
+		\begin{minipage}{0.5\linewidth}
+			\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_slave_A.pdf}}
+		\end{minipage}
+		\hfill
+		\begin{minipage}{0.5\linewidth}
+			\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_slave_B.pdf}}
+		\end{minipage}
+
+		\caption{Cross Section of the Slave lithosphere}
+		\label{slavecross}
+
+	\end{figure}
+
+	The Archaean aged Slave craton shows typical lithospheric craton signatures with higher than average velocities as shown in figure ~\ref{slavecross}. 
 	It shows in the entire region a positive velocity gradient with a maximum of ca. $4.8km.s^{-1}$ associated with a maximum of $\xi$ at ca. $150km$ deep.
@@
 	\paragraph{Western part}
-The Archaean aged Superior craton shows typical lithospheric craton signatures with higher than average velocities. 
+		\begin{figure}
+			\begin{minipage}{0.5\linewidth}
+				\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_west_superior_A.pdf}}
+			\end{minipage}
+			\hfill
+			\begin{minipage}{0.5\linewidth}
+				\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_west_superior_B.pdf}}
+			\end{minipage}
+
+			\caption{Cross Section of the western Superior lithosphere}
+			\label{westsupcross}
+
+		\end{figure}
+
+		The Archaean aged Superior craton shows typical lithospheric craton signatures with higher than average velocities as shown in figure ~\ref{westsupcross}. 
 		It shows in the entire region a positive velocity gradient with a maximum of ca. $4.8km.s^{-1}$  associated with a maximum of $\xi$ at ca. $150km$ deep.
+
 		Deeper a negative gradient is observed with a minimum of ca. $4.6km.s^{-1}$ at ca. $250km$ deep. 
@@
 		\paragraph{Eastern part}
-Compared to its western part, the structure here shows differences. 
+		\begin{figure}
+			\begin{minipage}{0.5\linewidth}
+				\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_east_superior_A.pdf}}
+			\end{minipage}
+			\hfill
+			\begin{minipage}{0.5\linewidth}
+				\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_east_superior_B.pdf}}
+			\end{minipage}
+
+			\caption{Cross Section of the western Superior lithosphere}
+			\label{eastsupcross}
+
+		\end{figure}
+
+		Compared to its western part, the structure here shows differences as shown in figure ~\ref{eastsupcross}. . 
 		Fast velocities are present down to $150km$deep but are slower. 
@@
 \subsubsection{Yavapai Mazatsal region}
+	\begin{figure}
+		\begin{minipage}{0.5\linewidth}
+			\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_yava-maza_A.pdf}}
+		\end{minipage}
+		\hfill
+		\begin{minipage}{0.5\linewidth}
+			\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_yava-maza_B.pdf}}
+		\end{minipage}
+
+		\caption{Cross Section of the Yavapai and Mazatzal lithosphere}
+		\label{yavacross}
+
+	\end{figure}
+
 	Yavapai-Mazatzal orogens from 1.8 to 1.6 Ga represent zones of accretion for Proterozoic material and mark the amalgamation of the archean cores. \citep{hoffman1988united} \\
+	Cross section of the region are represented in figure ~\ref{yavacross}.
 	West to the RMF, the region shows low velocities influenced by the active tectonic with a radially positive gradient. 
@@
 \subsubsection{The Great Meteor track}
-Our model images a slow velocity anomaly in agreement with the reconstruction of the Great Meteor Track proposed by \cite{heaman2000timing}.
+	\begin{figure}
+		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_GMTrack.pdf}}
+
+		\caption{Cross Section of the lithosphere along the Great Meteor track based on \cite{heaman2000timing} construction}
+		\label{gmtcross}
+
+	\end{figure}
+
+	Our model images a slow velocity anomaly in agreement with the reconstruction of the Great Meteor Track (thin red line in figure ~\ref{3d-VS}) proposed by \cite{heaman2000timing}.
 	Below $100km$ the pan african lithosphere is cut in two by a slow velocity anomaly beneath the New England seamounts (youngest part of the GMT).
@@
 	We observe at $250km$ that the slow anomaly go further inland and crosses the Superior Craton northward until the Hudson bay following the GMT.
+	See the cross section along the re-built GMT in figure ~\ref{gmtcross}.
 

### 2017-02-17 Paper_NA.tex
Paper with cleaned figures

@@
 The tomographic inversion results for $V_S$ and $\xi$ are presented in figures (to be defined).
-We divide the model description and interpretation into two parts, focusing first on the continent-wide structure and then, in more detail, on eastern different regions.
+We divide the model description and interpretation into two parts, focusing first on the continent-wide structure and then, in more detail, on different regions.
 
@@
 
-\begin{figure}
+\begin{figure}[h]
 	\begin{minipage}{0.5\linewidth}
@@
 While the beginning of the negative gradient shows a topography, the bottom of the gradient located beneath Slave, Hearne/Rae and western Superior cratons; appears very flat.
-As observed in \cite{gung2003global}; down to 250km negative anomalies of $dln(\xi)$ are observed in the whole craton in contrast with the oceanic surrounding. 
-Deeper, positive anomalies are emerging beneath archaean Slave/Rae blocks and proterozoic Yavapai/Mazatzal/Great Rhyolites provinces, however $dln\xi$ anomalies remain positives beaneath the archaean Superior block.
+
+As observed in \cite{gung2003global}; down to 250km low positive ($< 2\%$) anomalies of $dln(\xi)$ (Compared to an isotropic $\xi$) are observed in the whole craton.
+In contrast with the oceanic surrounding that shows strong positive anomalies ($> 4\%$).
+The gulf of Mexico and the southern part of Greenland show $\xi$ anomalies that are comparable to oceanic signature. 
+The Grenville and Appalachian shelf (including the eastern Superior craton) show also higher $\xi$ anomalies comparable to oceanic signature. 
+Deeper, positive anomalies are emerging beneath archaean Slave/Rae blocks/Eastern Superior and proterozoic North THO/Yavapai/Mazatzal/Great Rhyolites provinces, however $dln\xi$ anomalies remain low beaneath the archaean eastern Superior block.
 In the whole lithosphere, $\xi$ is higher than $1$ which means that $V_{SH} > V_{SV}$.
 
-\subsubsection{Wyoming}{h}
-\begin{figure}
+\subsubsection{Wyoming}
+\begin{figure}[h]
 	\begin{minipage}{0.5\linewidth}
@@
 
-	\begin{figure}
+	\begin{figure}[h]
 		\begin{minipage}{0.5\linewidth}
@@
 	\paragraph{Western part}
-		\begin{figure}
+		\begin{figure}[h]
 			\begin{minipage}{0.5\linewidth}
@@
 		\paragraph{Eastern part}
-		\begin{figure}
+		\begin{figure}[h]
 			\begin{minipage}{0.5\linewidth}
@@
 \subsubsection{Yavapai Mazatsal region}
-	\begin{figure}
+	\begin{figure}[h]
 		\begin{minipage}{0.5\linewidth}
@@
 \subsubsection{The Great Meteor track}
-	\begin{figure}
+	\begin{figure}[h]
 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_GMTrack.pdf}}
@@
 
-\subsubsection{Gulf of Mexico}
-Fast anomalies are observed beneath the whole gulf of Mexico down to $150km$ deep.
+% \subsubsection{Gulf of Mexico}
+% Fast anomalies are observed beneath the whole gulf of Mexico down to $150km$ deep.
 Despite the proximity with pan-african lithosphere, such high velocity anomalies represent older than $150Ma$ oceanic lithosphere.

### 2017-02-21 Paper_NA.tex
with geological map

@@
 The Grenville and Appalachian shelf (including the eastern Superior craton) show also higher $\xi$ anomalies comparable to oceanic signature. 
-Deeper, positive anomalies are emerging beneath archaean Slave/Rae blocks/Eastern Superior and proterozoic North THO/Yavapai/Mazatzal/Great Rhyolites provinces, however $dln\xi$ anomalies remain low beaneath the archaean eastern Superior block.
+Deeper, positive anomalies are emerging beneath archaean Slave/Rae blocks/Eastern Superior and proterozoic North THO/Yavapai/Mazatzal/Great Rhyolites provinces, following highest velocities anomailies.
+However $dln\xi$ anomalies remain low beaneath the archaean eastern Superior block.
 In the whole lithosphere, $\xi$ is higher than $1$ which means that $V_{SH} > V_{SV}$.

### 2017-02-24 Paper_NA_br.tex
Paper with B.R. corrections

--- 2017-02-17 Paper_NA.tex
@@
 
-The North American continent consists of an ancient cratonic core, formed during the Archean that has not been subjected to orogeny for at least 1.8 Ga, bordered to the southeast, by progressively younger, stable, Proterozoic provinces \citep{hoffman1988united}.
-In the west, the Rocky Mountain Front (RMF) separates the Proterozoic and Archean shield from younger, tectonically active provinces (Fig.
-1).
+\begin{figure}[p]
+	\centering
+	\centerline{\includegraphics[viewport=180 340 480 570,clip=true,width=\linewidth]{figures/Geologicalmap_Clouzet.pdf}}
+
+	% bb=llx lly urx ury :
+
+	\caption{\baselineskip 18pt
+	Large scale Precambrian architecture of North America (modified from \cite{whitmeyer2007tectonic}.
+Colours indicate ages of different regions, as shown in the legend.
+The stable cratonic core is defined by the Rocky Mountain Front to the west (dashed dark grey line), defined by the surface trace of deformation), and the appalachian fron to the east (solid light red line).
+The solid blue line demarcates the westward extent of the Grenville/Llano Front.
+The solid red line along the Cordillera province demarcates the breakup of Rodinia.
+Solid green lines indicate tectonic plate boundaries.
+Abreviations are as follows: RW, re-worked; GBA, Great Bear Arc; BHT, Buffalo Head Terrane; Taltson, Talston Arc; Rimbey, Rimbey Arc; Tornagat, Torngat Arc; L.B.A, Little Belt Arc; LRA; THO, Trans-Hudson Orogen; RGR, Rio Grande Rift; Reelfoot, Reelfoor Rift; Kwn.R, Kewanowan Rift; Sask, Sask Craton; WY, Wyoming Province; Med.
+Hat., Medicine Hat Block; Grt.
+Rhy, Granite–Rhyolite Province; Grenville, Grenville Province.
+	}
+
+	\label{mapgeol}
+
+\end{figure}
+
+The North American continent consists of an ancient cratonic core, formed during the Archean, that has not been subjected to orogeny for at least 1.8 Ga, bordered to the southeast and east, by progressively younger, stable, Proterozoic provinces \citep{hoffman1988united}.
+In the west, the Rocky Mountain Front (RMF) separates the Proterozoic and Archean shield from younger, tectonically active provinces (Fig.~\ref{mapgeol}).
 This relatively regular configuration makes North America an ideal target for seismic tomography, to investigate the relationship of lithospheric and asthenospheric structure to the geological features observed at the surface.
@@
 Differences in the seismic velocity structure between the craton and the western US extending down to at least 200 km were established already in the 1960's and 70's from the first studies of teleseismic P and S wave travel time anomalies  \citep[e.g.][]{cleary1966analysis,herrin1968regional,poupinet1979relation}, the first tomographic images based on P and S travel times \citep{romanowicz1979seismic} and long period surface waveforms \citep{woodhouse1984mapping}, as well as contrasted velocity-depth profiles obtained from the modeling of shear body waveforms at relatively short distances \citep[e.g.][]{grand1984upper}.
-
-More recent shear wave tomography, at scales from global \citep[e.g.][]{su1994degree,megnin2000three,shapiro2002monte,panning2006three,kustowski2008anisotropic,ritsema2010s40rts,lekic2011inferring,debayle2012global,schaeffer2013global} to continental \citep{lee2005surface,yuan20113,yuan2014lithospheric,schaeffer2014imaging}, confirm the presence of strong lateral variations in the thickness of the lithosphere across the RMF, with lithospheric roots extending down to 200-250 km under the craton, thinning abruptly to the west to less than 80-100 km.
-This is also found from azimuthal anisotropy tomography, where it is manifested by a change in the anisotropy fast axis direction across the lithosphere-asthenosphere boundary (LAB) \citep{marone2007depth,yuan2010lithospheric}.
-
-While the mechanism of craton formation is still widely debated \citep[e.g.][]{s2005archean,lee2011building}, the presence of finer scale lateral and depth variations in seismic structure suggest a complex history.
-Studies of azimuthal anisotropy have shown the presence of laterally varying layering within the cratonic lithosphere \citep[e.g.][]{levin1999shear,deschamps2008stratified,yuan2010lithospheric}, which may indicate different modes and/or times of formation of the top c.a.
+More recent body wave tomographic models have added considerable details to these images \citep{sigloch2011mantle,burdick2014model,porritt2014seismic,porritt2015lithospheric}
+
+Lateral variations in absolute velocities can be inferred from shear wave tomography, at scales from global \citep[e.g.][]{su1994degree,megnin2000three,shapiro2002monte,panning2006three,kustowski2008anisotropic,ritsema2010s40rts,lekic2011inferring,debayle2012global,schaeffer2013global} to continental \citep{lee2005surface,yuan20113,yuan2014lithospheric,schaeffer2014imaging}.
+These studies confirm the presence of strong lateral variations in the thickness of the lithosphere across the RMF, with lithospheric roots extending down to 200-250 km under the craton, thinning abruptly to the west to less than 80-100 km.
+This is also found from azimuthal anisotropy tomography, where  a change in the anisotropy fast axis direction across the lithosphere-asthenosphere boundary (LAB) \citep{marone2007depth,yuan2010lithospheric}.
+
+While the mechanism of craton formation is still widely debated \citep[e.g.][]{s2005archean,lee2011building}, the presence of finer scale lateral and depth variations {\color{red} (cite some body wave tomography studies} in seismic structure suggest a complex history.
+Moreover, studies of azimuthal anisotropy have shown the presence of laterally varying layering within the cratonic lithosphere \citep[e.g.][]{levin1999shear,deschamps2008stratified,yuan2010lithospheric}, which may indicate different modes and/or times of formation of the top c.a.
 100 km of this lithosphere.
-This is also suggested by fine scale studies of converted and reflected phases, which indicate the presence of a sharp mid-lithospheric discontinuity (MLD) within the craton, marking the top of a mid-lithospheric low velocity zone \citep{thybo1997seismic,bostock1998mantle,abt2010north,fischer2010lithosphere,rader2015characterization,ford2016midlithospheric} and even possibly several mid-lithospheric discontinuities \citep{calo2016layered}.
-In addition, there is evidence from shear wave radially anisotropic tomography for separation between blocks of different crustal ages extending to depths of at least 150 km \citep[e.g.][]{yuan2014lithospheric}.
-
-Much of this finer scale structure has been highlighted owing to the availability of data from the dense broadband USArray TA deployment.
-With the completion of the coverage of the conterminous US, as well as availability of data from Canada and Greenland, it is now possible to further refine shear wave tomographic images of North America, and in particular of the stable Proterozoic and Archean provinces, to try and improve our understanding of its formation.
+Fine scale studies of converted and reflected phases indicate the presence of a sharp mid-lithospheric discontinuity (MLD) within the craton, marking the top of a mid-lithospheric low velocity zone \citep{thybo1997seismic,bostock1998mantle,abt2010north,fischer2010lithosphere,rader2015characterization,ford2016midlithospheric} and even possibly several mid-lithospheric discontinuities \citep{calo2016layered}.
+In addition, there is evidence from radially anisotropic shear wave tomography for separation between blocks of different crustal ages extending to depths of at least 150 km \citep[e.g.][]{yuan2014lithospheric}.
+
+Much of this finer scale structure has been resolved owing to the availability of data from the dense broadband USArray TA deployment.
+With the completion of the coverage of the conterminous US, as well as availability of data from Canada and Greenland, it is now possible to further refine shear wave tomographic images of North America, and in particular of the stable Proterozoic and Archean provinces, to try and improve our understanding of the different stages of formation of the continent.
 It is also an unprecedented opportunity to experiment with improved waveform-based tomographic techniques.
 
-In a previous study, we presented a radially anisotropic shear velocity model of the North American upper mantle based on a combination of long period teleseismic and regional waveform data \citep{yuan2014lithospheric} The regional waveform data (down to 40 s period) were, for the first time at this scale, compared to 3D synthetics computed using RegSEM \citep{cupillard2012regsem}, a Spectral Element Method (SEM) code appropriate for continental scale wavefield computations.
-Due to the extremely lengthy computations that would have been necessary to compute it numerically at each iteration of the inversion, the predicted teleseismic wavefield was, instead, computed using Non-Linear Assymptotic Coupling Theory \citep[NACT,][]{li1995comparison}, a methodology based on normal mode perturbation theory, that is approximate, but computationally more efficient, and has been used in the development of several generations of global and continental scale shear velocity models \citep{li1996global,megnin2000three,gung2003global,panning2006three,yuan20113}
+In a previous study, we presented a radially anisotropic shear velocity model of the North American upper mantle based on a combination of long period teleseismic and regional waveform data \citep{yuan2014lithospheric}.
+The regional waveform data (down to 40 s period) were, for the first time at this scale, compared to 3D synthetics computed using RegSEM \citep{cupillard2012regsem}, a Spectral Element Method (SEM) code suitable for continental-scale wavefield computations.
+Due to the very heavy computations that would have been necessary to compute the predicted teleseismic wavefield numerically at each iteration of the inversion, the latter was, instead, computed using Non-Linear Assymptotic Coupling Theory \citep[NACT,][]{li1995comparison}, a methodology based on normal mode perturbation theory, that is computationally more efficient (albeit approximate), and has been used in the development of several generations of global and continental scale shear velocity models based on time-domain waveform inversion \citep{li1996global,megnin2000three,gung2003global,panning2006three,yuan20113}
 
 The resulting 2014 model of North America presents some interesting features, in particular a correlation of radial anisotropy structure with lithospheric blocks corresponding to different orogenies in the eastern US and continental shelf.
-While more rigorous than inversions practiced by most groups and based on the path-average surface wave approximation, this mixed methodology nevertheless presents some inconsistencies from the theoretical point of view, since the predicted wavefield through the target model space is computed with different theories for teleseismic and regional distance data.
-
- To better integrate teleseismic data into our regional tomographic inversions, we developed a general framework called "Box Tomography"  (see \cite{masson2013numerical}, Masson and Romanowicz, 2016a,b) that allow us to consistently model and invert both teleseismic (i.e.
+While more rigorous than inversions practiced by most groups and based on the path-average surface wave approximation \citep[PAVA,][]{woodhouse1984mapping}, this mixed methodology nevertheless presents some inconsistencies from the theoretical point of view, since the predicted wavefield through the target model space is computed with different theories for teleseismic and regional distance data, respectively.
+
+ To better integrate teleseismic data into our regional tomographic inversions, we developed a general framework called "Box Tomography"  \citep[see][]{masson2013numerical,masson2016box,masson2017fast}, that allow us to consistently model and invert both teleseismic (i.e.
 associated with sources or receivers outside of the imaged region) and regional (i.e.
 associated with sources and receivers within the imaged region) waveform data using accurate numerical methods such as SEM.
-In Box tomography, prior to the inversion, the seismic wavefield generated by teleseismic sources is first modeled numerically at the global scale for a given reference model and recorded at the surface of the region to be imaged.
+ In Box Tomography, prior to the inversion, the seismic wavefield generated by teleseismic sources is first modeled numerically at the global scale for a given reference 3D model and is recorded at the surface of the region to be imaged.
  This reference wavefield is then used to construct virtual sources lying at the boundary of the regional modeling domain and reproducing the original wavefield as illustrated in Figure~\ref{injection}.
-Once the telesismic sources have been moved (i.e.
-replaced by virtual sources) within the regional modeling domain, the tomographic inversion can be performed efficiently in a classic manner (i.e.
+ Once the teleseismic sources have been moved (i.e.
+replaced by virtual sources) within the regional modeling domain, the tomographic inversion can be performed efficiently in a classical manner (i.e.
 using regional modeling only) as the teleseismic data are accounted for seamlessly thanks to the virtual sources.
-Related concepts have been proposed to account for teleseismic data into regional imaging (e.g.
-Wang et al., 2016 Monteiller et al., 2015), however, in these studies the global modeling of the reference wavefield is performed using a faster method that do not account for the effects induced by the 3D structure of the Earth outside the regional modeling domain.
-In this paper, we present the first application of Box tomography, where the 3D structure is accounted for outside the imaging box, to continental scale waveform tomography in North America.
+ Related concepts have been proposed to account for teleseismic data in regional imaging (e.g.
+Wang et al., 2016 Monteiller et al., 2015), however, in these studies the global modeling of the reference wavefield is performed using a faster method that does not account for the effects induced by the 3D structure of the Earth outside the regional modeling domain.
+In this paper, we present the first application of Box Tomography to continental scale waveform tomography in North America.
 
@@
 When inverting waveforms, and in particular when a more accurate theory than the PAVA is considered, such a simple procedure is not possible, and it is necessary to recompute the synthetic teleseismic waveforms, at each iteration, in a 3D model which is fixed outside of the target region, and updated only within it.
-This leads to substantial computations, even in the case of an approximate theory such as NACT, and becomes prohibitively expensive, if the 3D teleseismic wavefield is computed using SEM or another accurate numerical method, as one aims to include increasingly shorter periods.
-
-To overcome this issue, we take advantage of the "Box Tomography" theory which allows to combine teleseismic and regional waveforms in a consistent manner to image regional targets at arbitrary locations.
-This general framework accomodates for arbitrary acquisition setups (i.e.
-with sources and receivers located outside the regional imaging box), is compatible with most popular numerical methods such as SEM or FD, and can produce exact seismograms and senitivity kernels (i.e.
+This leads to substantial computations, even in the case of an approximate theory such as NACT, and becomes prohibitively expensive, if the 3D teleseismic wavefield is computed using SEM or another accurate numerical method, as especially as one aims to include increasingly shorter periods.
+
+To overcome this problem, we take advantage of the "Box Tomography" theory which allows us to combine teleseismic and regional waveforms in a consistent manner to image regional targets at arbitrary locations.
+This general framework accommodates for arbitrary acquisition setups (i.e.
+with sources and receivers located outside the regional imaging box), is compatible with most popular numerical methods such as SEM or FD, and can produce exact seismograms and sensitivity kernels (i.e.
 similar to those obtained when solving the problem globally).
 In this study, we limit ourselves to the situation where both regional and teleseismic events are employed but where all the seismic stations lie within the regional computational domain. 
-Furthermore, we neglect some higher order scattering effects as proposed by Masson and Romanowicz (2016a).
-In this situation, Box Tomography is particularly efficient as the computational effort  to account for teleseismic data in regional inversions is limited to a few global scale simulations that are done once and for all prior to the inversion.
-Masson and Romanowicz (2016b) showed that this approach can produce accurate regional tomographic images even though the elastic structure outside the imaged region is neither fully known prior to the inversion nor updated during the inversion.
+Furthermore, we neglect some higher order scattering effects, as proposed by \cite{masson2016box}. 
+In this situation, Box Tomography is particularly efficient, as the computational effort  to account for teleseismic data in regional inversions is limited to a few global scale simulations that are done once and for all prior to the inversion.
+\cite{masson2016box} showed that this approach can produce accurate regional tomographic images even though the elastic structure outside the imaged region is neither fully known prior to the inversion, nor updated during the inversion.
 
@@
 This is the first time this methodology is applied in practice, and this study represents a proof of concept for this approach. 
-Adding teleseismic data allows better azimuthal coverage of the target region than can be achieved using only the regional dataset and eventually provides for the inclusion of constraints from additional phases such as, for example, teleseismic SS phases reflected inside the target region.
+Adding teleseismic data allows better azimuthal coverage of the target region than can be achieved using only the regional dataset, and eventually provides for the inclusion of constraints from additional phases such as, for example, teleseismic SS phases reflected inside the target region.
 
@@
 
-The waveforms are filtered between 40 and 400 s, with corner frequencies of 53 and 250 s, and windowed into wavepackets, according to the procedure of \cite{li1996global} allowing different weights to be applied according to relative amplitude of individual wavepackets, redundancy of paths and signal to noise ratio.
+The waveforms are filtered between 40 and 400 s, with corner frequencies at 53 and 250 s, and windowed into wavepackets, according to the procedure of \cite{li1996global} ,allowing different weights to be applied according to relative amplitude of individual wavepackets, redundancy of paths and signal to noise ratio.
 This weighting step is crucial, as it homogeneizes the data coverage within the region and provides a way to construct the data covariance matrix $C_D$, which we approximate as a diagonal matrix. 
-At each iteration of the inversion, only those wavepackets are considered that satisfy predefined goodness of fit criteria compared to synthetics calculated in the current 3D model, which avoids cycle slipping in our time domain waveform inversion procedure.
+At each iteration of the inversion, only those wavepackets are considered that satisfy predefined goodness of fit criteria compared to synthetics calculated in the current 3D model, in particular to avoid cycle slipping in our time domain waveform inversion procedure.
 As the model improves, at each iteration, additional wavepackets are included. 
-The total number of wavepackets of fundamental and overtone Love and Rayleigh waves is given in Table 1, and provides a good azimuthal coverage within the NA continent (Figure ~\ref{density_coverage}).
+The total number of wavepackets of fundamental and overtone Love and Rayleigh waves is given in Table 1, and provides a good coverage within the NA continent (Figure ~\ref{density_coverage}).
 
 \begin{figure}[p]
+	% bb=llx lly urx ury :
 	\begin{minipage}{0.47\linewidth}
-		\centerline{\includegraphics[bb=120 10 470 380,clip=,width=\linewidth]{figures/Coverage_L_iter3.png}}
+		\centerline{\includegraphics[viewport=120 10 470 420,clip=,width=\linewidth]{figures/Coverage_L_iter3.png}}
 	\end{minipage}
@@
 	\begin{minipage}{0.47\linewidth}
-		\centerline{\includegraphics[bb=120 10 470 380,clip=,width=\linewidth]{figures/Coverage_T_iter3.png}}
+		\centerline{\includegraphics[viewport=120 10 470 420,clip=,width=\linewidth]{figures/Coverage_T_iter3.png}}
 	\end{minipage}
@@
 	\begin{minipage}{0.47\linewidth}
-		\centerline{\includegraphics[bb=120 10 470 380,clip=,width=\linewidth]{figures/Coverage_Z_iter3.png}}
+		\centerline{\includegraphics[viewport=120 10 470 420,clip=,width=\linewidth]{figures/Coverage_Z_iter3.png}}
 	\end{minipage}
@@
 	\begin{minipage}{0.47\linewidth}
-		\centerline{\includegraphics[bb=120 10 470 380,clip=,width=\linewidth]{figures/Coverage_all_iter3.png}}
+		\centerline{\includegraphics[viewport=120 10 470 420,clip=,width=\linewidth]{figures/Coverage_all_iter3.png}}
 	\end{minipage}
@@
 In addition to long period waveforms, we also consider a group velocity dispersion dataset \citep{shapiro2002monte} provided in the form of $1^o \times 1^o $ maps between 25 and 100 s. 
-The shorter periods (below 60 s) are used to constrain our homogeneized crustal structure at each iteration, as described below, while the entire period range is included during the inversion for mantle structure, providing additional constraints for structure in the shallow parts of the mantle that are consistent with the treatment of the crust.
+The shorter periods (below 60 s) are used to constrain our homogeneized crustal structure at each iteration, as described below, while the entire period range is included during the inversion for mantle structure, providing additional constraints for structure in the shallow parts of the mantle, that are consistent with the treatment of the crust.
 
@@
 Instead, we compute a radially anisotropic smooth crustal model on a $2^o \times 2^o $ grid through Monte Carlo Markov Chain (MCMC) simulation constrained by the group velocity dispersion data.
-This process is a sort of "homogeneization" (e.g.
-Backus, 1961; Capdeville and Marigo, 2008), which yields a crustal model equivalent to a real crust within the period range of the data used to constrain it (i.e.
-for periods longer than 20s).
-The thickness of the crust is a priori chosen as that of model Crust2.0 \citep{basscrust2} where it is larger or equal to 30 km, and it is fixed at 30 km otherwise. 
-This results in a smooth, but realistic crust within most of the continent, and a crustal model that is not interpretable in the border regions of our model space (i.e.
-in the oceans and borders of the continent).
-We have shown that this does not bias the mantle structure obtained below 50 km \citep{french2014whole}. 
-At each iteration of the inversion, we only invert for structure below 30 km depth, and subsequently recompute the crustal model for the next iteration by applying our MCMC approach with mantle structure fixed to be that of the current iteration model.
+
+This process is called "homogeneization" \citep[e.g.][]{backus1962long,capdeville2008shallow}, and yields a crustal model equivalent to a real crust within the period range of the data used to constrain it (i.e.
+here for periods longer than 20s).
+The thickness of the crust is a priori fixed to that of model Crust2.0 \citep{basscrust2} when it is larger or equal to 30 km, and it is fixed at 30 km in regions where it is thinner. 
+This results in a smooth, but realistic crust within most of the continent, and a crustal model that is not directly interpretable in the border regions of our model space (i.e.
+in the oceans and borders of the continent where the real crust is thin).
+We have shown that this does not bias the mantle structure imaged below 50 km \citep{french2014whole}. 
+At each iteration of the inversion, we only invert for structure below 30 km depth, and subsequently recompute the crustal model for the next iteration by applying our MCMC approach with mantle structure fixed to that of the current iteration model.
 The procedure is described and discussed in detail in \cite{french2014whole}
 
-Assuming that our azimuthal coverage is everywhere sufficient within our target region (Figure ~\ref{density_coverage}), we here solve only for radially anisotropic structure and specifically for two of the 5 parameters that describe such a structure, the two parameters to which long period surface waveforms are the most sensitive: isotropic velocity $V_{S}$ and the anisotropic parameter $\xi = {\frac{V_{SH}}{V_{SV}}}^2$, where $V_{SH}$ is the velocity of waves polarized horizontally and $V_{SV}$ that of waves polarized vertically .
-The other three parameters and density are constrained through empirical scaling relationships, following \cite{montagner1989petrological}, and are based on laboratory measurements for upper mantle rocks.
-If (A,C,F,L and N) are the 5 elastic parameters describing a radially anisotropic (or VTI) medium (Love, 1927), then:
+Assuming that our azimuthal coverage is everywhere sufficient within our target region, we here solve only for radially anisotropic structure.
+If (A,C,F,L and N) are the 5 elastic parameters describing a radially anisotropic (or VTI) medium \citep{love2013treatise}, then:
 
@@
  \end{equation}
+
+We only consider two of the 5 independent anisotropic parameters , those to which long period surface waveforms are the most sensitive: isotropic velocity $V_{S}$ and the anisotropic parameter $\xi = {\frac{V_{SH}}{V_{SV}}}^2$, where $V_{SH}$ is the velocity of waves polarized horizontally and $V_{SV}$ that of waves polarized vertically .
+The other three parameters and density are constrained through empirical scaling relationships, following \cite{montagner1989petrological}, and are based on laboratory measurements for upper mantle rocks.
 
@@
 
-We prefer parametrization in terms of $V_S$ and $\xi$, rather than $V_{SH}$ and $V_{SV}$, as this allows us to consider different spatial resolution and apply higher damping in the inversion to the less well resolved anisotropic parameter $\xi$, rather than having to reconstruct this parameter from differences in perturbations in the two quantities $V_{SH}$ and $V_{SV}$, which would limit the resolution allowed in isotropic velocity.
+We chose a parametrization in terms of $V_S$ and $\xi$, rather than $V_{SH}$ and $V_{SV}$, as this allows us to consider different spatial resolution and apply higher damping in the inversion to the less well resolved anisotropic parameter $\xi$, rather than having to reconstruct this parameter from differences in perturbations in the two quantities $V_{SH}$ and $V_{SV}$, which would limit the resolution allowed in isotropic velocity.
 
@@
 The target model space is geographically defined as shown in Figure ~\ref{map_evt_stn}, and is limited in depth down to 800 km. 
-As in our previous tomographic studies, the model space is parametrized in terms of 26 cubic splines $\nu(r)_{q}$ vertically \citep{megnin2000three} from the core-mantle boundary to the Moho, although we invert for structure only in the top 16 splines, corresponding to the top 700-800 km of the mantle.
+As in our previous tomographic studies, the model space is parametrized in terms of 26 cubic splines $\nu_{q}(r)$ vertically \citep{megnin2000three} from the core-mantle boundary to 30 km deep. 
+Although we invert for structure only in the top 16 splines, corresponding to the top 700-800 km of the mantle. 
 The spline nodes are spaced more closely at shallow depths, and located at the following radii: 5690, 5810, 5910, 5900, 6061, 6101, 6131, 6161, 6191, 6221, 6241, 6261, 6281, 6301, 6321, 6341 km.  
@@
 The spherical spline parametrization has the advantage of allowing for variable grid parametrization, which can be adjusted according to data coverage \citep[e.g.][]{marone2007three}. 
-In our case, outside of the target region, we adopt a "level 6" spherical grid for $V_{S}$ ($2^{o}$ knot spacing) and a "level 4" spherical grid for $\xi$ ($8^{o}$ knot spacing), consistent with the parametrization of SEMUCB\_wm1.
+In our case, outside of the target region, we adopt a "level 6" spherical grid for $V_{S}$ ($2^{o}$ knot spacing) and a "level 4" spherical grid for $\xi$ ($8^{o}$ knot spacing), consistent with the parametrization of $SEMUCB\_wm1$ \citep{french2014whole}.
 Inside the well-sampled target region, we define a level 7 spherical grid ($1^{o}$ knot spacing) for $V_{S}$ and level 6 ($2^{o}$ knot spacing) for $\xi$.
@@
 
-During the inversion, all the synthetic seismograms are computed using the regional SEM code RegSEM \citep{cupillard2012regsem} that takes into account effects of oceans, topography/bathymetry, ellipticity, and anelasticity, and where the computational domain is truncated using absorbing boundaries.
+During the inversion, all the synthetic seismograms are computed using the regional SEM code RegSEM \citep{cupillard2012regsem} that takes into account effects of oceans, topography/bathymetry, ellipticity, and anelasticity, and where the limits of the computational domain are dealt with using absorbing boundaries.
 The synthetic seismograms associated with regional data (i.e.
-where both the seismic stations and the earthquake are located within the regional modeling domain) do not require any specific treatment and are modeled as in our previous studies \citep[e.g.][]{yuan2014lithospheric}.
+where both the seismic stations and the earthquake are located within the regional modeling domain) do not require any specific treatment and are computed as in our previous studies using RegSEM \citep[e.g.][]{yuan2014lithospheric}.
 The synthetic seismograms associated with teleseismic data (i.e.
-where the earthquake and the seismic stations are located outside and inside the regional modeling domain, respectively) are obtained using a two step procedure as proposed by Masson and Romanowicz (2016).
-Prior to the inversion, the wavefields generated by teleseismic earthquakes are computed globally within our starting model ($SEMucb\_wm1$) and using an adapted version of SPECFEM3D\_globe \citep{komatitsch2002spectrala}.
+where the earthquake and the seismic stations are located outside and inside the regional modeling domain, respectively) are obtained using a two step procedure as proposed by \cite{masson2016box} as follows.
+
+Prior to the inversion, the wavefields generated by teleseismic earthquakes are computed globally within our starting model ($SEMucb\_wm1$) and a version of $SPECFEM3D\_globe$ \citep{komatitsch2002spectrala} adapted to this model representation.
 During these global simulations, the 3 component displacement wavefield is recorded at a set of points with locations prescribed by the RegSEM code.
-Within the regional solver RegSEM, these points are the coallocation or Gauss-Lobatto-Legendre (GLL) points belonging to the one element thick surface surrounding the regional modeling domain \citep[see][]{masson2013numerical}.
-Notice that our procedure do not require to store the stress nor the strain wavefield which are computed naturally by the regional solver, this makes it easy to swap between different codes for modeling global wave propagation (i.e.
+Within the regional solver RegSEM, these points are the collocation or Gauss-Lobatto-Legendre (GLL) points belonging to the one element thick surface surrounding the regional modeling domain \citep[see][]{masson2013numerical}.
+Note that our procedure does not require to store either the stress or the strain wavefield, which are computed naturally by the regional solver.
+This makes it easy to swap between different codes for modeling global seismic wave propagation (i.e.
 any code that outputs displacement seismograms can be used as such).
-Because the Courant criteria which ensure the stability in SEM often leads to oversampling, we compress the recordings versus time using a least square B-spline transform (Unser et al., 1993ab), we found this  approach more efficient and practical than the more classic decimation/interpolation scheme.
-We typically achieve data compression ratio lying between 10 and 100 with no significant loss of accuracy. 
-During the inversion, the global recordings are transformed to virtual sources that regenerate the global wavefield regionally, this operation is done on the fly by the RegSEM code and illustrated in Figure ~\ref{injection}b.
-
-Overall, the additional computational effort to account for teleseismic event consists of one global simulation per teleseismic event that are done once and for all before the inversion starts.
+Because the Courant criteria which ensure the stability in SEM often lead to oversampling, we compress the recordings versus time using a least square B-spline transform \citep{unser1993a,unser1993b}.
+We found this  approach more efficient and practical than the more classic decimation/interpolation scheme.
+We typically achieve a data compression ratio of between 10 and 100 with no significant loss of accuracy. 
+During the inversion, the global recordings are transformed to virtual sources that regenerate the global wavefield regionally. 
+This operation is done on the fly by the RegSEM code.
+
+Overall, the additional computational effort to account for teleseismic events consists of one global simulation per teleseismic event.
+These simulations are done once and for all before the inversion starts.
 
@@
 
-In continuity of our previous work \citep[e.g][]{marone2007three,yuan20113,yuan2014multiscale}, we use a hybrid iterative inversion scheme where, at each iteration, the forward wavefield is computed accurately using SEM, but the inverse step is solved using the formalism of \cite{tarantola1982generalized}  with sensitivity kernels calculated approximately using normal mode perturbation theory.
+In continuity of our previous work \citep[e.g][]{marone2007three,yuan20113,yuan2014lithospheric}, we use a hybrid iterative inversion scheme where, at each iteration, the forward wavefield is computed accurately using SEM, but the inverse step is solved using the formalism of \cite{tarantola1982generalized} with sensitivity kernels calculated approximately using normal mode perturbation theory.
 This allows us to apply a fast converging Gauss-Newton quadratic optimization scheme. 
-As shown in \cite{tarantola2005inverse}, and in Appendix A of \cite{lekic2011inferring}, it is far more important to use an accurate forward modelling scheme to compute the misfit function, while inaccuracies in the theoretical treatment of kernels result in smoothing errors that can be compensated for in subsequent iterations.
-Currently popular "adjoint tomographic" approaches \citep{zhu2015seismic,bozdag2016global} do compute the numerically exact gradient, however, they also need to apply smoothing operators and regularization, which degrades the accuracy of their kernels.
+As shown in \cite{tarantola2005inverse}, and in Appendix A of \cite{lekic2011inferring}, it is far more important to use an accurate forward modelling scheme to compute the misfit function, provided of course that the theory captures the effects of the background long wavelength structure accurately \citep[unlike the examples shown in][]{valentine2016impact}.
+Inaccuracies in the theoretical treatment of kernels then result in smoothing errors that can be compensated for in subsequent iterations.
+In contrast, currently popular "adjoint tomographic" approaches \citep{zhu2015seismic,bozdag2016global} compute the numerically exact gradient, however, they also need to apply smoothing operators and regularization, which degrades the accuracy of their kernels.
 Importantly, they rely on a linear optimization scheme (conjugate gradient method), which is characterized by very slow convergence.
 
-Our misfit function is defined in the time domain from the point by point differences between observed and synthetic waveforms by:
-
+Our misfit function is defined in the time domain from the point by point differences between observed and synthetic waveforms as follows:
 \begin{equation}
@@
 
-where $m_{k}$ represents the model estimate at the k-th iteration, $d$ is the data vector (waveform discretized in time or group velocity as a function of period) and $g(m_k)$ is the corresponding discretized wavefield computed using SEM, or the predicted group velocity dispersion.
-$m_p$ is the model prior (i.e.
+\noindent where $m_{k}$ represents the model estimate at the k-th iteration, $d$ is the data vector (waveform discretized in time or group velocity as a function of period) and $g(m_k)$ is the corresponding discretized wavefield computed using SEM, or the predicted group velocity dispersion.
+The model prior is $m_p$  (i.e.
 the 3D starting model) and $C_m$ and $C_d$ represent a priori model and data covariance matrices, respectively.
@@
 
-where $G$ is the matrix of Frechet derivatives of $g(m)$ calculated at $m_k$.
-We compute G using NACT or the Path Average Approximation (PAVA,\cite{woodhouse1984mapping}), depending on the distance range of the corresponding source-station path.
-NACT and PAVA are both asymptotic (high frequency) approximations to normal mode perturbation theory. 
-The PAVA includes along branch mode coupling only and is the standard surface wave approximation \citep[e.g][]{mochizuki1986free,romanowicz1987multiplet}, as used for example in \cite{woodhouse1984mapping}, in which a frequency shift and distance shift are introduced for each mode to account for the effects of heterogeneous structure.
+\noindent where $G$ is the matrix of Frechet derivatives of $g(m)$ calculated at $m_k$. We compute G using NACT or PAVA, depending on the distance range of the corresponding source-station path.
+NACT and PAVA are both asymptotic (high frequency) approximations to first order normal mode perturbation theory. 
+The PAVA includes along branch mode coupling only and is equivalent to the standard surface wave approximation \citep[e.g][]{mochizuki1986free,romanowicz1987multiplet}, as used for example in \cite{woodhouse1984mapping}, in which a frequency shift and distance shift are introduced for each mode to account for the effects of heterogeneous structure.
 The corresponding kernels are "1D", i.e.
@@
 These effects become important for accurate amplitude fitting.
-With our choice of misfit function, we are first and foremost fitting the phase, and for that, the 2D effects in the vertical plane are dominant, and important especially for overtones, as illustrated in \cite{romanowicz2008computation}.
+With our choice of misfit function, we are first and foremost fitting the phase, and for that, the 2D effects in the vertical plane are dominant, and important especially for overtones, as illustrated in \cite{megnin1999effects,romanowicz2008computation}.
 
@@
 
-\begin{figure}[p]
+\begin{figure}[ht]
 	\centering
@@
 	3D isotropic shear wave velocity structure of the continent.
-Map views are shown from 60 km down to 300 km, shown as variations with respect to the global mean (thick black line in Figure ~\ref{1daverage}).
+Map views are shown from 60 km down to 300 km, shown as variations with respect to the global mean (thick red line in Figure ~\ref{1daverage}).
 	% Labels are: BP, Blake Plateau; NQ, New Quebec Orogen; THO, Trans-Hudson Orogen. State abbreviations are: CO, Colorado; MT, Montana; ND, North Dakota; SD, South Dakota; and WY, Wyoming. 
@@
 
-The tomographic inversion results for $V_S$ and $\xi$ are presented in figures (to be defined).
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=1\textwidth]{figures/Xi-60-300-for-paper.png}
+
+	\caption{\baselineskip 18pt
+	3D isotropic radial anisotropy structure of the continent.
+Map views are shown from 60 km down to 300 km, shown as variations with respect to the regional mean (thick red line in Figure ~\ref{1daverage}).
+	% Labels are: BP, Blake Plateau; NQ, New Quebec Orogen; THO, Trans-Hudson Orogen. State abbreviations are: CO, Colorado; MT, Montana; ND, North Dakota; SD, South Dakota; and WY, Wyoming. 
+	% The geological limits of the cratonic crust are indicated as follows: (to be defined when using final figures... 
+	}
+
+	\label{3d-Xi}
+
+\end{figure}
+
+The tomographic inversion results for $V_S$ and $\xi$ are presented in figures ~\ref{3d-VS,3d-Xi}.
 We divide the model description and interpretation into two parts, focusing first on the continent-wide structure and then, in more detail, on eastern different regions.
@@
 	\centering
-	\includegraphics[width=1\textwidth]{figures/1D_profile_withmean.png}
-
-	\caption{1D profiles of the global mean and the continental mean}
+	\includegraphics[width=1\textwidth]{figures/1D_profile_with_regions.png}
+
+	\caption{\baselineskip 18pt 
+	1D average depth profiles for isotropic shear velocity (left) and radial anisotropy (right), for the SEMucb\_wm1 global model (thick black), the continental average (thick red); and depth profiles for different regions in the model (see legend and Figure~\ref{mapgeol}).
+The continental average is always higher than the global average in terms of $V_S$, while for $\xi$, they are similar deeper than $250 \: km$.}
 	\label{1daverage}
@@
 
-As observed in many other studies, average profiles of $V_S$ within North America (NA) in the shallow upper mantle is higher compared to Global average profiles biased by ocean structures as shown in Figure ~\ref{1daverage}.
-The difference is clearly distinct until ca. $270km$ and then fade away deeper, though, NA remains slightly higher than the global average. 
-The global negative gradient of $V_S$ near $150km$ is here less pronounced beneath the continent by the influence of the positive gradient beneath the cratonic part. 
-NA average of $\xi$ is lower than the global average, the latter influenced by the strong $\xi > 1$ oceanic signature where shearing is more important. 
-Both trends are similar, while maxima and minima are always slightly deeper than globally and the change from $V_{SH} > V_{SV}$ to $V_{SH} < V_{SV}$ is reached at ca. $420km$.
+As observed in many other studies, the average isotropic shear velocity $V_S$ in the shallow upper mantle beneath North America (NA) is high compared to the global average, which contains contributions from large oceanic regions (Figure ~\ref{1daverage}).
+The difference persists until ca. $270 \: km$ depth and then fades away, though, $V_S$ remains slightly higher in NA than in the global average. 
+Compared to the continental average, precambrian provinces like Wyoming, Yavapai, Mazatzal, Great Rhyolites and Grenville are caracterized by slower velocities. 
+The cordillera region shows the slowest velocities, close the global average as for $\xi$. 
+Closer to the center of the continent, region like Slave, T.H.O, Juvenile arcs and orogens (e.g.
+Wopmay), Superior; show higher velocities with the Slave province caracterized by a positive gradient above $100 \: km$ deep.
+
+%The global negative gradient of $V_S$ near $150km$ is here less pronounced beneath the continent by the influence of the positive gradient beneath the cratonic part.
+(I don't think we need stress this -discussing differences between east and west will be more interesting.
+
+The average of $\xi$ is higher than $1$ (which means that $V_{SH} > V_{SV}$), but lower than the global average, the latter influenced by the strong $\xi > 1$ oceanic signature.
+% where shearing is more important. 
+Both trends are similar with depth, but the maximum in $\xi$ is slightly deeper than in the global average and the change from $V_{SH} > V_{SV}$ to $V_{SH} < V_{SV}$ is reached at ca.
+$420km$.
+{\color{red} This feature was already present in the global model, and perturbations below that depth are <1\% So I dont think the model adds more informations compared to SEMucb}.
+Regions showing highest $\xi$ values
 
@@
 
-Our study confirms the presence of a thick cratonic root with faster than average $V_S$ down to $250km$ deep, with highest velocities, ca.
-$4.8km.s^{-1}$, located at depths shalower than at most $150km$.
-Laterally, the cratonic root in terms of $V_S$, coincides with precambrian aged crust. 
-The western and eastern/south eastern edges follow the Rocky Mountain Front (RMF) and the Llano-Grenville Front respectively. 
-The northern part is delimited by the coastine of the Arctic ocean including the arctic archipelago and Greenland. 
-These lateral edges are well observed until $150km$, below, the cratonic root retracts toward the center of the archaean continent beneath Slave, Hearne/Rae and western Superior cratons. 
-Everywhere within the lithosphere, a negative gradient of $V_S$ is observed with a minimum of ca. $4.6km.s^{-1}$ reached at $300km$. 
-Deeper, this gradient switches to a positive trend.
-While the beginning of the negative gradient shows a topography, the bottom of the gradient located beneath Slave, Hearne/Rae and western Superior cratons; appears very flat.
-As observed in \cite{gung2003global}; down to 250km negative anomalies of $dln(\xi)$ are observed in the whole craton in contrast with the oceanic surrounding. 
-Deeper, positive anomalies are emerging beneath archaean Slave/Rae blocks and proterozoic Yavapai/Mazatzal/Great Rhyolites provinces, however $dln\xi$ anomalies remain positives beaneath the archaean Superior block.
-In the whole lithosphere, $\xi$ is higher than $1$ which means that $V_{SH} > V_{SV}$.
-
-\subsubsection{Wyoming}{h}
+Our study confirms the presence of a thick lithospheric root in the craton with faster than average $V_S$ down to $250km$ depth, with highest velocities, ca.
+$4.8km.s^{-1}$, located at depths shallower than  $150km$ (e.g.
+Figure~\ref{cratoncross}).
+Laterally, the cratonic root in $V_S$ coincides roughly with precambrian aged crust (see Figure~\ref{3d-VS} at from 60 to 150km deep)
+The western edge of the craton follows the RMF, instead of the western breakup line (See the red thin line in Figure~\ref{mapgeol} which coincides with RMF in Canada, but is diverging westward in the U.S.).
+While the eastern/south-esatern coincides with the Llano-Grenville Front (See the blue thin line in Figure~\ref{mapgeol}).
+The northern part ends at the coastline of the Arctic ocean including the arctic archipelago and Greenland.
+
 \begin{figure}
 	\begin{minipage}{0.5\linewidth}
-		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_wyoming_A.pdf}}
+		\centerline{\includegraphics[width=\linewidth]{figures/profiles_WE-50.pdf}}
 	\end{minipage}
@@
 	\begin{minipage}{0.5\linewidth}
-		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_wyoming_B.pdf}}
+		\centerline{\includegraphics[width=\linewidth]{figures/profiles_SN-100.pdf}}
 	\end{minipage}
 
-	\caption{Cross Section of the Wyoming lithosphere}
-	\label{wyomingcross}
+	\caption{Cross Section of the lithospheric core showing the core retraction}
+	\label{cratoncrossedge}
 
@@
 
-The Archaean aged Wyoming craton does not show as high anomalies as within the craton nucleus as shown in figure ~\ref{wyomingcross}. 
+These lateral edges are well observed down to $150km$ depth.
+Below that, the cratonic root retracts toward the center of the archaean continent and is still present beneath Slave, Hearne/Rae and western Superior cratons as shown in Figure~\ref{cratoncrossedge}.
+Everywhere within the craton, a negative gradient of $V_S$ is observed, marking the base of the lithosphere, with a minimum of $\sim 4.6km.s^{-1}$ reached at $\sim 300km$ depth.
+While the base of the lithosphere (and top of the low velocity zone- LVZ) shows strong topography, the bottom of the LVZ appears largely flat, in particular under the  Slave, Hearne/Rae and western Superior cratons.
+As observed in \cite{babuvska1998age,gung2003global,yuan2014lithospheric}, there are large lateral variations  $\xi$ down to $\sim 200km$ depth, with regions of reduced radial anisotropy, in contrast to the more uniform, positive $dln(\xi)$ structure in this depth range in the oceans.
+Generally, $\xi$ always appear higher than 1.
+Around 100 km negative $dln(\xi)$ compared to the isotropic case are observed in Mexico, at the southern part of Florida, the southern part of the Superior craton comprising the Mid Continental rifting, the southern part of the Granit Rhyiolite Province comprising the Failed Rifting area and finally beneath Baffin Island.
+
+\subsubsection{Wyoming}{h}
+% \begin{figure}
+% 	\begin{minipage}{0.5\linewidth}
+% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_wyoming_A.pdf}}
+% 	\end{minipage}
+% 	\hfill
+% 	\begin{minipage}{0.5\linewidth}
+% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_wyoming_B.pdf}}
+% 	\end{minipage}
+
+% 	\caption{Cross Section of the Wyoming lithosphere}
+% 	\label{wyomingcross}
+
+% \end{figure}
+
+The Archaean aged Wyoming craton does not show as high anomalies as within the craton nucleus as shown in figure ~\ref{3d-VS}. 
 It is split into 2 distinctive part. A western part showing low velocities influenced by the active tectonic with a radially positive gradient. 
@@
 
-	\begin{figure}
-		\begin{minipage}{0.5\linewidth}
-			\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_slave_A.pdf}}
-		\end{minipage}
-		\hfill
-		\begin{minipage}{0.5\linewidth}
-			\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_slave_B.pdf}}
-		\end{minipage}
-
-		\caption{Cross Section of the Slave lithosphere}
-		\label{slavecross}
-
-	\end{figure}
-
-	The Archaean aged Slave craton shows typical lithospheric craton signatures with higher than average velocities as shown in figure ~\ref{slavecross}. 
+	% \begin{figure}
+	% 	\begin{minipage}{0.5\linewidth}
+	% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_slave_A.pdf}}
+	% 	\end{minipage}
+	% 	\hfill
+	% 	\begin{minipage}{0.5\linewidth}
+	% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_slave_B.pdf}}
+	% 	\end{minipage}
+
+	% 	\caption{Cross Section of the Slave lithosphere}
+	% 	\label{slavecross}
+
+	% \end{figure}
+
+	The Archaean aged Slave craton shows typical lithospheric craton signatures with higher than average velocities as shown in figure ~\ref{3d-VS} and in figure~\ref{cratoncross}. 
 	It shows in the entire region a positive velocity gradient with a maximum of ca. $4.8km.s^{-1}$ associated with a maximum of $\xi$ at ca. $150km$ deep.
@@
 	\paragraph{Western part}
-		\begin{figure}
-			\begin{minipage}{0.5\linewidth}
-				\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_west_superior_A.pdf}}
-			\end{minipage}
-			\hfill
-			\begin{minipage}{0.5\linewidth}
-				\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_west_superior_B.pdf}}
-			\end{minipage}
-
-			\caption{Cross Section of the western Superior lithosphere}
-			\label{westsupcross}
-
-		\end{figure}
-
-		The Archaean aged Superior craton shows typical lithospheric craton signatures with higher than average velocities as shown in figure ~\ref{westsupcross}. 
+		% \begin{figure}
+		% 	\begin{minipage}{0.5\linewidth}
+		% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_west_superior_A.pdf}}
+		% 	\end{minipage}
+		% 	\hfill
+		% 	\begin{minipage}{0.5\linewidth}
+		% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_west_superior_B.pdf}}
+		% 	\end{minipage}
+
+		% 	\caption{Cross Section of the western Superior lithosphere}
+		% 	\label{westsupcross}
+
+		% \end{figure}
+
+		The Archaean aged Superior craton shows typical lithospheric craton signatures with higher than average velocities as shown in figure ~\ref{3d-VS} and in figure~\ref{cratoncross}. 
 		It shows in the entire region a positive velocity gradient with a maximum of ca. $4.8km.s^{-1}$  associated with a maximum of $\xi$ at ca. $150km$ deep.
@@
 		The band of minimum velocity within the cratonic root has similar amplitudes, in terms of velocity, than beneath the younger lithosphere but at shalower depths between $100$ and $150km$. 
-		Folowing the Great Meteor track built by \cite{heaman2000timing}, which crosses the Superior craton, it is interseting to note that the low velocity channel beneath the base of the lithosphere shows lower velocties than usualy observed beneath the other Archaean cratons.
+		Following the Great Meteor track built by \cite{heaman2000timing}, which crosses the Superior craton, it is interseting to note that the low velocity channel beneath the base of the lithosphere shows lower velocties than usualy observed beneath the other Archaean cratons.
 		This slow anomaly corresponds to the V-shape dent, or divot, observed by \cite{lee1997upper} who explained the feature by the presence of water in the mantle. 
@@
 		\paragraph{Eastern part}
-		\begin{figure}
-			\begin{minipage}{0.5\linewidth}
-				\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_east_superior_A.pdf}}
-			\end{minipage}
-			\hfill
-			\begin{minipage}{0.5\linewidth}
-				\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_east_superior_B.pdf}}
-			\end{minipage}
-
-			\caption{Cross Section of the western Superior lithosphere}
-			\label{eastsupcross}
-
-		\end{figure}
-
-		Compared to its western part, the structure here shows differences as shown in figure ~\ref{eastsupcross}. . 
+		% \begin{figure}
+		% 	\begin{minipage}{0.5\linewidth}
+		% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_east_superior_A.pdf}}
+		% 	\end{minipage}
+		% 	\hfill
+		% 	\begin{minipage}{0.5\linewidth}
+		% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_east_superior_B.pdf}}
+		% 	\end{minipage}
+
+		% 	\caption{Cross Section of the western Superior lithosphere}
+		% 	\label{eastsupcross}
+
+		% \end{figure}
+
+		Compared to its western part, the structure here shows differences as shown in figure ~\ref{cratoncrossedge} 
 		Fast velocities are present down to $150km$deep but are slower. 
@@
 While most of the proterozoic root start to thin out at $150km$, the THO at the vicinity of the hudson bay shows as high anomalies as Rae/Hearne and Superior areas.
-
-\subsubsection{Yavapai Mazatsal region}
-	\begin{figure}
-		\begin{minipage}{0.5\linewidth}
-			\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_yava-maza_A.pdf}}
-		\end{minipage}
-		\hfill
-		\begin{minipage}{0.5\linewidth}
-			\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_yava-maza_B.pdf}}
-		\end{minipage}
-
-		\caption{Cross Section of the Yavapai and Mazatzal lithosphere}
-		\label{yavacross}
-
-	\end{figure}
-
-	Yavapai-Mazatzal orogens from 1.8 to 1.6 Ga represent zones of accretion for Proterozoic material and mark the amalgamation of the archean cores. \citep{hoffman1988united} \\
-	Cross section of the region are represented in figure ~\ref{yavacross}.
+	Je suis en train de construire des profils le long de THO pour voir!
+
+\subsubsection{Yavapai/Mazatzal/Granit-Rhyolite region}
+	% \begin{figure}
+	% 	\begin{minipage}{0.5\linewidth}
+	% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_yava-maza_A.pdf}}
+	% 	\end{minipage}
+	% 	\hfill
+	% 	\begin{minipage}{0.5\linewidth}
+	% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_yava-maza_B.pdf}}
+	% 	\end{minipage}
+
+	% 	\caption{Cross Section of the Yavapai and Mazatzal lithosphere}
+	% 	\label{yavacross}
+
+	% \end{figure}
+
+	Yavapai/Mazatzal/Granit-Rhyolite orogens from 1.8 to 1.6 Ga represent zones of accretion for Proterozoic material and mark the amalgamation of the archean cores. \citep{hoffman1988united} \\
 	West to the RMF, the region shows low velocities influenced by the active tectonic with a radially positive gradient. 
@@
 
-\subsection{Surrounding oceanic}
-	Our inversion is focusing on the continental part, as shown in our spatial parameterization (i.e. see figure...) where the region is sampled by data.
+\subsection{Surrounding oceanic} This section needs to be redone...
+
+	Our inversion is focusing on the continental part, as shown in our spatial parameterization (i.e. see figure~\ref{density_coverage}) where the region is sampled by data.
 	Ecxept the oceanic borders of NA, further away, the structure has not been updated from SEMucb\_wm1 \citep{french2014whole}.

### 2017-02-28 Paper_NA_br.tex
No Linked figures

@@
 
-\begin{figure}[ht]
+\begin{figure}[p]
 	\centering
@@
 
-During the inversion, all the synthetic seismograms are computed using the regional SEM code RegSEM \citep{cupillard2012regsem} that takes into account effects of oceans, topography/bathymetry, ellipticity, and anelasticity, and where the limits of the computational domain are dealt with using absorbing boundaries.
+During the inversion, all the synthetic seismograms are computed using the regional SEM code RegSEM \citep{cupillard2012regsem} that takes into account effects of oceans, topography/bathymetry, ellipticity, and anelasticity, and where the limits of the computational domain (white box around North America in Figure~\ref{injection} and red box in Figure ~\ref{map_evt_stn}) are dealt with using absorbing boundaries.
 The synthetic seismograms associated with regional data (i.e.
@@
 These simulations are done once and for all before the inversion starts.
+A comparison between regular $SPECFEM3D\_globe$ and ``RegSEM + virtual sources`` synthetics is shown in Figure~\ref{injection}.
 
@@
 	3D isotropic shear wave velocity structure of the continent.
-Map views are shown from 60 km down to 300 km, shown as variations with respect to the global mean (thick red line in Figure ~\ref{1daverage}).
+Map views are shown from 60 km down to 300 km, shown as variations with respect to the regional mean (thick red line in Figure ~\ref{1daverage}).
 	% Labels are: BP, Blake Plateau; NQ, New Quebec Orogen; THO, Trans-Hudson Orogen. State abbreviations are: CO, Colorado; MT, Montana; ND, North Dakota; SD, South Dakota; and WY, Wyoming. 
@@
 
-The tomographic inversion results for $V_S$ and $\xi$ are presented in figures ~\ref{3d-VS,3d-Xi}.
+The tomographic inversion results for $V_S$ and $\xi$ are presented in figures \ref{3d-VS} and \ref{3d-Xi} respectvely. 
 We divide the model description and interpretation into two parts, focusing first on the continent-wide structure and then, in more detail, on eastern different regions.
@@
 % where shearing is more important. 
-Both trends are similar with depth, but the maximum in $\xi$ is slightly deeper than in the global average and the change from $V_{SH} > V_{SV}$ to $V_{SH} < V_{SV}$ is reached at ca.
-$420km$.
+Both trends are similar with depth, but the maximum in $\xi$ is slightly deeper than in the global average and the change from $V_{SH} > V_{SV}$ to $V_{SH} < V_{SV}$ is reached at ca. $420km$.
+
 {\color{red} This feature was already present in the global model, and perturbations below that depth are <1\% So I dont think the model adds more informations compared to SEMucb}.
-Regions showing highest $\xi$ values
+
+I am on this part..
 
@@
 Figure~\ref{cratoncross}).
-Laterally, the cratonic root in $V_S$ coincides roughly with precambrian aged crust (see Figure~\ref{3d-VS} at from 60 to 150km deep)
-The western edge of the craton follows the RMF, instead of the western breakup line (See the red thin line in Figure~\ref{mapgeol} which coincides with RMF in Canada, but is diverging westward in the U.S.).
+Laterally, the cratonic root in $V_S$ coincides roughly with precambrian aged crust (see Figure~\ref{3d-VS} from 60 to 150km deep)
+The western edge of the craton follows the RMF (black thick dashed line in Figure \ref{mapgeol}), instead of the western breakup line (See the red thin line in Figure~\ref{mapgeol} which coincides with RMF in Canada, but is diverging westward in the U.S.).
 While the eastern/south-esatern coincides with the Llano-Grenville Front (See the blue thin line in Figure~\ref{mapgeol}).
@@
 These lateral edges are well observed down to $150km$ depth.
-Below that, the cratonic root retracts toward the center of the archaean continent and is still present beneath Slave, Hearne/Rae and western Superior cratons as shown in Figure~\ref{cratoncrossedge}.
+Below that, the cratonic root retracts toward the center of the archaean continent and is still present beneath Slave, Hearne/Rae and western Superior cratons as shown in Figure \ref{cratoncrossedge}.
 Everywhere within the craton, a negative gradient of $V_S$ is observed, marking the base of the lithosphere, with a minimum of $\sim 4.6km.s^{-1}$ reached at $\sim 300km$ depth. 
 While the base of the lithosphere (and top of the low velocity zone- LVZ) shows strong topography, the bottom of the LVZ appears largely flat, in particular under the  Slave, Hearne/Rae and western Superior cratons.
-As observed in \cite{babuvska1998age,gung2003global,yuan2014lithospheric}, there are large lateral variations  $\xi$ down to $\sim 200km$ depth, with regions of reduced radial anisotropy, in contrast to the more uniform, positive $dln(\xi)$ structure in this depth range in the oceans.
+
+As observed in \cite{babuvska1998age,gung2003global,yuan2014lithospheric}, there are large lateral variations  $\xi$ down to $\sim 200km$ depth, with regions of reduced radial anisotropy, in contrast to the more uniform, positive $dln(\xi)$ structure in this depth range in the oceans (See figure \ref{3d-Xi}.
+Below 250km, it is interesting to note that there is a large band of positive anomaly crossing the whole continent (across the US), making the connection between the Pacific and Atlantic mantles.
+
 Generally, $\xi$ always appear higher than 1. 
-Around 100 km negative $dln(\xi)$ compared to the isotropic case are observed in Mexico, at the southern part of Florida, the southern part of the Superior craton comprising the Mid Continental rifting, the southern part of the Granit Rhyiolite Province comprising the Failed Rifting area and finally beneath Baffin Island.
+Except around 100 km, negative absolute values of $\xi$ (lower than -$1\%$) are observed in Mexico, at the southern part of Florida, the southern part of the Superior craton comprising the Mid Continental rifting, the southern part of the Granit Rhyiolite Province comprising the Failed Rifting area and finally beneath Baffin Island.
 

### 2017-03-01 Paper_NA_br.tex
Update 01.03.17

@@
 \usepackage{natbib} % Powerful citation management
+% \usepackage{float}
+% \floatplacement{figure}{p}
 \bibpunct[; ]{(}{)}{,}{a}{}{;} % Custom minimal bibliography styling
@@
 \section{Introduction}
-
-\begin{figure}[p]
-	\centering
-	\centerline{\includegraphics[viewport=180 340 480 570,clip=true,width=\linewidth]{figures/Geologicalmap_Clouzet.pdf}}
-
-	% bb=llx lly urx ury :
-
-	\caption{\baselineskip 18pt
-	Large scale Precambrian architecture of North America (modified from \cite{whitmeyer2007tectonic}.
-Colours indicate ages of different regions, as shown in the legend.
-The stable cratonic core is defined by the Rocky Mountain Front to the west (dashed dark grey line), defined by the surface trace of deformation), and the appalachian fron to the east (solid light red line).
-The solid blue line demarcates the westward extent of the Grenville/Llano Front.
-The solid red line along the Cordillera province demarcates the breakup of Rodinia.
-Solid green lines indicate tectonic plate boundaries.
-Abreviations are as follows: RW, re-worked; GBA, Great Bear Arc; BHT, Buffalo Head Terrane; Taltson, Talston Arc; Rimbey, Rimbey Arc; Tornagat, Torngat Arc; L.B.A, Little Belt Arc; LRA; THO, Trans-Hudson Orogen; RGR, Rio Grande Rift; Reelfoot, Reelfoor Rift; Kwn.R, Kewanowan Rift; Sask, Sask Craton; WY, Wyoming Province; Med.
-Hat., Medicine Hat Block; Grt.
-Rhy, Granite–Rhyolite Province; Grenville, Grenville Province.
-	}
-
-	\label{mapgeol}
-
-\end{figure}
 
@@
 
-\begin{figure}[p]
-	\centering
-	\begin{tabular}{ccc}
-		\includegraphics[width=1\textwidth]{figures/map_events_stn_global.png}
-	\end{tabular}
-
-	\caption{\baselineskip 18pt 
-	Source and station distribution for the new North American inversion.
-Black triangles show the seismic stations.
-Stars are 155 regional events, in addition to 122 teleseismic events.
-Orange stars indiscate events shallower than $30km$ deep, Yellow stars for deeper than $300km$ and red one are in between.
-Thick red line indicates the boundaries used for the RegSEM forward simulation, which extends 89◦ by 89◦ horizontally and down to 1600 km and contains all stations and regional events.}
-
-	\label{map_evt_stn}
-\end{figure}
-
 The dataset includes three component acceleration waveforms from 2860 permanent and temporary broadband stations located in the target region, a $89^o \times 89^o $ area encompassing most of north America (Figure ~\ref{map_evt_stn}).
@@
 The total number of wavepackets of fundamental and overtone Love and Rayleigh waves is given in Table 1, and provides a good coverage within the NA continent (Figure ~\ref{density_coverage}).
-
-\begin{figure}[p]
-	% bb=llx lly urx ury :
-	\begin{minipage}{0.47\linewidth}
-		\centerline{\includegraphics[viewport=120 10 470 420,clip=,width=\linewidth]{figures/Coverage_L_iter3.png}}
-	\end{minipage}
-	\hfill
-	\begin{minipage}{0.47\linewidth}
-		\centerline{\includegraphics[viewport=120 10 470 420,clip=,width=\linewidth]{figures/Coverage_T_iter3.png}}
-	\end{minipage}
-
-	\begin{minipage}{0.47\linewidth}
-		\centerline{\includegraphics[viewport=120 10 470 420,clip=,width=\linewidth]{figures/Coverage_Z_iter3.png}}
-	\end{minipage}
-	\hfill
-	\begin{minipage}{0.47\linewidth}
-		\centerline{\includegraphics[viewport=120 10 470 420,clip=,width=\linewidth]{figures/Coverage_all_iter3.png}}
-	\end{minipage}
-
-\caption{Density coverage for each component and all gathered.} 
-\label{density_coverage}
-
-\end{figure}
 
@@
 
-\begin{figure}[p]
-% \centering
-	\begin{tabular}{ccc}
-		(a)&&(b)\\
-		\includegraphics[width=.4\textwidth]{figures/snap_glob.pdf} && \includegraphics[width=.4\textwidth]{figures/snap_reg.pdf} \\ 
-		\includegraphics[width=.4\textwidth]{figures/seismo.pdf}    && \includegraphics[width=.4\textwidth]{figures/seismo.pdf}
-	\end{tabular}
-
-	\caption{\baselineskip 18pt 
-		Comparison between a global scale simulation and a regional scale simulation of wave propagation following an earthquake in south America.
-The seismograms corresponds to the recordings at 37 station located in north America (red spheres).
-		The pink sphere shows the epicenter of the earthquake.
-In (a), the wavefield is modeled globally using the spectral element method Specfem3D\_globe \citep{komatitsch2002spectrala}.
-In (b), the wavefield is modeled regionally using the RegSEM.
-During the inversion, all the synthetic seismograms are computed using the regional SEM code RegSEM \citep{cupillard2012regsem} code and regenerated thanks to virtual sources located around the computational domain \citep[see][]{masson2013numerical}.
-The sismograms in (b) are exactly similar to those in (a), however, the computational effort is greatly reduced in the regional simulation (b).}
-
-	\label{injection}
-
-\end{figure}
-
 During the inversion, all the synthetic seismograms are computed using the regional SEM code RegSEM \citep{cupillard2012regsem} that takes into account effects of oceans, topography/bathymetry, ellipticity, and anelasticity, and where the limits of the computational domain (white box around North America in Figure~\ref{injection} and red box in Figure ~\ref{map_evt_stn}) are dealt with using absorbing boundaries.
@@
 
-\begin{figure}[ht]
-	\centering
-	\includegraphics[width=1\textwidth]{figures/Vs-60-300-for-paper.png}
-
-	\caption{\baselineskip 18pt
-	3D isotropic shear wave velocity structure of the continent.
-Map views are shown from 60 km down to 300 km, shown as variations with respect to the regional mean (thick red line in Figure ~\ref{1daverage}).
-	% Labels are: BP, Blake Plateau; NQ, New Quebec Orogen; THO, Trans-Hudson Orogen. State abbreviations are: CO, Colorado; MT, Montana; ND, North Dakota; SD, South Dakota; and WY, Wyoming. 
-	% The geological limits of the cratonic crust are indicated as follows: (to be defined when using final figures... 
-	}
-
-	\label{3d-VS}
-
-\end{figure}
-
-\begin{figure}[ht]
-	\centering
-	\includegraphics[width=1\textwidth]{figures/Xi-60-300-for-paper.png}
-
-	\caption{\baselineskip 18pt
-	3D isotropic radial anisotropy structure of the continent.
-Map views are shown from 60 km down to 300 km, shown as variations with respect to the regional mean (thick red line in Figure ~\ref{1daverage}).
-	% Labels are: BP, Blake Plateau; NQ, New Quebec Orogen; THO, Trans-Hudson Orogen. State abbreviations are: CO, Colorado; MT, Montana; ND, North Dakota; SD, South Dakota; and WY, Wyoming. 
-	% The geological limits of the cratonic crust are indicated as follows: (to be defined when using final figures... 
-	}
-
-	\label{3d-Xi}
-
-\end{figure}
-
 	The tomographic inversion results for $V_S$ and $\xi$ are presented in figures \ref{3d-VS} and \ref{3d-Xi} respectvely. 
@@
 	\subsection{Continental model}
-\begin{figure}[ht]
-	\centering
-	\includegraphics[width=1\textwidth]{figures/1D_profile_with_regions.png}
-
-	\caption{\baselineskip 18pt 
-	1D average depth profiles for isotropic shear velocity (left) and radial anisotropy (right), for the SEMucb\_wm1 global model (thick black), the continental average (thick red); and depth profiles for different regions in the model (see legend and Figure~\ref{mapgeol}).
-The continental average is always higher than the global average in terms of $V_S$, while for $\xi$, they are similar deeper than $250 \: km$.}
-	\label{1daverage}
-\end{figure}
 
@@
 
-I am on this part..
-
 	\subsection{Beneath the craton}
-
-\begin{figure}
-	\begin{minipage}{0.5\linewidth}
-		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_A.pdf}}
-	\end{minipage}
-	\hfill
-	\begin{minipage}{0.5\linewidth}
-		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_B.pdf}}
-	\end{minipage}
-
-	\caption{Cross Section of the lithospheric core}
-	\label{cratoncross}
-
-\end{figure}
 
@@
 		The western edge of the craton follows the RMF (black thick dashed line in Figure \ref{mapgeol}), instead of the western breakup line (See the red thin line in Figure~\ref{mapgeol} which coincides with RMF in Canada, but is diverging westward in the U.S.).
-While the eastern/south-esatern coincides with the Llano-Grenville Front (See the blue thin line in Figure~\ref{mapgeol}).
+While the eastern/south-eastern coincides with the Llano-Grenville Front (See the blue thin line in Figure~\ref{mapgeol}).
 		The northern part ends at the coastline of the Arctic ocean including the arctic archipelago and Greenland.
-
-\begin{figure}
-	\begin{minipage}{0.5\linewidth}
-		\centerline{\includegraphics[width=\linewidth]{figures/profiles_WE-50.pdf}}
-	\end{minipage}
-	\hfill
-	\begin{minipage}{0.5\linewidth}
-		\centerline{\includegraphics[width=\linewidth]{figures/profiles_SN-100.pdf}}
-	\end{minipage}
-
-	\caption{Cross Section of the lithospheric core showing the core retraction}
-	\label{cratoncrossedge}
-
-\end{figure}
 
@@
 
-As observed in \cite{babuvska1998age,gung2003global,yuan2014lithospheric}, there are large lateral variations  $\xi$ down to $\sim 200km$ depth, with regions of reduced radial anisotropy, in contrast to the more uniform, positive $dln(\xi)$ structure in this depth range in the oceans (See figure \ref{3d-Xi}.
+		As observed in \cite{babuvska1998age,gung2003global,yuan2014lithospheric}, there are large lateral negative variations $\xi$ down to $\sim 200km$ depth, with regions of reduced radial anisotropy, in contrast to the more uniform, positive $dln(\xi)$ structure in this depth range in the oceans (See figure \ref{3d-Xi}.
 		Below 250km, it is interesting to note that there is a large band of positive anomaly crossing the whole continent (across the US), making the connection between the Pacific and Atlantic mantles. 
+		This feature is not present in most of Canada, where negative $dln(\xi)$ structure persist down to 300km.
+		Generally, $\xi$ always appear higher than 1. 
+		Except between 60 and 100 km, lower than 1 of $\xi$ ($V_{SV} > V_{SH}$) are observed in Mexico, at the southern part of Florida, the southern part of the Superior craton comprising the Mid Continental rifting, the southern part of the Granit Rhyiolite Province comprising the Failed Rifting area, a large area comprising southern Hearne, Medicine hat, northern Wyoming, B.H.T, Cordillera; and finally beneath Baffin Island.
+This feature can be better observed in Figure \ref{3d-xi-iso-100km} where perturbations are shown according the Isotropic case ($\xi = 1$)
 
-Generally, $\xi$ always appear higher than 1. 
-Except around 100 km, negative absolute values of $\xi$ (lower than -$1\%$) are observed in Mexico, at the southern part of Florida, the southern part of the Superior craton comprising the Mid Continental rifting, the southern part of the Granit Rhyiolite Province comprising the Failed Rifting area and finally beneath Baffin Island.
+	\subsubsection{Archaean provinces}
 
-\subsubsection{Wyoming}{h}
-% \begin{figure}
-% 	\begin{minipage}{0.5\linewidth}
-% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_wyoming_A.pdf}}
-% 	\end{minipage}
-% 	\hfill
-% 	\begin{minipage}{0.5\linewidth}
-% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_wyoming_B.pdf}}
-% 	\end{minipage}
-
-% 	\caption{Cross Section of the Wyoming lithosphere}
-% 	\label{wyomingcross}
-
-% \end{figure}
-
-The Archaean aged Wyoming craton does not show as high anomalies as within the craton nucleus as shown in figure ~\ref{3d-VS}. 
+		\paragraph{Wyoming and Medecine Hat}
+			The highly reworked Wyoming craton does not show as high anomalies as within the craton nucleus (+2\% intstead of +4-8\%) as shown in figure ~\ref{3d-VS}. 
 			It is split into 2 distinctive part. A western part showing low velocities influenced by the active tectonic with a radially positive gradient. 
-In contrast with the eastern part showing higher velocities with a radially negative then positive gradient. 
+			In contrast with the eastern part showing higher velocities with a radially negative then positive gradient which marks the western border of the craton nucleus. 
 			The depth of negative gradient's minima is variable, from 70 (western part) to 250km (eastern part) deep. 
 			The separation between the two areas coincides with the RMF.
+			In terms of $\xi$, it is characterized by negative anomalies down to 250km deep. 
+			The northern part shows highest anomalies and $V_{SV}$ appears higher than $V_{SH}$ of about 2\% between 60 and 100km.
+This feature is common with the Medicine Hat province (See Figure \ref{3d-xi-iso-100km}).
 
-\subsubsection{Slave}
-
-	% \begin{figure}
-	% 	\begin{minipage}{0.5\linewidth}
-	% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_slave_A.pdf}}
-	% 	\end{minipage}
-	% 	\hfill
-	% 	\begin{minipage}{0.5\linewidth}
-	% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_slave_B.pdf}}
-	% 	\end{minipage}
-
-	% 	\caption{Cross Section of the Slave lithosphere}
-	% 	\label{slavecross}
-
-	% \end{figure}
-
-	The Archaean aged Slave craton shows typical lithospheric craton signatures with higher than average velocities as shown in figure ~\ref{3d-VS} and in figure~\ref{cratoncross}. 
+		\paragraph{Slave}
+			The Slave craton shows typical lithospheric craton signatures with higher than average velocities (See figures \ref{3d-VS} and \ref{cratoncross}), and negative perturbations of $\xi$ (See Figure \ref{3d-VS}.
+			We can note that the average of $\xi$ is the only one higher than the continental mean, compared to the other Archaean provinces (See Figure \ref{1daverage}.
 			It shows in the entire region a positive velocity gradient with a maximum of ca. $4.8km.s^{-1}$ associated with a maximum of $\xi$ at ca. $150km$ deep. 
 			Deeper, a negative gradient is observed with a minimum of ca. $4.6km.s^{-1}$ at ca. $250km$ deep. 
-	The topography of the extrema appears flat. (Est ce que c'est une characteristique d'un craton stable non deforme?)
+			The topography of the extrema appears flat.
 			Although the Slave lithosphere shows highest velocities at $150km$, above $100km$ velocities are slightly higher than the continental average ($dln(V_S)< 4\%$).
 
-\subsubsection{Superior Craton}
+		\paragraph{Superior Craton}
 			In our model, the Superior craton can be split in two parts: The western part showing typical cratonic structures while the eastern part shows slower velocities at all depths.
@@
 		\paragraph{Western part}
-		% \begin{figure}
-		% 	\begin{minipage}{0.5\linewidth}
-		% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_west_superior_A.pdf}}
-		% 	\end{minipage}
-		% 	\hfill
-		% 	\begin{minipage}{0.5\linewidth}
-		% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_west_superior_B.pdf}}
-		% 	\end{minipage}
-
-		% 	\caption{Cross Section of the western Superior lithosphere}
-		% 	\label{westsupcross}
-
-		% \end{figure}
-
-		The Archaean aged Superior craton shows typical lithospheric craton signatures with higher than average velocities as shown in figure ~\ref{3d-VS} and in figure~\ref{cratoncross}. 
-		It shows in the entire region a positive velocity gradient with a maximum of ca. $4.8km.s^{-1}$  associated with a maximum of $\xi$ at ca. $150km$ deep.
+			The western Superior craton shows typical lithospheric craton signatures with highest velocities as shown in figure ~\ref{3d-VS} and in figure~\ref{cratoncross}.
+In terms of $\xi$, slight negative to zero perturbations are observed at the very western part, while close to the Mid continental rifting, highest negative perturbations indicate that $V_{SV}$ is higher than $V_{SH}$.
+This last feature is associated the thinning of the cratonic lithosphere when crossing the Grenville front.
+			In the entire region there is a positive velocity gradient with a maximum of ca. $4.8km.s^{-1}$  associated with a maximum of $\xi$ at ca. $150km$ deep.
 
@@
 			Beneath the maximum velocity region of ca. $150km$ the negative gradient appears stipper than beneath the $70km$ one, while the topography of the deeper positive gradient appears flat.
-		At the south-eastern part of the western Superior, close the Mid-Lithospheric rift, we can observe the thinning of the cratonic lithosphere when crossing the Grenville front. 
-		The band of minimum velocity within the cratonic root has similar amplitudes, in terms of velocity, than beneath the younger lithosphere but at shalower depths between $100$ and $150km$. 
+
+			The band of minimum velocity within the cratonic root has similar amplitudes, in terms of velocity, than beneath the younger lithosphere but at shalower depths between $100$ and $150 \: km$. 
 			Following the Great Meteor track built by \cite{heaman2000timing}, which crosses the Superior craton, it is interseting to note that the low velocity channel beneath the base of the lithosphere shows lower velocties than usualy observed beneath the other Archaean cratons.
-		This slow anomaly corresponds to the V-shape dent, or divot, observed by \cite{lee1997upper} who explained the feature by the presence of water in the mantle. 
+			This slow anomaly could correspond to an extension the V-shape dent, or divot, observed by \cite{lee1997upper} who explained the feature by the presence of water in the mantle. 
 			We observe at $250km$ that the slow anomaly go further inland and crosses the Superior Craton northward until the Hudson bay following the GMT.
@@
 		\paragraph{Eastern part}
-		% \begin{figure}
-		% 	\begin{minipage}{0.5\linewidth}
-		% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_east_superior_A.pdf}}
-		% 	\end{minipage}
-		% 	\hfill
-		% 	\begin{minipage}{0.5\linewidth}
-		% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_east_superior_B.pdf}}
-		% 	\end{minipage}
-
-		% 	\caption{Cross Section of the western Superior lithosphere}
-		% 	\label{eastsupcross}
-
-		% \end{figure}
-
 			Compared to its western part, the structure here shows differences as shown in figure ~\ref{cratoncrossedge} 
-		Fast velocities are present down to $150km$deep but are slower. 
-		The negative gradient is steeper and the minimum is radially thicker. 
-		However at the center, the minimum channel is not continuous and shows higher (still slower than the center of the western Superior) velocities.
+			Fast velocities are present down to $150 \: km$ deep but are slower and no perturbations are observed below $200 \: km$. 
+			The negative gradient is steeper and the minimum is radially thicker (more than $100 \: km$ thick compare to $\sim 50 \: km$ beneath the craton nucleus). 
+			% However at the center, the minimum channel is not continuous and shows higher (still slower than the center of the western Superior) velocities.
 			% the gradual trend in velocity is replaced by an almost constant slow anomaly from $150$ to $300km$ deep.
 
-\subsubsection{Trans Hudson Orogen}
-	The Trans Hudson Orogen (THO) is a belt of Paleo-Proterozoic aged region. It refers to a major tectonic event (between $2.0$ to $1.8Ga$) that welded Archaean blocks to form the Laurentia Craton.
+		\paragraph{Trans Hudson Orogen}
+			The Trans Hudson Orogen (THO) is a belt of reworked archaean aged region.
+It refers to a major Paleo Proterozoic tectonic event (between $2.0$ to $1.8Ga$) that welded the Hearne, Wyoming, and Superior archaean blocks to form core of Laurentia \citep{hoffman1988united,whitmeyer2007tectonic}.
 			% At shallow depths ($ < 60km$) shows lower velocities than archaean surroundings. 
-	It is interesting to notice that at shallower depths than $100km$, the proterozoic THO (including the Sask craton) which welded the Superior and Rae/Hearne cratons, shows lighter (i.e, $+4\%$ compared to $+8\%$) positive anomalies than its archaean surrounding.
-While most of the proterozoic root start to thin out at $150km$, the THO at the vicinity of the hudson bay shows as high anomalies as Rae/Hearne and Superior areas.
-	Je suis en train de construire des profils le long de THO pour voir!
+			It shows typical lithospheric craton signatures with higher than average velocities (See figures \ref{3d-VS} and \ref{cratoncross} associated with negative $\xi$ perturbations. 
+			It is interesting to notice that at shallower depths than $100km$, the THO (including the Sask craton) which welded the Superior and Rae/Hearne cratons, shows lighter (i.e, $+4\%$ compared to $+8\%$) positive anomalies than its archaean surrounding.
+This feature is associated with negative anomalies of $\xi$.
+While most of the proterozoic root start to thin out at $150km$, the THO at the vicinity of the hudson bay shows as high anomalies as Rae/Hearne and Superior areas associated with zero to positive anomalies in term of $\xi$.
 
 	\subsubsection{Yavapai/Mazatzal/Granit-Rhyolite region}
-	% \begin{figure}
-	% 	\begin{minipage}{0.5\linewidth}
-	% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_yava-maza_A.pdf}}
-	% 	\end{minipage}
-	% 	\hfill
-	% 	\begin{minipage}{0.5\linewidth}
-	% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_yava-maza_B.pdf}}
-	% 	\end{minipage}
-
-	% 	\caption{Cross Section of the Yavapai and Mazatzal lithosphere}
-	% 	\label{yavacross}
-
-	% \end{figure}
 
@@
 
-\subsection{Surrounding oceanic} This section needs to be redone...
+\subsubsection{Surrounding oceanic} This section needs to be redone...
 
 	Our inversion is focusing on the continental part, as shown in our spatial parameterization (i.e. see figure~\ref{density_coverage}) where the region is sampled by data.
-	Ecxept the oceanic borders of NA, further away, the structure has not been updated from SEMucb\_wm1 \citep{french2014whole}.
+	Except the oceanic borders of NA, further away, the structure has not been updated from SEMucb\_wm1 \citep{french2014whole}.
 
@@
 		As in \cite{schaeffer2014imaging,yuan2014lithospheric,james2014rayleigh,silveira2002anisotropic,mocquet1990three}, we observe along the negative anomaly structure following the Appalachian continental shelf, a stretched positive anomaly structure from the Suwanne terrane in Florida to the Avalonian terranes of New-Scotland.
-		At $60km$ it is made of several blocks of at most $+3\%$ anomalies emdedded in a $+2\%$ structure.
+		At $60km$ it is made of several blocks of at most $+3\%$ anomalies emdedded in a $+2\%$ structure, associated with zero to positive $\xi$ anomalies.
 At $150km$, the structure is split into 2 blocks by a slow anomaly comprising both the Great Meteor (GMT) and Bermuda tracks, and further southeast, the Low Velocities Fingers (LVF) observed in \cite{french2013waveform}.
@@
 		On the continental shelf, authors \citep[e.g.][]{o1983avalon,nance2002cordilleran} identified the west Avalonian terranes of pan-african affinity as in Florida \citep[see][]{smith1982review} that might have been underlayed by a thin (ca.
-$100km$) lithosphere before the Pangaea assembling.
+$100 \: km$) lithosphere before the Pangaea assembling.
 \citep{mckenzie2015lithospheric}
 
-		Fast anomalies are observed beneath the whole gulf of Mexico down to $150km$ deep.
-Despite the proximity with pan-african lithosphere, such high velocity anomalies represent older than $ca.
-150Ma$ oceanic lithosphere.
+		Fast veolicity associated with typical oceanic radial anisotropic structures are observed beneath the whole gulf of Mexico down to $150 \: km$ deep.
+Despite the proximity with pan-african lithosphere, such high velocity anomalies represent older than $\sim 150 \: Ma$ oceanic lithosphere.
 \citep{muller2008age,pindell2009tectonic}
@@
 \subsubsection{The Great Meteor track}
-	\begin{figure}
-		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_GMTrack.pdf}}
-
-		\caption{Cross Section of the lithosphere along the Great Meteor track based on \cite{heaman2000timing} construction}
-		\label{gmtcross}
-
-	\end{figure}
 
@@
 \citep{rondenay2000lithospheric}
-% This hypothesis is based on different points:
-% \begin{itemize}
-% \item The lithosphere is moving above a fixed mantle where plume originate \citep[e.g.][]{morgan1972deep}
-% \item Age progression along seamounts \citep[see for example][]{crough1980kimberlites,morgan1983hotspot,duncan1984age,sleep1990monteregian}
-% \item Concordance between kimberlite ages and the continental drifting rate of North America. \citep{crough1980kimberlites,duncan1984age}
-% \item Evidence of hotspot related uplift along both continental and oceanic structures. \citep{crough1981mesozoic,sleep1990monteregian}
-% \item Similarities in initial Sr and Nd isotopic ratios between land and sea portions. (Foland et al., 1988)
-% \end{itemize}
 
-% Continuity: 
-% Chercher anomalie de flux thermique (Jaupart 2000 2002 2014) ne montrent rien
-% Parler de l'hypothese rifting of the continental plate qui est en opposition avec Mantle plume.
-% -- Lien avec Vs (effet thermique/chimique/fabrics) ou il faudrait prendre en compte les effets anharmonicity/anelasticity. Voir Ebinger et Sleep 1998
-% -- Haydar voit une anomalie negative de Q dans la region Great Meteor + Finger!! 
-% -- Trouver model de conductivite, anomalie de Gravi? Pour appuyer 
-% -- Faire le lien avec Kaban 2015 Byriol 2016 Poritt 2014
-% -- Fingers From SEMum2 se dirigent plus vers le Bermuda hotspot que GMT, Meme si il y a un lien entre les 2 a 150km.
-% -- Aucune update du Model de GMT depuis Haeman 2000...
-% This feature, parrallel the Absolute Plate Motion (APM) corresponds to the Low Velocity Fingers (LVF) introduced by \cite{french2013waveform} and it shows highest negative anomalies between $100km$ and $200km$ and seems to vanish below $250km$.
-While in \cite{french2013waveform}, the North Atlantic LVF was stopped at the shelf of the US, our model shows that the slow velocity anomaly goes further through the continental lithosphere with a contrast of $-2\%$ to $-3\%$ compared to the surrounding cratonic root.
-This slow anomaly corresponds to the V-shape dent, or divot, observed by \cite{lee1997upper}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% SECTION FOR FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Figure}
 
-% Some authors [Sykes,1978; Croughet al., 1980] have suggested a genetic relation between
-% kimberlites and hotspot tracks, here defined as lin-
-% ear, anorogenicvolcanic chains mapped over oceanic
-% and young continental lithospheres. The termination
-% of these features over thickening continental roots has been cited as evidence supporting the origin of hotspot tracks in the interaction of fixed mantle plumes with an overriding plate [Morgan,1972;Burke and Wilson,
-% 1976; Crough et al., 1980; Sleep, 1990a; Duncan and
-% Richards, 1991], although compelling arguments may also be made for association of hotspot tracks with an-
-% cient rifts and subvertical lithospheric discontinuities [Curtie,1976;Sykes,1978;Anderson1,998
+	% Geological Map
+	\begin{figure}
 
-% Hotspot tracks preserve records of past plate motion and provide markers with which the relative motion between a plate’s surface and underlying mantle regions may be examined
+		\centering
+		\centerline{\includegraphics[viewport=180 340 480 570,clip=true,width=\linewidth]{figures/Geologicalmap_Clouzet.pdf}}
 
-% Quand tu vas parler de l'implication thermale et chimqieu il faut citer le boulot de Haydar avec son anomalue negative de Q.
+		% bb=llx lly urx ury :
+		\caption{\baselineskip 18pt
+		Large scale Precambrian architecture of North America (modified from \cite{whitmeyer2007tectonic}.
+Colours indicate ages of different regions, as shown in the legend.
+The stable cratonic core is defined by the Rocky Mountain Front to the west (dashed dark grey line), defined by the surface trace of deformation), and the appalachian fron to the east (solid light red line).
+The solid blue line demarcates the westward extent of the Grenville/Llano Front.
+The solid red line along the Cordillera province demarcates the breakup of Rodinia.
+Solid green lines indicate tectonic plate boundaries.
+Abreviations are as follows: RW, re-worked; GBA, Great Bear Arc; BHT, Buffalo Head Terrane; Taltson, Talston Arc; Rimbey, Rimbey Arc; Tornagat, Torngat Arc; L.B.A, Little Belt Arc; LRA; THO, Trans-Hudson Orogen; RGR, Rio Grande Rift; Reelfoot, Reelfoor Rift; Kwn.R, Kewanowan Rift; Sask, Sask Craton; WY, Wyoming Province; Med.
+Hat., Medicine Hat Block; Grt.
+Rhy, Granite–Rhyolite Province; Grenville, Grenville Province.
+		}
 
-% At least two Mesozoic hotspot tracks have been postulated for eastern North America; the Juras- sic Verde and Cretaceous Great Meteor tracks [38^40,42,43].
-According to Morgan [39], these two parallel and closely spaced hotspot tracks passed through Ontario and the New England region at slightly di¡erent times (c.
-40 Myr apart).
-The location of the continental portion of these tracks is based on several lines of evidence includ- ing: (1) the current position of Atlantic Ocean hotspots; (2) the record of ocean £oor magnetic anomalies; (3) the assumption that the position of hotspots remains relatively ¢xed on Earth with time; and (4) the relative motions of North Amer- ica, South America, and Africa during the various stages recorded in opening of the Atlantic Ocean since about 200 Myr [50].
-In the case of these Atlantic hotspots, the initial portion of the track typically begins on the continent and eventually moves toward the current oceanic position of the corresponding mantle plume.
+		\label{mapgeol}
 
-% \subsubsection{Gulf of Mexico}
-% Fast anomalies are observed beneath the whole gulf of Mexico down to $150km$ deep.
-Despite the proximity with pan-african lithosphere, such high velocity anomalies represent older than $150Ma$ oceanic lithosphere.
+	\end{figure}
 
- % \subsection{Previous models}
- % Since \cite{grand1984upper}, one well established feature within NA lithosphere is the contrast between the high velocity craton and the low velocity western active part.
-This feature is observed in many global models (e.g., \cite{megnin2000three}, \cite{shapiro2002monte},  \cite{panning2006three}, \cite{houser2008shear}  \cite{nettles2008radially}, \cite{kustowski2008anisotropic}, \cite{ritsema2010s40rts}, \cite{simmons2010gypsum}, \cite{lekic2011inferring}, \cite{debayle2012global}, \cite{schaeffer2013global}, \cite{french2014whole}) where the cratonic lithosphere is caracterized by higher than average (ca.
-$4.8 km.s^{-1}$) velocities down to at least $200km$ deep.
+	% Map of events and stations
+	\begin{figure}
+		\centering
+		\begin{tabular}{ccc}
+			\includegraphics[width=1\textwidth]{figures/map_events_stn_global.png}
+		\end{tabular}
 
- % Measurements of heat flux and heat production over North America ( e.g \cite{jaupart2014building} ) showed low heat flux (i.e $30-50 mW.m^{-2}$) compared to surrounding area, with a contribution at the base of the archean lithosphere brought by convection and a small internal contribution from heat producing elements (e.g \cite{michaut2007transient}).
-This indicates that the cratonic lithosphere that have cooled down by conduction through time can reach a $400km$ thickness and is cold (\cite{rudnick1998thermal} \cite{s2005archean}).
+		\caption{\baselineskip 18pt 
+		Source and station distribution for the new North American inversion.
+Black triangles show the seismic stations.
+Stars are 155 regional events, in addition to 122 teleseismic events.
+Orange stars indiscate events shallower than $30km$ deep, Yellow stars for deeper than $300km$ and red one are in between.
+Thick red line indicates the boundaries used for the RegSEM forward simulation, which extends 89◦ by 89◦ horizontally and down to 1600 km and contains all stations and regional events.}
 
-% Preservation of the keel through time:
-% - depth 400-500km incompatible with sismo and thermobarometric models.
-% - Diamond inclusions in kimberlites and Re–Os isotope analyses of peridotite xenoliths indicate that xenoliths with Late-Archean ages erupted from depths up to 200km.
-% 	-  If we assume that these xenoliths represent samples of the cratonic keel, this suggests that cratonic keels formed soon after (or along with) the craton and that subcratonic lithosphere was cold as early as the Late-Archean
-% 		- Thus, cratonic keels would have appeared to have formed soon after the craton formed and have moved with the craton, relatively undeformed, through time. 
-% 		- However, recent evidence suggests that Re–Os ages may over-estimate the age of xenoliths because zircon grown, and presumably diamond, does not reset the Re–Os system.
-It is difficult, at best, to preserve the diamond stability conditions for 3 Gyr, even in a stable continental keel.
-Perhaps the apparent ages are over-estimates; this would eliminate the need to maintain stable, low temperature keels for time periods up to 3 Gyr in the dynamical modeling.
+		\label{map_evt_stn}
+	\end{figure}
 
-% Jordan's tectosphere
-% root of the cratonic lithosphere have remained unmixed with the rest of the upper mantle.
-% 	Problem: No positive geoid anomaly observed so no mass excess
-% Roots composition must be different in term of composition and iron depleted
-% Chemistry of the tectosphere is mechanically strong and buoyant as the submantle but chemically distinct
-% The effect of composition offsets the increase in density that results from cooling, leading to the isopycnic hypothesis, cratons and keels are neutrally buoyant.
-% 	Problem: Neutral buyancy is not sufficiant to stabilyse cratonic keels from the shearing of mantle flow.
-% 			Thermal conduction is a slow process (for the cooling), it is unlikely that compositional process can balance thermal cooling through time that would lead to neutral buoyancy.
-% 			Geodynamical models cant maintain a stable cratonic root through time.
+	 % Density coverage
+	\begin{figure}
+		\begin{minipage}{0.47\linewidth}
+			\centerline{\includegraphics[viewport=120 10 470 420,clip=,width=\linewidth]{figures/Coverage_L_iter3.png}}
+		\end{minipage}
+		\hfill
+		\begin{minipage}{0.47\linewidth}
+			\centerline{\includegraphics[viewport=120 10 470 420,clip=,width=\linewidth]{figures/Coverage_T_iter3.png}}
+		\end{minipage}
 
-% geochemistry measures
-% 	Cold and dehydrated keels appear to be residual from varying degrees of partial melting over a range of pressure {}with some post depletion metasomatic activity (King 2005).
+		\begin{minipage}{0.47\linewidth}
+			\centerline{\includegraphics[viewport=120 10 470 420,clip=,width=\linewidth]{figures/Coverage_Z_iter3.png}}
+		\end{minipage}
+		\hfill
+		\begin{minipage}{0.47\linewidth}
+			\centerline{\includegraphics[viewport=120 10 470 420,clip=,width=\linewidth]{figures/Coverage_all_iter3.png}}
+		\end{minipage}
 
-% 	Thermobarometric studies indicate that until ca. $175km$ the mantle is chemically depleted. \cite{kopylova1999petrology}
+		\caption{Density coverage for each component and all gathered.} 
+		\label{density_coverage}
 
-% 	Beneath 175km peridotite more fertile: Top of the convective mantle or archean oceanic crust undeplating craton or metasomatic activities...
-Re-Os datattion on eclogites showed that they are older than some peridotites -> keels are subducted archean ocean crust.
-\cite{shirey2001archean}
+	\end{figure}
 
-% 	A number of petrological studies confirm the relative buoyancy of xenoliths thought to represent the composition of the keel [4,24,51–53].
-Reconciling the petrological models of root composition with the seismic velocities remains elusive (e.g., [20,24]).
-Dehydration can lead to several order of magnitude increase in viscosity, relative to mantle at the same pressure and temperature conditions [54,55].
-Hence if the cratonic keels represent the depleted residual from a large melting event, this would be consistent with the buoyant and viscous properties that are needed to maintain stable keels over a long time period.
+	% Box Tomo
+	\begin{figure}
+		\begin{tabular}{ccc}
+			(a)&&(b)\\
+			\includegraphics[width=.4\textwidth]{figures/snap_glob.pdf} && \includegraphics[width=.4\textwidth]{figures/snap_reg.pdf} \\ 
+			\includegraphics[width=.4\textwidth]{figures/seismo.pdf}    && \includegraphics[width=.4\textwidth]{figures/seismo.pdf}
+		\end{tabular}
 
-% Electrical conductovity:
-% The bootom of upper mantle is wetter than above beneath craton and dryer than ocean until 250km \cite{yang2015three} \cite{hirth2000comparison}
-% Jordan [23] hypothesized that subcratonic keels or
-% roots are attached to the cratonic lithosphere and that
-% these roots or keels have remained isolated from and
-% unmixed with the rest of the mantle since the
-% formation of the craton. If the cratonic root is a thick
-% thermal boundary layer and the subcratonic mantle is
-% average upper mantle composition, there should be a
-% 10-m positive geoid anomaly over cratons [29].
-% Because there is no correlation between continents
-% and the geoid [30–32], the subcratonic mantle must be
-% compositionally distinct from the surrounding upper
-% mantle. Jordan called this subcratonic material tectosphere
-% to distinguish it from normal upper mantle. In
-% the tectosphere hypothesis these keels are chemically
-% distinct from the convective mantle and are both
-% compositionally buoyant and mechanically strong.
-% The effect of composition offsets the increase in
-% density that results from cooling, leading to the
-% disopycnicT hypothesis [33]—cratons and keels are
-% neutrally buoyant. Numerical studies have clearly
-% demonstrated that neutral buoyancy is not sufficient to
-% stabilize cratonic keels from the shearing of mantle
-% flow. Thermal conduction is a slow and continuous
-% process and it seems highly improbable that a
-% compositional process would exactly balance thermal
-% cooling through time, leading to neutral buoyancy at
-% all depths envisioned in the isopycnic hypothesis.
-% While both strength and buoyancy act to stabilize
-% keels in theory, it has been quite challenging to
-% produce geodynamical models that maintain stable
-% cratonic keels in a convecting mantle over several
-% billion years of mantle evolution (e.g., [3] and see
-% later section on craton stability).
+		\caption{\baselineskip 18pt 
+			Comparison between a global scale simulation and a regional scale simulation of wave propagation following an earthquake in south America.
+The seismograms corresponds to the recordings at 37 station located in north America (red spheres).
+			The pink sphere shows the epicenter of the earthquake.
+In (a), the wavefield is modeled globally using the spectral element method Specfem3D\_globe \citep{komatitsch2002spectrala}.
+In (b), the wavefield is modeled regionally using the RegSEM.
+During the inversion, all the synthetic seismograms are computed using the regional SEM code RegSEM \citep{cupillard2012regsem} code and regenerated thanks to virtual sources located around the computational domain \citep[see][]{masson2013numerical}.
+The sismograms in (b) are exactly similar to those in (a), however, the computational effort is greatly reduced in the regional simulation (b).}
 
-% In the steady state thermal conditions that prevail in old continents, the man- tle heat flux, i.e., the heat flux at the Moho, is the sum of the heat flux at the base of the lithosphere, which is brought by mantle convection, and heat generated by heat producing elements in the lithospheric mantle.
-The first contribution originates deeper than 200 km and potential lateral variations would get smoothed out by horizontal diffusion.
-The contribution of heat producing elements in the lithospheric mantle is very small, as shown by xenolith studies [Michaut et al., 2007], and horizontal anomalies would also be smeared out.
-Joint analysis of surface heat flux data and seismic wave velocity variations confirms that Moho heat flux variations are constrained to be less than about ± 3 mW m−2 , so that the total variation is 6 mW m−2 across the province [Perry et al., 2006b; Lévy and Jaupart, 2011].
+		\label{injection}
 
- % \subsection{Craton caracteristic}
+	\end{figure}
 
- % Through times, Lauretia craton remained undeformed while being part of the earliest supercontinent Nuna-Columbia \citep{meert2012whats}, then Rodinia and finally Pangaea.
+	 % 3d map of Vs
+	\begin{figure}
+		\centering
+		\includegraphics[width=1\textwidth]{figures/Vs-60-300-for-paper.png}
 
-% Archean part more buoyant than proterozoic one
-% The archean province may reflect a relatively depleted underlying mantle lithosphere.
-% The presume secular decline in mean temperature of the asthenosphere impliesa greater depth and volume of melting accompanying mantle upwelling in the Archean, and consequently thicker, more depleted mantle lithosphere.
+		\caption{\baselineskip 18pt
+		3D isotropic shear wave velocity structure of the continent.
+Map views are shown from 60 km down to 300 km, as variations with respect to the regional mean (thick red line in Figure ~\ref{1daverage}) as $\frac{dV_S}{V_{S_{O}}}$ .
+		Below each map, depth and its regional shear velocity mean is indicated. Minimum and maximum perturbations are also indicated.
+		}
 
-% Most importantly, the 2.7 Ga age of the Greenland eclogites supports the observation by Shirey and Richardson (2011) that mantle eclogite materials older than 3.0 Ga appear absent from the cratonic record.
-They suggest that the appearance of cratonic eclogites in most cratons after 3.0 Ga marks the onset of the Wilson Cycle of plate tectonics, when more frequent subduction and collision events became the most prominent continental growth mechanism.
-(Tappe et al., 2011)
+		\label{3d-VS}
 
-% Previous studies
-% Link with upper mantle tomography
-% layering and radial intro to aximuthal
+	\end{figure}
 
-% In the early 1960s, theoretical work (e.g.
-Anderson 1961) on the dispersive properties of transversely isotropic media suggested that radial anisotropy in the upper mantle could have a pronounced ef- fect on the shape of surface wave dispersion curves and could lead to an apparent discrepancy between Love and Rayleigh wave data.
+	% 3d map of Xi
+	\begin{figure}
+		\centering
+		\includegraphics[width=1\textwidth]{figures/Xi-60-300-for-paper.png}
 
-% For  more  than  thirty  years,  seismologists  have  imaged  the  earth's  interior  using  seismic
-% waves generated by earthquakes,  and traveling through di erent structures of the planet.
-% A  remaining  challenge  in  seismology  is  to  interpret  the  recovered  Earth  models  in  terms
-% of physical properties (e.g.  temperature, density, mineral composition) that are needed for
-% understanding  mantle  dynamics  and  plate-tectonics.   For  example,  a  region  of  slow  wave
-% speed can be either interpreted as anomalously warm, or rich in water, or iron.
-% Although seismic waves are sensitive to a large number of visco-elastic parameters as well
-% as density, the mantle models constructed from seismic tomography are only parameterized
-% with a few physical parameters, for example average isotropic shear wave velocity and radial
-% anisotropy (e.g. French et al., 2013). This is because given the available information observed
-% at the surface, there is not enough resolution to entirely describe the local elastic tensor.  In
-% addition to the limitednumber of resolvable elastic (and anelastic) parameters, there is also
-% the question of spatial resolution, namely the smallest spatial scale at which heterogeneities
-% can be imaged.
-% The  number  of  independent  elastic  parameters  that  can
+		\caption{\baselineskip 18pt
+		3D radial anisotropic structure of the continent.
+Map views are shown from 60 km down to 300 km, as variations with respect to the regional mean (thick red line in Figure ~\ref{1daverage}) as $\frac{d\xi}{\xi_{O}}$ .
+		Below each map, depth and its regional shear velocity mean is indicated. Minimum and maximum perturbations are also indicated.
+		}
 
-% \section{Methodology}
-% % \subsection{Kernels}
-% In the continuity of previous work (e.g Marone et al.,2007; Yuan et al.,2011; Yuan et al., 2014) we invert 3-component fundamental and overtone surface waves in the time domain to image North American upper mantle 3D structure using an hybrid inversion scheme.
-\\
-% Inferring the elastic structure of the earth from the analysis of propagating waveforms is a well-studied inverse problem in geophysics. 
-% % A non linear inverse problem that can be solved iteratively, where the accuracy of the different solution's terms may come along with the expensiveness of their calculation. 
-% Our inverse problem is solved iteratively following the generalized least-squares formalism of Tarantola and Valette, 1982. We define a discrete L2 misfit functional:
+		\label{3d-Xi}
 
-% \begin{equation}
-% 2\Phi(m_{k}) = [d - g(m_{k})]^{T} C^{-1}_{d} [d-g(m_{k})] + [m_{p}-m_{k}]^{T} C_{m}^{-1} [m_{p}-m_{k}]
-% \end{equation}
+	\end{figure}
 
-% where \ensuremath{m_{k}} represents the \ensuremath{k^{th}} iterative model estimate, \ensuremath{d} and \ensuremath{g(m_{k})} are the observed data and data predictions (time-discretized waveforms and period-discretized group velocities), \ensuremath{m_{p}} is the model prior (1D earth or 3D starting model), and \ensuremath{C_{m}} and \ensuremath{C_{d}} represent a priori model and data covariance operators respectively.
+	% 1D profiles
+	\begin{figure}
+		\centering
+		\includegraphics[width=1\textwidth]{figures/1D_profile_with_regions.png}
 
-% We solve for successive updates to the current model estimate \ensuremath{m_{k} \to m_{k+1}} following the update equation that minimizes $\Phi$, namely:
-% \begin{equation}
-% m_{k+1} = m_{k} + (C_{m}G^{T}_{k}C^{-1}_{d}G_{k} + I)^{-1} (C_{m}G_{k}^{T}C^{-1}[d-g(m_{k})]+ m_{p}-m_{k})
-% \end{equation}
-% We employ a hybrid waveform inversion scheme, in which “exact” forward modeling of the wavefield (i.e.
-the waveform elements of $g(m_{k})$ is achieved using the spectral element method (SEM: e.g.
-Komatitsch and Vilotte, 1998; Komatitsch and Tromp, 2002a,b), while approximate sensitivity kernels (i.e.
-elements of G) are calculated using non-linear asymptotic coupling theory (NACT: Li and Romanowicz, 1995).
+		\caption{\baselineskip 18pt 
+		1D average depth profiles for isotropic shear velocity (left) and radial anisotropy (right), for the SEMucb\_wm1 global model (thick black), the continental average (thick red); and depth profiles for different regions in the model (see legend and Figure~\ref{mapgeol}).
+The continental average is always higher than the global average in terms of $V_S$, while for $\xi$, they are similar deeper than $250 \: km$.}
+		\label{1daverage}
+	\end{figure}
 
-% NACT is a normal-mode perturbation approach, which takes into account coupling between modes both along and across dispersion branches, allowing the computation of 2-D broad-band sensitivity kernels which more rigorously reproduce the sensitivity of body waveforms to structure along the ray geometrical path in the vertical plane containing the source and the receiver.
-% However, when the distance between a source and a station is below $15^{o}$, NACT breaks down and we instead use the more standard Path Average Approximation (PAVA).
-Developed by Woodhouse \& Dziewonski, 1984; PAVA kernels are sensitive to the average structure between the source and the receiver and along the great circle path, that is, the corresponding sensitivity kernels are ‘1-D’ in the vertical plane containing the source and the receiver.
-This approximation is best for single-mode seismograms, such as fundamental mode surface waves (Romanowicz, 2008).
+	% Xsections of the craton
+	\begin{figure}
+		\begin{minipage}{0.5\linewidth}
+			\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_A.pdf}}
+		\end{minipage}
+		\hfill
+		\begin{minipage}{0.5\linewidth}
+			\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_B.pdf}}
+		\end{minipage}
 
-% PAVA/NACT do not consider off great circle path sensitivity in the horizontal direction, which gives rise 
-% to focusing and defocusing effects (e.g. Zhou et al. 2005). These effects are important for amplitude 
-% fitting. Here, we are, however, primarily concerned with fitting the phase and for that, the 2-D effects in
-% the vertical plane are dominant, especially for overtones. The dominance of 2-D effects in the
-% vertical plane can be assessed by considering an asymptotic expansion of normal mode first order
-% perturbation theory, which shows that 2-D coupling effects in the vertical plane are of zeroth 
-% order, while focusing effects are of higher order (e.g. Romanowicz, 1987; Li \& Romanowicz, 1995; Romanowicz, 2008).
+		\caption{Cross Section of the lithospheric core}
+		\label{cratoncross}
 
-% On the other hand, we use the spectral element method to accurately compute the forward seismic wavefield $g(m_{k})$. 
-% Lekić \& Romanowicz., 2011 showed, that using approximate methods, such as normal mode perturbations theory, to calculate synthetic seismograms can introduce significant errors in the modeling of the wavefield in a heterogeneous 3-D earth.
-These errors can be seen as additional noise in the inversion and may introduce some model dependent errors.
-In that situation, convergence of the inverse problem can be compromised.
-We use a modified version of the spectral element code RegSEM (Cupillard et al., 2012) which initially was designed to calculate synthetic seismograms in a $90^{o} \times 90^{o}$ area within which both receivers and event sources are located.
-RegSEM takes into account effects of the oceans, topography / bathymetry, ellipticity, gravity, rotation and anelasticity.
-It also includes Perfectly Matched Layers (PML) absorbing boundary conditions, which allow us to restrict the computation laterally and radially to the region of interest with the vanishing of spurious reflections at the borders.
-(Patera \& Vilotte, 2005)
+	\end{figure}
 
-% Using only regional events is an opportunity to image the North American upper mantle without the influence on the wavefield of the global 3-D structure.
-However the current data coverage from a pure regional data set is not sufficient to constrain our region.
-To address this problem, we chose to include teleseismic events and developed a new method in order to calculate teleseismic synthetics where, through iterations, the wavefield is calculated only in the region of interest by the use of secondary sources.
-% Developed by Masson et al., 2013, the idea is to construct an equivalent of time-reversal mirrors (Fink et al., 1989) at the boundary of our region of interest where the wavefield is then re-generated forwardly.
-% Time reversal is a two step process.
-First, waves propagating in a medium from a source are recorded thanks to an array of transducers (also named time-reversal mirrors).
-Second, after having reversed in time our records, transducers emit back into the medium toward the source where the ernergy is refocused in time and space.
-The feasability of such procedure relies in the reciprocity in space and reversibility in time of the wave equation.
-(Robertson \& Chapman, 2000)
+	\begin{figure}
+		\begin{minipage}{0.5\linewidth}
+			\centerline{\includegraphics[width=\linewidth]{figures/profiles_WE-50.pdf}}
+		\end{minipage}
+		\hfill
+		\begin{minipage}{0.5\linewidth}
+			\centerline{\includegraphics[width=\linewidth]{figures/profiles_SN-100.pdf}}
+		\end{minipage}
 
-% % As in Robertson \& Chapman, 2000 we use a two step method.
-The first step performs wave propagations in the global medium, then as a second step the wavefield is used as an input (time-reversal mirrors) to obtain the wavefield into the medium under study.
+		\caption{Cross Section of the lithospheric core showing the core retraction}
+		\label{cratoncrossedge}
 
-% In the elastic case, the ideal response of a time-reversal mirror can be expressed in the frequency domain with the use of the Helmotz-Kirchhoff representation theorem:
-% \begin{equation}\label{eq:repres}
-% \begin{split}
-% 	u_{i}^{*}(\bm x) =&  \int_{V} G_{in}(\bm x, \bm x')f_{n}^{*}(\bm x')dV' \\
-% 	 +& \oint_{S} G_{in}(\bm x, \bm x') n_{j} C_{njkl} \partial_{k}^{'} u_{l}^{*}(\bm x') dS' \\
-% 	 -& \oint_{S} u_{n}^{*}(\bm x') n_{j} C_{njkl} \partial_{k}^{'} G_{il}(\bm x, \bm x')dS'
-% \end{split}
-% \end{equation}
+	\end{figure}
 
-% where $u$ is the displacement vector, $G_{in}(\bm x, \bm x')$ is the Green's tensor denoting the displacement at location $\bm x$ in the $i$ direction due to a unit point source at $\bm x'$ in the $n$ direction, $C_{ijkl}(\bm x)$ is the stiffness tensor at the $\bm x$ location, $\partial_{k}^{'}$ stands the derivative with respect to the $x_{k}^{'}$ coordinate and the star symbol (*) stands for the complex conjugate.
-Eq.\ref{eq:repres} shows that, if the exciting force $f_{n}(\bm x')$ is known throughout the volume $V$ and that the wavefield $u_{i}(\bm x)$, the associated traction $n_{j} C_{njkl} \partial_{k}^{'} u_{l}(\bm x')$ are known on the surface $S$; we are then able to regenerate the wavefield anywhere within the volume $V$.
-(Masson et al., 2013)
+	% Xsection of the GMT track
+	\begin{figure}
+		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_GMTrack.pdf}}
 
-% Masson et al., 2013 proposed to introduce a spatial window function into the wave equation (original wavefield) in order to determine the exciting forces, that applied to the surface $S$, will generate the response $u_{i}(\bm x)$ defined in equation \ref{eq:repres}.
+		\caption{Cross Section of the lithosphere along the Great Meteor track based on \cite{heaman2000timing} construction}
+		\label{gmtcross}
 
-% Starting with the wave equation in the frequency domain:
+	\end{figure}
 
-% \begin{equation}\label{eq:wavefreq}
-% 	\rho \omega^{2} \bm u_n + \partial_j (C_{njkl}\partial_k \bm u_l) = -f_n,
-% \end{equation}
+	% dIso for Xi
+	\begin{figure}
+		\centerline{\includegraphics[width=\linewidth]{figures/Xi_NASEM-iter3-Xi-iso_100km.png}}
 
-% the generated wavefield in $M \in V$, is defined as:
+		\caption{3D radial anisotropic structure of the continent at 100km compared to the Isotroic case. 
+		Map view shows variations with respect to the Isotropic case as $\frac{\xi - 1}{1}$ in \%. 
+		Map view is saturated to better show where $V_{SV}$ is stronger than $V_{SH}$.
+		Below each map, depth and its regional shear velocity mean is indicated. Minimum and maximum perturbations are also indicated.}
+		\label{3d-xi-iso-100km}
 
-% \begin{equation}\label{eq:um}
-% 	\bm u^M (\bm x ,t) = \bm u(\bm x ,t)\omega(\bm x).
-% \end{equation}
-
-% with the use of a bi-valued function $\omega$:
-
-% \begin{equation}\label{eq:window}
-% 	 \omega(\bm x) = \left\{\begin{array}{ll}
-% 						1 	&\mbox{ for all \quad $\bm x \in M$} \\
-% 						0   &\mbox{ for all \quad $\bm x \notin M$}
-% 						\end{array}
-% 						\right.
-% \end{equation}
-
-% such as:
-
-% \begin{equation}\label{eq:um_in_u}
-%  \bm u^M (\bm x ,t) = \left\{\begin{array}{cl}
-% 						\bm u (\bm x ,t) &\mbox{ for all \quad $\bm x \in M$} \\
-% 						0                &\mbox{ for all \quad $\bm x \notin M$}
-% 						\end{array}
-% 						\right.
-% \end{equation}
-
-% where $M$ is the regional subvolume bounded by $\delta M$ within the global volume $V$.
-
-% The mirror's excitation force that generates $\bm u^M$ is introduced as:
-% \begin{equation}\label{eq:exc_force_reg}
-%  	f_n^M = -\rho \omega^2 \bm u_n^M - \partial_j (C_{njkl}\partial_k\bm u_l^M)
-%  \end{equation}
-
-%  Using eq. \ref{eq:wavefreq} - \ref{eq:exc_force_reg} and after developing:
-%  \begin{equation}
-%    	f_n^M = \omega f_n - C_{njkl} \partial_{j}\omega \partial_k u_l - \partial_j(u_l C_{njkl} \partial_k \omega).
-%  \end{equation}
-
-%  Finally, convolving eq. \ref{eq:exc_force_reg} with the Green's function of the medium and then integrate over the volume (see Masson et al., 2013 for a detailed derivation) gives :
-
-%  \begin{equation}\label{eq:umfinal}
-%  \begin{split}
-%   	u_{i}^{M}(\bm x) =& \int_{M} G_{in}(\bm x , \bm x{'})f_{n} (\bm x') dV' \\
-%   					 +& \oint_{\delta M} G_{in} (\bm x, \bm x') n_{j} C_{njkl} \partial_{k}^{'} u_l(\bm x')dS' \\
-%   					 -& \oint_{\delta M} u_{n}  (\bm x') n_{j}   C_{njkl} \partial_{k}^{'} G_{il}(\bm x, \bm x')dS'
-%  \end{split}
-%  \end{equation}
-
-% In practice, one global simulation per event is needed to generate such mirrors, they are then used as secondary sources to generate the wavefield in our region of interest.
-These secondary sources are calculated once and for all.
-The main advantage of using such a method lies in the fact that we update the elastic structure of the region iteratively, we only need to calculate synthetics in the region.
-Since we're not updating the global structure (following the zeroth iteration ) of the earth but only North America, this allow us to save significant computation time and therefore attain shorter periods.
-
-% The hybrid inversion scheme represents a compromise between speed and accuracy.
-The use of exact numerical methods to calculate the wavefield remains expensive, while waveform partial derivatives computed using NACT are approximate and inexpensive.
-Despite the approximate nature of NACT, we point out two important properties: sensitivity kernels, that are recalculated at each iteration in the current 3-D model) accommodate non-linearity due to multiple forward scattering (Romanowicz et al., 2008); and it includes finite-frequency effects in the 2D great-circle plane, critical to modeling surface-wave overtones and long-period body waves (Li and Romanowicz, 1995).
-
-% % \subsection{Effect of the crust}
-% % From Scott's thesis 
-% % Because SEM simulations are so precise, we require a model of crustal structure that accurately predicts the seismic response of the earth’s crust – particularly for highly sensitive fundamental- mode surface waves.
-Indeed, fidelity to crustal effects is especially critical for accurate retrieval of anisotropic mantle structure, as often noted when discussing the shortcomings of standard linear crustal corrections (Bozdag and Trampert, 2008; Lekic ́ et al., 2010; Ferreira et al., 2010).
-Here, we adopt the smooth crustal mod- eling technique proposed in Lekic ́ and Romanowicz (2011) and later refined by French et al.
-(2013).
-The thin layers of the earth’s crust slow down SEM computations considerably, particularly in the oceans.
-To mitigate this effect, we construct models of the crust that contain no interior discontinuities and are potentially thicker than the real crust in some regions.
-These smooth mod- els are calibrated to mimic the seismic response of the earth’s crust, as seen through observations of surface-wave group-velocity dispersion, by introducing crustal radial anisotropy (Backus, 1962; Capdeville and Marigo, 2007).
-% % In the continents, our smooth crust matches the Moho depth of Crust2.0 (Bassin et al., 2000), with slight smoothing to avoid aliasing in the SEM mesh, while Moho depth saturates at 30 km in oceanic regions.
-A similar technique, but calibrated to mimic the dispersion of an a priori crustal model and not restricted to match continental Moho topography, has previously been shown to be effective for regional-scale waveform inversions (Fichtner et al., 2009).
-The disadvantages of such a technique are: (a) that the crustal model is only valid over the period range it is calibrated, in this case periods at or above 25 s, and (b) susceptibility to noise in the dispersion data upon which it is calibrated.
-The first concern should not be an issue for the 40 s and 60 s SEM modeling em- ployed here, confirmed in the global and North-American region validation simulations reported by French et al.
-(2013).
-To address the second concern, these authors produced ensembles of crustal models calibrated on dispersion data with realistic-amplitude syn- thetic noise added, and examined variation in the resulting SEM synthetics, concluding that such errors could not significantly af- fect long-period waveforms.
-Thus, we are confident that the crustal modeling approach is able to attain SEM speedup, while maintain- ing fidelity to the observed surface-wave dispersion characterizing crustal structure, and without negatively affecting the resulting long-period fundamental and higher mode waveforms and the re- sulting upper mantle shear velocity and radial anisotropy structure.
-
-% \subsection{Parameterization}
-% Propagation of seismic waves through an arbitrary Hookean medium depends on 21 parameters of the stiffness tensor,
-% and inferring the values of all these parameters at all locations within the mantle is not feasible with available seismic data. 
-% However, by approximating the Earth as a transversely isotropic medium with a vertical axis of symmetry ( VTI ), we can reduce the number of free parameters,
-% while capturing the first order observation that horizontally polarized surface waves travel, on average, faster than vertically polarized ones (e.g. Anderson 1961; McEvilly 1964).
-
-% A VTI medium is described by density \ensuremath{\rho} and the five Love coefficients (\ensuremath{A}, \ensuremath{C}, \ensuremath{F},
-%  \ensuremath{L} and \ensuremath{N} ) (Love 1927). 
-% We parametrize our model in terms of \ensuremath{\rho}, the Voigt average isotropic shear ($V_S$) and compressional ($V_P$) velocities: \\
-% \begin{align}
-% V_{S} =& \sqrt{\frac{2V^{2}_{SV}+V^{2}_{SH}} {3}} \\
-% V_{P} =& \sqrt{\frac{V^{2}_{PV} + 4V^{2}_{PH}}{5}}
-% \end{align}
-
-%  and the anisotropic parameters:
-
-%  \begin{equation}
-%  \xi = \frac{N}{L}, \quad \varphi = \frac{C}{A}, \quad \eta = \frac{F}{A-2L} 
-%  \end{equation}
-
-% Love and Rayleigh waves are primarily sensitive to shear wave structure, whereas their sensitivities are significantly weaker for other elastic parameters (Dahlen \& Tromp 1998). 
-% Thus, we decrease the number of parameters of interest by choosing not to invert for lateral variations in the poorly constrained $V_{P}$ $\varphi$, $\rho$ and $\eta$ parameters. 
-% Instead, we parametrize the elastic structure of the mantle in terms of $V_{S}$ and $\xi$.
-% The parameter $\eta$ governs the variation of wave speed at directions intermediate to the horizontal and vertical. 
-% When $\eta$ and $\varphi$ are approximately equal to one, which is very likely the case in the mantle, we can approximately relate Voigt average velocities to those of vertically and horizontally polarized waves.
-% Following Montagner \& Anderson, 1989 we impose an a priori information on other parameters perturbations by scaling them with inverted parameters:
-
-% \begin{equation}
-% \frac{\delta(V_{P})}{\delta(V_{S})} = 0.5, \quad
-% \frac{\delta(\rho)}{\delta(V_{S})} = 0.33, \quad
-% \frac{\delta(\eta)}{\delta(\xi)} = -2.5,   \quad
-% \frac{\delta(\varphi)}{\delta(\xi)} = -1.5,   \quad
-% \end{equation}
-
-% We opt for $V_{S}$ and $\xi$ parameterization instead of the more traditional $V_{SV}$ and $V_{SH}$, to avoid additional uncertainties on the amplitude and possibly even sign of the anisotropic parameter $\xi$ introduced by damping in the inversion process.
-(Marone \& Romanowicz, 2007)
-% Indeed, if $\xi$ is computed after the inversion from the derived $V_{SV}$ and $V_{SH}$, its value can be biased by the different effect of damping on the two velocity parameters.
-( Panning \& Romanowicz; 2006).
-
-% In theory, the retrieved radial anisotropic model could be biased by the presence of azimuthal anisotropy, neglected at the present stage of this study.
-In practice, with sufficient azimuthal coverage, this possible effect is reduced because the azimuthal variations will be averaged out at each point of the model.
-This study, limited to radial anisotropy, should be regarded as a first step towards a more complex model of anisotropy, where the assumption of a vertical symmetry axis will be dropped and which will, therefore, still be characterized by radial anisotropy but with a symmetry axis with an arbitrary orientation (e.g.
-Montagner \& Nataf 1988).\\
-
-% From the Core Mantle Boundary up to the Moho, the model is expressed on 26 cubic splines $\nu_{q}(r)$ defined in Megnin \& Romanowicz (2000), though we invert for structure only in the top 16 splines; deeper structure is fixed to that of our starting model $SEMucb_wm1$ (French \& Romanowicz, 2014).
-% The knot locations are at radii: 5690, 5810, 5910, 5990, 6061, 6101, 6131, 6161, 6191, 6221, 6241, 6261, 6281, 6301, 6321, 6341 km. 
-% Laterally, we parametrize our model spatially in terms of spherical splines $\beta_{p}(\theta, \phi))$ (Wang \& Dahlen 1995). 
-% Thus, the value of a given model parameter $m$ at any location in the Earth $(\theta, \phi, r)$ can then be calculated from a set of spline coefficients $m_{pq}$:
-
-% \begin{equation}
-% m(\theta,\varphi,r) = \sum_{p}\sum_{q}m_{pq}\beta_{p}(\theta,\varphi)\nu_{q}(r)
-% \end{equation}
-
-% The splines are a local basis, and thus help minimize the mapping of structure in one region into structure in distant regions, which can be an undesirable effect of global parametrization such as spherical harmonics (Lekic et al., 2011) .
-% % In general when parameterizing our model, we put strict a priori constraints on the minimum length scale of structure allowed in our model.
-This truncation results in spectral leakage (aliasing) of short scale heterogeneity into longer length scales (Trampert \& Snieder 1996), though the use of splines reduces this aliasing when compared to spherical harmonics or spherical pixels (Chiao \& Kuo 2001).
-To further reduce the aliasing of retrieved structure, we allow structure to vary at shorter length-scales than those that we can reasonably expect to image and interpret (Spetzler \& Trampert 2003).
-
-% In our regional tomographic studies, the mesh for the target region can be refined by adding knots halfway between two neighboring grid points, until the desired knot spacing has been reached, while maintaining a coarser mesh throughout the rest of the globe (Marone \& Romanowicz 2007a).
-This approach allows flexibility: the grid spacing can, for instance, be chosen as a function of the data coverage.
-Therefore, we maximize the resolution within the ideally sampled region, without need for over parameterizing the areas with insufficient data density.
-For the $V_{S}$ parameter we use a regional level 7 spherical grid (~1 degree knots spacing) and level 6 ( ~2 degree ) for $\xi$.
-
-% \subsection{A priori information}
-% When solving an inverse problem, the importance of the a priori information depends greatly on the use one plans for the estimated model parameters.
-% If one simply wants one example of a solution to the problem, the choice of a priori information is unimportant.
-However, if one wants to develop arguments that depend on the uniqueness of the estimates, the validity of the a priori assumptions is of paramount importance.
-(Menke, 2012)
-
-% As a starting model, we opted for the global 3D radially anisotropic shear wave velocity model $SEMUCBwm1$ (French \& Romanowicz, 2014).
-This global whole mantle model was built in the continuity of SEMum generation models (Lekić \& Romanowicz, 2011) with an extended data set of :
-% \begin{itemize}
-% 	\item 3 components 60-400s periods waveforms
-% 	\item 3 components 30-400s periods body-waves 
-% 	\item group-velocity dispersion maps 25-150s 
-% \end{itemize}
-
-% Concerning the isotropic structure, the a priori information on which SEMum is based on is a VTI 1D earth model.
-The idea was to fit long-period ($T > 60 s$) waveforms (Cammarano \& Romanowicz 2007) starting from one of the physical reference models of Cammarano et al.
-(2005), which are calculated from a fixed composition (dry pyrolite) and a thermal profile using the elastic and anelastic properties of principal mantle minerals.
-
-% As for the the anisotropic part, the a priori information of $\xi$ was obtained by testing several values against a set of observed frequencies of spheroidal and toroidal modes, keeping fixed the elastic structure.
-% In this process, $\xi$ was allowed to be stronger or equal to 1 (isotropic case) at depth shallower than 320 km as verified in numerous previous seismic studies (Dziewonski \& Anderson 1981).
-
-% When solving the inverse problem, we also include a priori information on the model in a covariance matrix $C_{M}$.
-It specifies the expected deviation of true mantle structure from that specified by our starting model.
-This matrix is defined by the variance $\sigma^{2}_{0}$ (the diagonal entries) and the horizontal and vertical correlation lengths, $h_{0}$ and $v_{0}$, associated with each spline knot.
-
-%  These correlation lengths are spatially tuned according to data-coverage density and quality.
-As in \cite{lekic2011inferring}, the element of $C_{M}$ jointly describing the state of a priori information on model coefficients i and j is given by:
-
-% \begin{equation}
-% C_{m} = \sigma_{ij} exp(\frac{cos\Delta_{ij}-1}{h^{2}_{ij}}) exp(\frac{-2d^{2}_{ij}}{v^{2}_{ij}})
-% \end{equation}
-
-% where $\Delta_{ij}$ is the minor-arc distance separating model coefficients i and j, $d_{ij}$ their radial separation, and $v_{ij}$ and $h_{ij}$ the averages of their associated radial and lateral correlation lengths.
-
-% In order to increase image resolution, we allowed the horizontal length of correlations to vary between $100 km$ and $300 km$ for $V_{S}$; $300 km$ to $600 km$ for $\xi$.
-Thus, within the model, $h_{ij}$ can vary based upon an estimator for relative sensitivity.
-The idea is to take the diagonal of the approximated Gauss-Newton Hessian $G^{T}C^{-1}G$ associated with the data misfit term, to represent aggregate sensitivity of the data to each model coefficient.
-Contributions from different data type are weighted accordingly to the quality and uncertainty estimates appearing in the data covariance matrix $C_{D}$.
-These values are, in turn, used to select local correlation-length estimates by scaling to the interval between prescribed minimum and maximum values for a given model parameter.
-
-% \subsection{data}
-
-% In this study, long-period waveforms and group velocity dispersion maps are used to constrain shear wave velocities and radial anisotropy in both crust and upper mantle of North America.
-
-% The group-velocity dispersion data are from Ritzwoller et al., 2002 and are in the form of $1^o \times 1^o$ maps between 25 and 150s.
-The use of short period dispersion curves ( below 80s ) constrained our homogenized crustal structure, while the entire period range was included during the inversion of the mantle structure.
-
-% The long-period waveform data set comprises Rayleigh and Love accelerograms including both fundamental and higher modes recorded on 3 components broad-band seismographs. 
-% Fundamental modes are primarly sensitive to the first 300km while overtones have sensitivity into the transition zone. We used 2 different data sets:
-
-% \begin{itemize}
-% 	\item a regional data set consisiting of 155 events $(4.5 < M_{w} < 6.0)$ where sources are located within the $90^o \times 90^o $RegSEM mesh.
-% 	\item a teleseismic data set consisting of 122 events $(5.5 < M_{w} < 6.9)$ for which secondary sources are calculating around the $90^o \times 90^o $ RegSEM mesh using the procedure of \cite{masson2013numerical}.
-% \end{itemize}
-
-% Such data set, recorded at more than 2400 stations, provides an exellent azimuthal coverage of NA continent, with a total of 19.991 Love wavepackets, 55.176 Rayleigh wavepackets for fundamental modes, 82.809 overtones with 20 per cent from the T component and 63.179 wavepackets mixing fundamental and overtone energies.
-
-% We filter accelerograms between 40 and 400s with corner frequencies at 53 and 250s, and separate them into wave-packets.
-Before inversio, the wavepackets are weighted according to their relative amplitudess, in order to avoid dominance by the larger amplitude fundamental modes in the inversion.
-In our weighting scheme, we also account for background noise, and path redundancy, as described in \cite{li1996global} This weighting step is crucial as it homogenizes the data coverage within the region and provides a way to construct the data covariance matrix $C_D$, which we approximate as a diagonal matrix.
-
+	\end{figure}
 \bibliographystyle{apalike}

### 2017-03-03 Paper_NA_br.tex
Update 03.03.17

@@
 		The average of $\xi$ is higher than $1$ (which means that $V_{SH} > V_{SV}$), but lower than the global average, the latter influenced by the strong $\xi > 1$ oceanic signature. 
+		Both trends are similar with depth, but the maximum in $\xi$ is slightly deeper than in the global average and the change from $V_{SH} > V_{SV}$ to $V_{SH} < V_{SV}$ is reached at $\sim 420 \: km$. 
+		At shallower depths than $\sim 200 - 220 \: km$, strongest variations relative the continental average are observed ($+/- 4 \: \text{to} \: 8 \%$, compared to $+/- 1 \: \text{to} \: 3 \%$). 
+		Alternations of positive and negative gradient are observed from $50 \: \text{to} \: 400 \: km$ with maxima diminishing along depth.
+Below than $420 \: km$, negative to isotropic structures with very small lateral variations (i.e $ dln(\xi) < 0.5 \%$) are observed.
+
+		% A negative gradient is regionally observed with its minima at $100 \Km$ of $\sim \xi = 1.02$ evolving to a positive gradient with a maximum of $\xi = 1.03$ located at $150 \: km$. 
 		% where shearing is more important.
-		Both trends are similar with depth, but the maximum in $\xi$ is slightly deeper than in the global average and the change from $V_{SH} > V_{SV}$ to $V_{SH} < V_{SV}$ is reached at ca. $420km$.
 
@@
 
-		Our study confirms the presence of a thick lithospheric root in the craton with faster than average $V_S$ down to $250km$ depth, with highest velocities, ca.
-$4.8km.s^{-1}$, located at depths shallower than  $150km$ (e.g.
+		Our study confirms the presence of a thick lithospheric root in the craton with faster than average $V_S$ down to $250 \: km$ depth, with highest velocities, ca.
+$4.7 \: km.s^{-1}$, located at depths shallower than  $150 \: km$ (e.g.
 Figure~\ref{cratoncross}).
-		Laterally, the cratonic root in $V_S$ coincides roughly with precambrian aged crust (see Figure~\ref{3d-VS} from 60 to 150km deep)
+		Laterally, the cratonic root in $V_S$ coincides roughly with precambrian aged crust (see Figure~\ref{3d-VS} from 60 to 150km deep).
 		The western edge of the craton follows the RMF (black thick dashed line in Figure \ref{mapgeol}), instead of the western breakup line (See the red thin line in Figure~\ref{mapgeol} which coincides with RMF in Canada, but is diverging westward in the U.S.).
@@
 
-		These lateral edges are well observed down to $150km$ depth. 
+		These lateral edges are well observed down to $150 \: km$ depth. 
 		Below that, the cratonic root retracts toward the center of the archaean continent and is still present beneath Slave, Hearne/Rae and western Superior cratons as shown in Figure \ref{cratoncrossedge}.
-		Everywhere within the craton, a negative gradient of $V_S$ is observed, marking the base of the lithosphere, with a minimum of $\sim 4.6km.s^{-1}$ reached at $\sim 300km$ depth. 
-		While the base of the lithosphere (and top of the low velocity zone- LVZ) shows strong topography, the bottom of the LVZ appears largely flat, in particular under the  Slave, Hearne/Rae and western Superior cratons.
+		Everywhere within the lithospheric core, a negative gradient of $V_S$ is observed, marking the base of the lithosphere \citep{eaton2009elusive}, with a minimum of $\sim 4.6 \: km.s^{-1}$ reached at $\sim 300 \: km$ depth.
+		While the base of the lithosphere (and top of the low velocity zone - LVZ) shows strong topography, the bottom of the LVZ appears largely flat, in particular under the Slave, Hearne/Rae and western Superior cratons.
 
-		As observed in \cite{babuvska1998age,gung2003global,yuan2014lithospheric}, there are large lateral negative variations $\xi$ down to $\sim 200km$ depth, with regions of reduced radial anisotropy, in contrast to the more uniform, positive $dln(\xi)$ structure in this depth range in the oceans (See figure \ref{3d-Xi}.
-		Below 250km, it is interesting to note that there is a large band of positive anomaly crossing the whole continent (across the US), making the connection between the Pacific and Atlantic mantles. 
+		As observed in \cite{babuvska1998age,gung2003global,yuan2014lithospheric}, there are large lateral negative variations $\xi$ down to $\sim 200km$ depth, with regions of reduced radial anisotropy, in contrast to the more uniform, positive $dln(\xi)$ structure in this depth range in the oceans (See figure \ref{3d-Xi}).
+
+		{\color{red} I would like to point that we dont see the same features as in \cite{marone2007three} with the craton showing same $\xi \: (> 1)$ than oceanic lithopshhere, while western active part and greenvile-appalachian area shows reduced $\xi < 1$.
+Different causes to me: starting model, I need to compare my and her starting model, or simply the lateral parameterisation (8deg (federica) compared to 2deg (Me)).}
+
+		Below 250km, it is interesting to note that there is a large band of positive $dln(\xi)$ crossing the whole continent (across the US), making the connection between the Pacific and Atlantic mantles.
 		This feature is not present in most of Canada, where negative $dln(\xi)$ structure persist down to 300km.
-		Generally, $\xi$ always appear higher than 1. 
+		Generally, $\xi$ always appears higher than 1. 
 		Except between 60 and 100 km, lower than 1 of $\xi$ ($V_{SV} > V_{SH}$) are observed in Mexico, at the southern part of Florida, the southern part of the Superior craton comprising the Mid Continental rifting, the southern part of the Granit Rhyiolite Province comprising the Failed Rifting area, a large area comprising southern Hearne, Medicine hat, northern Wyoming, B.H.T, Cordillera; and finally beneath Baffin Island.
@@
 		\paragraph{Wyoming and Medecine Hat}
-			The highly reworked Wyoming craton does not show as high anomalies as within the craton nucleus (+2\% intstead of +4-8\%) as shown in figure ~\ref{3d-VS}. 
-			It is split into 2 distinctive part. A western part showing low velocities influenced by the active tectonic with a radially positive gradient. 
+			The Wyoming craton was strongly affected by a regional metamorphic event related to the Trans-Hudson orogen, which juxtaposed a number of Archean cratons during the Paleoproterozoic development of Laurentia \citep{hoffman1989precambrian}.
+			This highly reworked craton does not show as high anomalies as within the craton nucleus (+2\% instead of +4 \:\text{to}\: 8\%) as shown in figure ~\ref{3d-VS}. 
+			It is split into 2 distinctive part. A western part showing low velocities influenced by the active tectonic with a radially positive gradient in terms of $V_S$. 
 			In contrast with the eastern part showing higher velocities with a radially negative then positive gradient which marks the western border of the craton nucleus. 
@@
 			In terms of $\xi$, it is characterized by negative anomalies down to 250km deep. 
-			The northern part shows highest anomalies and $V_{SV}$ appears higher than $V_{SH}$ of about 2\% between 60 and 100km.
+			The northern part shows highest negative $dln(\xi)$ and $V_{SV}$ appears higher than $V_{SH}$ of about 2\% between 60 and 100km.
 This feature is common with the Medicine Hat province (See Figure \ref{3d-xi-iso-100km}).
@@
 		\paragraph{Slave}
-			The Slave craton shows typical lithospheric craton signatures with higher than average velocities (See figures \ref{3d-VS} and \ref{cratoncross}), and negative perturbations of $\xi$ (See Figure \ref{3d-VS}.
-			We can note that the average of $\xi$ is the only one higher than the continental mean, compared to the other Archaean provinces (See Figure \ref{1daverage}.
-			It shows in the entire region a positive velocity gradient with a maximum of ca. $4.8km.s^{-1}$ associated with a maximum of $\xi$ at ca. $150km$ deep. 
-			Deeper, a negative gradient is observed with a minimum of ca. $4.6km.s^{-1}$ at ca. $250km$ deep. 
-			The topography of the extrema appears flat.
-			Although the Slave lithosphere shows highest velocities at $150km$, above $100km$ velocities are slightly higher than the continental average ($dln(V_S)< 4\%$).
+			The Slave craton shows typical lithospheric craton signatures with higher than average velocities (See figures \ref{3d-VS} and \ref{cratoncross}), and negative perturbations of $\xi$ (See Figure \ref{3d-Xi}).
+			We can note that the average of $\xi$ is the only one higher than the continental mean, compared to the other Archaean provinces (See Figure \ref{1daverage}).
+			It shows in the entire region a positive velocity gradient with a maximum of ca.
+$4.7 \: km.s^{-1}$ associated with a maximum of $\xi$ at ca.
+$150 \: km$ deep.
+However we don't observe, both the Central Slave Mantle Conductor and the $4.8 \: km.s^{-1} \: V_S$ profile extending from $160 \: \text{to} \: 250 \: km$, obtained from joint inversion of receiver function, surface wave dispersion curve, and magnetotelluric data of \cite{moorkamp2010joint}.
+			Deeper, a negative gradient is observed with a minimum of ca.
+$4.65 \: km.s^{-1}$ at ca.
+$250 - 270\: km$ deep in agreement with \cite{chen2007new}.
+Not sure I should cite error bars are .0.2kms-1...
+			Although the Slave lithosphere shows highest velocities at $150 \: km$, above $100 \: km$ velocities are slightly higher than the continental average ($dln(V_S)< 4\%$).
+
+			Je voudrais ici mettre en lien avec les mesures de kimberlites/ eclogites de \cite{faure2011seismic,heaman2003timing} Aulbach et al., 2007, Jacob 2004 pour montrer que leur mesures estimant la profondeur de la pLAB (200 a 220) peut ne pas etre compatible avec les mesures tomo car leurs mesures datent d'entre 650Ma et 50Ma.
+Il peut y avoir eu un epaississement du mantle keel, mais je n'ai pas encore trouve d'arguments favoarbles...
+Ou alors que ce genre d'information peut etre mieux contraint via l'aniso azim mais tous leurs papiers comparent la LAB sismique avec Vs only.
+
+			% Two firm LAB estimates come from MT.
+The star data point in the centre of the Slave Craton in northern Canada (Fig.
+2e) is the LAB estimate of Moorkamp et al.
+(2010), derived from joint inver- sion of receiver function data, surface wave dispersion curve data, and magnetotelluric data from the EKTN POLARIS station.
+This eLAB + sLABrf + sLABsw value of 230 km (220–235 km) exceeds by 20–30 km the greatest pLAB values on the Slave Craton of 202 km for the Neoproterozoic Anuri pipe in NW Slave, 208 km from the Jurassic-aged Gahcho Kue pipe (formerly known as Kennedy Lake) in SW Slave, and the 202 km pLAB from the Eocene-aged Diavik pipe in the central Slave (Snyder and Grütter 2010).
+It is though well within the error limits of the poorly-resolved sLABsw esti- mate of 220 ± 60 km of Chen et al.
+(2007) based on only Rayleigh wave inversion from the same station (EKTN); here, the joint in- version of Moorkamp et al.
+(2010) has served to reduce the uncer- tainty considerably.
+			{\color{red}Either the pLAB estimates are too low by 20–30 km, which is likely, or the lithosphere has grown by that order over the last 50 million years since the Eocene kimberlite eruptions in the central Slave, which is less likely, or aspects of both are operating.} \cite{jones2014electrical}
 
@@
 
-		\paragraph{Western part}
+		\paragraph{Western Ontario part}
 			The western Superior craton shows typical lithospheric craton signatures with highest velocities as shown in figure ~\ref{3d-VS} and in figure~\ref{cratoncross}.
 In terms of $\xi$, slight negative to zero perturbations are observed at the very western part, while close to the Mid continental rifting, highest negative perturbations indicate that $V_{SV}$ is higher than $V_{SH}$.
-This last feature is associated the thinning of the cratonic lithosphere when crossing the Grenville front.
-			In the entire region there is a positive velocity gradient with a maximum of ca. $4.8km.s^{-1}$  associated with a maximum of $\xi$ at ca. $150km$ deep.
+This last feature is associated with the thinning of the cratonic lithosphere when crossing the Grenville front as observed by \cite{darbyshire2007new}.
+			In the entire region there is a positive velocity gradient with a maximum of ca. $4.7 \: km.s^{-1}$  associated with a maximum of $\xi$ at ca. $150km$ deep.
 
 			Deeper a negative gradient is observed with a minimum of ca. $4.6km.s^{-1}$ at ca. $250km$ deep. 
-			Opposed to the Slave craton, the topography of the extrema is variable. 
+			Opposed to the Slave craton, the topography of these extrema is variable. 
 			From SW to NE, the core of maximum velocity can reach beween 70 and 150km of thickness. 
@@
 			Following the Great Meteor track built by \cite{heaman2000timing}, which crosses the Superior craton, it is interseting to note that the low velocity channel beneath the base of the lithosphere shows lower velocties than usualy observed beneath the other Archaean cratons.
-			This slow anomaly could correspond to an extension the V-shape dent, or divot, observed by \cite{lee1997upper} who explained the feature by the presence of water in the mantle. 
-			We observe at $250km$ that the slow anomaly go further inland and crosses the Superior Craton northward until the Hudson bay following the GMT.
+			Xenoliths analysis from kimberlites from Kirkaland Lake (south eastern part of the western Superior) exhibit some evidence of diamonds, i.e, the base of the lithosphere was deeper than 170 km at the time of kimberlitic eruption \citep{jones2014electrical} estimated at $160 - 152 \: Ma$ by \cite{heaman2000timing,heaman2003timing}.
+			This can be interpreted as either the lithosphere is today thick, and the lower lithosphere was not sampled by xenolith entrainment in the kimberlitic magmas (but nevertheless diamonds were entrained), or the lowermost lithosphere has been eroded away since eruption, possibly by the movement of the North American plate over the Great Meteor hotspot \citep{heaman2000timing,jones2014electrical}.
+The latter being in agreement with our of observations.
 
-		\paragraph{Eastern part}
+			% This slow anomaly could correspond to an extension the V-shape dent, or divot, observed by \cite{lee1997upper} who explained the feature by the presence of water in the mantle. 
+			% We observe at $250km$ that the slow anomaly go further inland and crosses the Superior Craton northward until the Hudson bay following the GMT.
+
+		\paragraph{Eastern Quebec part}
 			Compared to its western part, the structure here shows differences as shown in figure ~\ref{cratoncrossedge} 
@@
 			It shows typical lithospheric craton signatures with higher than average velocities (See figures \ref{3d-VS} and \ref{cratoncross} associated with negative $\xi$ perturbations. 
-			It is interesting to notice that at shallower depths than $100km$, the THO (including the Sask craton) which welded the Superior and Rae/Hearne cratons, shows lighter (i.e, $+4\%$ compared to $+8\%$) positive anomalies than its archaean surrounding.
+			It is interesting to notice that at shallower depths than $100km$, the THO which welded the Superior and Rae/Hearne cratons, shows lighter (i.e, $+4\%$ compared to $+8\%$) positive anomalies than its archaean surrounding.
 This feature is associated with negative anomalies of $\xi$.
 While most of the proterozoic root start to thin out at $150km$, the THO at the vicinity of the hudson bay shows as high anomalies as Rae/Hearne and Superior areas associated with zero to positive anomalies in term of $\xi$.
+			The THO comprises the archaean aged Sask craton (See Figure \ref{mapgeol}. 
+			The mantle beneath this exotic craton is interpreted to have been detached and replaced by both the Superior and the Hearne mantle lithosphere during final collision of the orogen \citep{nemeth2005lithospheric}.
+			This interpretation, according to \cite{faure2011seismic} can be extrapolated farther to the east and may explain why Trans-Hudson orogen has no imprint in the upper mantle.
 
@@
 
-\subsubsection{Surrounding oceanic} This section needs to be redone...
+\subsubsection{Surrounding oceanic}
 
@@
 Stars are 155 regional events, in addition to 122 teleseismic events.
-Orange stars indiscate events shallower than $30km$ deep, Yellow stars for deeper than $300km$ and red one are in between.
-Thick red line indicates the boundaries used for the RegSEM forward simulation, which extends 89◦ by 89◦ horizontally and down to 1600 km and contains all stations and regional events.}
+Orange stars indiscate events shallower than 30~km deep, yellow stars for deeper than 300~km and red one are in between.
+Thick red line indicates the boundaries used for the RegSEM forward simulation, which extends 89◦ by 89◦ horizontally and down to 1600~km and contains all stations and regional events.}
 
@@
 	\begin{figure}
+		\begin{minipage}{0.5\linewidth}
+			\centerline{\includegraphics[width=\linewidth]{figures/Xi_NASEM-iter3-Xi-iso_60km.png}}
+		\end{minipage}
+		\hfill
+		\begin{minipage}{0.5\linewidth}
 			\centerline{\includegraphics[width=\linewidth]{figures/Xi_NASEM-iter3-Xi-iso_100km.png}}
+		\end{minipage}
+		% \begin{minipage}{0.5\linewidth}
+		% 	\centerline{\includegraphics[width=\linewidth]{figures/Xi_NASEM-iter3-Xi-iso_150km.png}}
+		% \end{minipage}
 
-		\caption{3D radial anisotropic structure of the continent at 100km compared to the Isotroic case. 
+		\caption{\baselineskip 18pt 
+		3D radial anisotropic structure of the continent at 60 and 100 \: km compared to the Isotropic case. 
 		Map view shows variations with respect to the Isotropic case as $\frac{\xi - 1}{1}$ in \%. 

### 2017-03-22 Paper_NA_br.tex
for 3.2 part

--- 2017-03-01 Paper_NA_br.tex
@@
 \usepackage{textcomp}
-\usepackage{graphicx}
+\usepackage[draft]{graphicx}
 \usepackage{natbib} % Powerful citation management
+
+\usepackage{hyperref}
+\hypersetup{pdftitle={Introduction à LaTeX},
+           pdfauthor={Pierre Clouzet},
+           colorlinks=true,
+           plainpages=false,
+           colorlinks=true,
+           urlcolor=blue,
+           citecolor=red,
+           pdfstartview=FitH}
 % \usepackage{float}
@@
 	We divide the model description and interpretation into two parts, focusing first on the continent-wide structure and then, in more detail, on eastern different regions.
+	Purpose:
+	Is there a link between age of the crust and the lipthospheric structure
+	Do images give us insight on cratonozation process - to compare with known process and thermal - chemistry data
+	Deformation of the lithosphere by hotspot basal drag
+
+% 	[42] Throughout the Precambrian it is expected that the processes responsible for the formation of continental lithosphere evolved as Earth cooled and modern-style plate tectonics began.
+In the broadest sense, this is well established seismologically in that the mantle beneath the shields is characterized by elevated seismic wave speeds compared to younger Phanerozoic mantle.
+The high wave speeds beneath the cratons are usually explained in the context of a depleted composition [e.g.,Jordan, 1988].
+
+% [43] In southern Africa, Archean domains are usually characterized by faster seismic wave speeds than surrounding younger Proterozoic mobile belts [e.g., Li and Burke, 2006; Chevrot and Zhao, 2007], with the implication that mantle composition and/or plate formation processes evolved between Archean and Proterozoic times.
+In contrast, in Australia, there is no clear distinction between Archean and Proterozoic mantle lithospheric wave speed structure [e.g., Simons et al., 1999; Fishwick et al., 2005].
+The same is true for Fennoscandia [e.g., Bruneton et al., 2004], whereas Lebedev et al.
+[2009] found systematically slower wave speeds beneath the Proterozoic fold belts of central Asia compared to the cratonic Russian platform.
+
+% [44] The tectonic history of our study area spans ∼3 Ga of the geological record, from Archean formation of the Superior craton to the Paleozoic accretion of the Appalachian terranes to Laurentia.
+Our tomographic images thus provide an excellent opportunity to test whether there is a systematic age-dependence on seismic wave speed anomalies.
 
@@
 
+		En attendant le code de cluster analysisde Marco je n'ai pas touche a cette partie
+
 		As observed in many other studies, the average isotropic shear velocity $V_S$ in the shallow upper mantle beneath North America (NA) is high compared to the global average, which contains contributions from large oceanic regions (Figure ~\ref{1daverage}).
-		The difference persists until ca. $270 \: km$ depth and then fades away, though, $V_S$ remains slightly higher in NA than in the global average. 
+		The difference persists until $\sim$270~km depth and then fades away, though, $V_S$ remains slightly higher in NA than in the global average. 
 		Compared to the continental average, precambrian provinces like Wyoming, Yavapai, Mazatzal, Great Rhyolites and Grenville are caracterized by slower velocities. 
@@
 		Closer to the center of the continent, region like Slave, T.H.O, Juvenile arcs and orogens (e.g.
-Wopmay), Superior; show higher velocities with the Slave province caracterized by a positive gradient above $100 \: km$ deep.
+Wopmay), Superior; show higher velocities with the Slave province caracterized by a positive gradient above 100~km deep.
 
@@
 
-		The average of $\xi$ is higher than $1$ (which means that $V_{SH} > V_{SV}$), but lower than the global average, the latter influenced by the strong $\xi > 1$ oceanic signature.
+		The average of $\xi$ is higher than 1 (which means that $V_{SH} > V_{SV}$), but lower than the global average, the latter influenced by the strong $\xi > 1$ oceanic signature. 
+		Both trends are similar with depth, but the maximum in $\xi$ is slightly deeper than in the global average and the change from $V_{SH} > V_{SV}$ to $V_{SH} < V_{SV}$ is reached at $\sim$420~km. 
+		At shallower depths than 200 to 220~km, strongest variations relative the continental average are observed ($+/- 4 \: \text{to} \: 8 \%$, compared to $+/- 1 \: \text{to} \: 3 \%$). 
+		Alternations of positive and negative gradient are observed from 50 to 400 km with peak values diminishing along depth.
+
+		% A negative gradient is regionally observed with its minima at $100 \Km$ of $\sim \xi = 1.02$ evolving to a positive gradient with a maximum of $\xi = 1.03$ located at $150 \: km$. 
 		% where shearing is more important.
-		Both trends are similar with depth, but the maximum in $\xi$ is slightly deeper than in the global average and the change from $V_{SH} > V_{SV}$ to $V_{SH} < V_{SV}$ is reached at ca. $420km$.
-
-		{\color{red} This feature was already present in the global model, and perturbations below that depth are <1\% So I dont think the model adds more informations compared to SEMucb}.
 
@@
 	% Gung 2003 Positive anomalies of dlnxi down to 400km in the craton.
-
-		Our study confirms the presence of a thick lithospheric root in the craton with faster than average $V_S$ down to $250km$ depth, with highest velocities, ca.
-$4.8km.s^{-1}$, located at depths shallower than  $150km$ (e.g.
-Figure~\ref{cratoncross}).
-		Laterally, the cratonic root in $V_S$ coincides roughly with precambrian aged crust (see Figure~\ref{3d-VS} from 60 to 150km deep)
-		The western edge of the craton follows the RMF (black thick dashed line in Figure \ref{mapgeol}), instead of the western breakup line (See the red thin line in Figure~\ref{mapgeol} which coincides with RMF in Canada, but is diverging westward in the U.S.).
+		\paragraph{age of the crust and its corresponding lithosphere}
+		Our study confirms the presence of a thick lithospheric root in the craton with faster than average $V_S$ down to 250~km depth, with highest velocities, $\sim$4.7~km.s$^{-1}$, located at depths shallower than  150 km (e.g.
+Figure~\ref{cratoncrossedge} and \ref{cratoncrosscenter}).
+		Laterally, the cratonic root in $V_S$ coincides roughly with precambrian aged crust (see Figure~\ref{3d-VS} from 60 to 150~km deep and Figure~\ref{cratoncrossedge}.a ).
+		The western edge of the craton follows the RMF (black thick dashed line in Figure~\ref{mapgeol}), instead of the western breakup line (See the red thin line in Figure~\ref{mapgeol} which coincides with RMF in Canada, but is diverging westward in the U.S.).
 		While the eastern/south-eastern coincides with the Llano-Grenville Front (See the blue thin line in Figure~\ref{mapgeol}). 
@@
 
-		These lateral edges are well observed down to $150km$ depth. 
-		Below that, the cratonic root retracts toward the center of the archaean continent and is still present beneath Slave, Hearne/Rae and western Superior cratons as shown in Figure \ref{cratoncrossedge}.
-		Everywhere within the craton, a negative gradient of $V_S$ is observed, marking the base of the lithosphere, with a minimum of $\sim 4.6km.s^{-1}$ reached at $\sim 300km$ depth. 
-		While the base of the lithosphere (and top of the low velocity zone- LVZ) shows strong topography, the bottom of the LVZ appears largely flat, in particular under the  Slave, Hearne/Rae and western Superior cratons.
+		These lateral edges, coinciding with precambrian crust, are well observed down to 150 km depth. 
+		Below that, the lithospheric core retracts toward the center of the precambrian continent comprising north western Juvenile arcs and orogens (e.g.
+Wopmay and BHT), north-eastern part of Juvenile crusts (e.g Yavapai/Mazatzal/Gr.
+Rhyolite), Slave, Hearne/Rae and western Superior cratons; while eastern Superior and Wyoming are cast out, as shown in Figure \ref{3d-VS}.
+		There are no obvious correlations between the age of a province and the amplitudes of anomalies and their extension in depth. 
+		For example, archaean eastern Superior craton has $\sim 0\% \: V_S$ anomalies below 200 km (compared to $\sim+$ 4 $\%\:V_S$ beneath the western Superior), similar to the proterozoic Grenville province and archaean Wyoming ( also observed in Gao et al.
+2002, Griffin et al.
+1998, Menzies et al.
+1993, Menzies et al.
+2007); while the proterozoic belt of THO shows as high anomalies at the same depth as within the archaean Slave or western Superior lithosphere.
+		It is interesting to notice that at shallower depths than 100 km, the THO which welded the Superior and Rae/Hearne cratons \citep{hoffman1988united}, shows lighter (i.e, $+4\%$ compared to $+8\%$) positive anomalies than its archaean surrounding.
+		The THO comprises the archaean aged Sask craton (See Figure \ref{mapgeol}).
+		The mantle beneath this exotic craton is interpreted to have been detached and replaced by both the Superior and the Hearne mantle lithosphere during final collision of the orogen \citep{nemeth2005lithospheric}.
+		This interpretation, according to \cite{faure2011seismic} can be extrapolated farther to the east and may explain why Trans-Hudson orogen has no imprint in the upper mantle.
 
-		As observed in \cite{babuvska1998age,gung2003global,yuan2014lithospheric}, there are large lateral negative variations $\xi$ down to $\sim 200km$ depth, with regions of reduced radial anisotropy, in contrast to the more uniform, positive $dln(\xi)$ structure in this depth range in the oceans (See figure \ref{3d-Xi}.
-		Below 250km, it is interesting to note that there is a large band of positive anomaly crossing the whole continent (across the US), making the connection between the Pacific and Atlantic mantles. 
-		This feature is not present in most of Canada, where negative $dln(\xi)$ structure persist down to 300km.
-		Generally, $\xi$ always appear higher than 1. 
+		As observed in \cite{babuvska1998age,gung2003global,yuan2014lithospheric}, there are large lateral negative variations $\xi$ down to $\sim 200km$ depth, with regions of reduced radial anisotropy, in contrast to the more uniform, positive $dln(\xi)$ structure in this depth range in the oceans (See figure \ref{3d-Xi} where $dln(\xi)$ are shown according the Isotropic case ($\xi = 1$)).
+
+		Below 250km, it is interesting to note that there is a large band of positive $dln(\xi)$ crossing the whole continent (across the US), making the connection between the Pacific and Atlantic mantles.
+		This feature is not present in most of Canada, where negative $dln(\xi)$ structure persist down to 400km and can be linked with the deep strong $\xi$ layer of \cite{gung2003global}.
+Our higher resolution model shows that this layer of strong $\xi$ may not be present beneath the whole cratonic lithosphere and it seems that this feature is marking a connection with Low Velocity Fingers (LVF) observed by \cite{french2013waveform}.
+% Below, this feature is still slightly visible beneath the US, while negative values of $dln\xi$ are observed beneath Canada
+		% In particular, surface wave anisotropy studies show that cratons are underlain by a layer of radial anisotropy, where horizontally polarized shear waves are faster than vertically polarized shear waves in the depth range of 250–400 km
+		Generally, $\xi$ always appears higher than 1. 
 		Except between 60 and 100 km, lower than 1 of $\xi$ ($V_{SV} > V_{SH}$) are observed in Mexico, at the southern part of Florida, the southern part of the Superior craton comprising the Mid Continental rifting, the southern part of the Granit Rhyiolite Province comprising the Failed Rifting area, a large area comprising southern Hearne, Medicine hat, northern Wyoming, B.H.T, Cordillera; and finally beneath Baffin Island.
-This feature can be better observed in Figure \ref{3d-xi-iso-100km} where perturbations are shown according the Isotropic case ($\xi = 1$)
+		% This feature can be observed in Figure \ref{3d-Xi} where perturbations are shown according the Isotropic case ($\xi = 1$)
 
-	\subsubsection{Archaean provinces}
+		\paragraph{Lithospheric layering}
+		Everywhere within the lithospheric core, from 60 to 100-150 km $V_S$ increases with depth (positive radial gradient) with a maximum of $\sim$4.7-4.8 km.s$^{-1}$ between 100 and 150 km (See Figure~\ref{cratoncrosscenter}).
+This can be an indication of a Mid lithospheric discontinuity within the lithopsheric core.
+		% However, MLD imaging is beyond the scope of this study.  
+		Deeper, a negative gradient of $V_S$ is observed, marking the base of the lithospheric core \citep{eaton2009elusive}, with a minimum of $\sim$4.6 km.s$^{-1}$ reached at $\sim$270 km depth. 
+		Such structure is very well illustrated at the center of the continent under the Slave, Hearne/Rae, THO and western Superior cratons; which is in contrast with the surrounding of the core, where velocities decrease with depths down to at most 100-150 km and deeper increase down to 400 km (see Figure~\ref{cratoncrosscenter} along the FF' and GG' cross sections of radial $V_S$ gradient showing that within the lithospheric core of NA is characterized by a positive gradient down to $\sim$150~km ,and then, a negative gradient down to at least $\sim$250~km.).
+		Beneath other Archean provinces like Wyoming craton (see Figure~\ref{cratoncrosscenter} along the FF' cross section), west to the RMF, such structure is not present. 
+		Velocities decrease down to 120 km and then increase deeper than 400 km. 
+		East of the RMF in Wyoming craton, velocities are monotically decreasing down to 250 km. 
+		Also, beneath the eastern Superior craton, high velocities are confined at shallower depths than 100 km, and the radial structure is monotically decreasing with depth. 
+		% The lithospheric core in North
+		While the base of the lithosphere (and top of the low velocity zone - LVZ) shows strong topography, the bottom of the LVZ appears largely flat, in particular under the Slave, Hearne/Rae and western Superior cratons (See cross sections in Figure~\ref{cratoncrosscenter} and \ref{cratoncrossedge}.
 
-		\paragraph{Wyoming and Medecine Hat}
-			The highly reworked Wyoming craton does not show as high anomalies as within the craton nucleus (+2\% intstead of +4-8\%) as shown in figure ~\ref{3d-VS}. 
-			It is split into 2 distinctive part. A western part showing low velocities influenced by the active tectonic with a radially positive gradient. 
-			In contrast with the eastern part showing higher velocities with a radially negative then positive gradient which marks the western border of the craton nucleus. 
-			The depth of negative gradient's minima is variable, from 70 (western part) to 250km (eastern part) deep. 
-			The separation between the two areas coincides with the RMF.
-			In terms of $\xi$, it is characterized by negative anomalies down to 250km deep. 
-			The northern part shows highest anomalies and $V_{SV}$ appears higher than $V_{SH}$ of about 2\% between 60 and 100km.
-This feature is common with the Medicine Hat province (See Figure \ref{3d-xi-iso-100km}).
+		Regarding the radial evolution of $\xi$, cross section of the $\frac{d\xi}{dz}$ in Figure~\ref{cratoncrosscenter} shows alternances of trends.
+		$\xi$ generally decrease from 50 to $\sim$70-100 km, then increase down to 150 km, decrease again down to $\sim$220-250 km and finally decrease to the Isotropy down to the Transition zone.
+Compared to $\frac{dV_S}{dz}$, $\frac{d\xi}{dz}$ in terms of amplitude is 4 order of magnitude below.
+It appears the band of negative $\frac{d\xi}{dz}$ between 150 and 250 km shows a topography at its upper limit.
+Around the lithopheric core, the band is $\sim$100 km thick between 150 and 250 km deep; while within the lithospheric core, the band is $\sim$50 km thick between 170 and 220 km deep.
+This feature might be a characteristic of the $\xi$ structure beneath a lithospheric core which needs to be confirmed in other cratonic studies.
 
-		\paragraph{Slave}
-			The Slave craton shows typical lithospheric craton signatures with higher than average velocities (See figures \ref{3d-VS} and \ref{cratoncross}), and negative perturbations of $\xi$ (See Figure \ref{3d-VS}.
-			We can note that the average of $\xi$ is the only one higher than the continental mean, compared to the other Archaean provinces (See Figure \ref{1daverage}.
-			It shows in the entire region a positive velocity gradient with a maximum of ca. $4.8km.s^{-1}$ associated with a maximum of $\xi$ at ca. $150km$ deep. 
-			Deeper, a negative gradient is observed with a minimum of ca. $4.6km.s^{-1}$ at ca. $250km$ deep. 
-			The topography of the extrema appears flat.
-			Although the Slave lithosphere shows highest velocities at $150km$, above $100km$ velocities are slightly higher than the continental average ($dln(V_S)< 4\%$).
+		To summarize, as observed in many previous tomographic studies of North America, we observe a lithospheric core beneath the Precambrian shield caracterized by high $V_S$ and reduced $\xi$. 
+		According to our higher resolution model, we dont observe a direct link between the age of the crust and its underlying lithosphere.
+Most of the core is located beneath precambrian crust but the history of the amalgamation and destruction (e.g convective removal, see \cite{lee2011building}) of the craton complexified the structure of its lithosphere.
+		This is the case for the archean Wyoming craton who does not share the same characteristic as other Archaean blocks (e.g Slave, Hearne, Rae, western Superior) and may have suffered from basal erosion \citep{s2005archean,lee2011building}.
+		Moreover, we observe that the eastern lithospheric block of the archean Superior craton, also doesn't share the cratonic lithospheric structure.
+In the next chapter we will focus on this craton that may have endured basal deformation, or weakening, induced by the passage of the Great Meteor chain of hotspot.
 
+	\subsection{Basal traction beneath the lithosphere}
+	% Its thickness is defined by the limit at which the cooler temperatures of the boundary layer can be resolved from the mantle adiabat (which is pinned to the potential temperature of the mantle, i.e., the temperature of the convecting mantle if it were adiabatically decompressed to the surface of Earth in the solid state).
+
+	Notes for interprating Xenoliths studies. SEE DATATION OF GMT Model
+	% Determining the formation age of the underlying continental mantle is more challenging because mantle xenoliths, before entrainment in the host magma, typically reside at temperatures above the closure temperature of most radiometric systems; hence their internal isochron ages often reflect eruption ages
+
+	% Isotopic memory of older events is more likely preserved in whole rocks or mineral inclusions in diamonds.
+For example, radiogenic and unradiogenic whole-rock Sr and Nd isotopes, respectively, indicate Ga isolation in an incompatible-element- enriched environment (Cohen et al.
+1984, Roden & Murthy 1985).
+Similarly, Nd model ages of garnet grains isolated in diamonds give Archean ages (Richardson et al.
+1984).
+However, because Sr and Nd isotopes are incompatible elements, model ages based on these isotopic systems usually date only metasomatic events.
+
+	% Better constraints on the timing of melt extraction and initial isolation from the convecting mantle are obtained from the Re-Os
+
+	Compare studies between ReOs and UPb and what they indicate
 		\paragraph{Superior Craton}
 			In our model, the Superior craton can be split in two parts: The western part showing typical cratonic structures while the eastern part shows slower velocities at all depths.
-Below $150km$, velocity anomalies are below $2\%$ (almost $0\%$).
+Below 200 km, the distinction between the eastern and western part correlates the GMT.
 
-		\paragraph{Western part}
+		\paragraph{Western Ontario part}
 			The western Superior craton shows typical lithospheric craton signatures with highest velocities as shown in figure ~\ref{3d-VS} and in figure~\ref{cratoncross}.
 In terms of $\xi$, slight negative to zero perturbations are observed at the very western part, while close to the Mid continental rifting, highest negative perturbations indicate that $V_{SV}$ is higher than $V_{SH}$.
-This last feature is associated the thinning of the cratonic lithosphere when crossing the Grenville front.
-			In the entire region there is a positive velocity gradient with a maximum of ca. $4.8km.s^{-1}$  associated with a maximum of $\xi$ at ca. $150km$ deep.
+This last feature is associated with the thinning of the cratonic lithosphere when crossing the Grenville front as observed by \cite{darbyshire2007new}.
+			In the entire region there is a positive velocity gradient with a maximum of ca. $4.7 \: km.s^{-1}$ underneathed by a negative gradient  with a minimum of ca. $4.6km.s^{-1}$ at ca. $250km$ deep.
+% 
+			% associated with a maximum of $\xi$ at ca. $150km$ deep.
 
-			Deeper a negative gradient is observed with a minimum of ca. $4.6km.s^{-1}$ at ca. $250km$ deep. 
-			Opposed to the Slave craton, the topography of the extrema is variable. 
+			Opposed to the Slave craton, the topography of the top of the negative gradient is variable. 
 			From SW to NE, the core of maximum velocity can reach beween 70 and 150km of thickness. 
-			Beneath the maximum velocity region of ca. $150km$ the negative gradient appears stipper than beneath the $70km$ one, while the topography of the deeper positive gradient appears flat.
+			Beneath the maximum velocity region of  $\sim$ 150 km the negative gradient appears stipper than beneath the $70km$ one, while the topography of the deeper positive gradient appears flat.
+The topography of the bottom of the negative gradient is observed flat {\color{red} It is the same in Slave - One linkk to share with other regions}
 
-			The band of minimum velocity within the cratonic root has similar amplitudes, in terms of velocity, than beneath the younger lithosphere but at shalower depths between $100$ and $150 \: km$. 
-			Following the Great Meteor track built by \cite{heaman2000timing}, which crosses the Superior craton, it is interseting to note that the low velocity channel beneath the base of the lithosphere shows lower velocties than usualy observed beneath the other Archaean cratons.
-			This slow anomaly could correspond to an extension the V-shape dent, or divot, observed by \cite{lee1997upper} who explained the feature by the presence of water in the mantle. 
-			We observe at $250km$ that the slow anomaly go further inland and crosses the Superior Craton northward until the Hudson bay following the GMT.
+			% The band of minimum velocity within the cratonic root has similar amplitudes, in terms of velocity, than beneath the younger lithosphere but at shalower depths between $100$ and $150 \: km$. 
+			Following the Great Meteor track built by \cite{heaman2000timing}, which crosses the Superior craton, it is interesting to note that the low velocity channel beneath the base of the lithosphere shows lower velocties than usualy observed beneath the other Archaean cratons.
+			Xenoliths analysis from kimberlites at Kirkaland Lake (south eastern part of the western Superior) exhibit some evidence of diamonds, i.e, the base of the lithosphere was deeper than 170 km at the time of kimberlitic eruption \citep{jones2014electrical} estimated at $160 - 152 \: Ma$ by \cite{heaman2000timing,heaman2003timing}.
+			This can be interpreted as either the lithosphere is today thick, and the lower lithosphere was not sampled by xenolith entrainment in the kimberlitic magmas (but nevertheless diamonds were entrained); or the lowermost lithosphere has been eroded away since eruption, possibly by the movement of the North American plate over the Great Meteor hotspot \citep{heaman2000timing,jones2014electrical}.
+In the study of \cite{faure2011seismic}, 23 $\%$ of kimberlite samples from southern Superior region contain diamonds.
+In contrast with 78 $\%$ at the northern Superior.
+This means that the diamond stability field within the mantle has been deteriorated through time and kimberlites that have passed across the mantle during/after such deterioration have less chances to intercept diamonds.
+\citep{faure2011seismic}
 
-		\paragraph{Eastern part}
-			Compared to its western part, the structure here shows differences as shown in figure ~\ref{cratoncrossedge} 
-			Fast velocities are present down to $150 \: km$ deep but are slower and no perturbations are observed below $200 \: km$. 
-			The negative gradient is steeper and the minimum is radially thicker (more than $100 \: km$ thick compare to $\sim 50 \: km$ beneath the craton nucleus). 
-			% However at the center, the minimum channel is not continuous and shows higher (still slower than the center of the western Superior) velocities.
-			% the gradual trend in velocity is replaced by an almost constant slow anomaly from $150$ to $300km$ deep.
+			\cite{kaban2015cratonic} proposed that basal drag induced by plate motion caused a south-western shift of the Superior's cratonic root.
+A feature that was observed by the local inversion of gravity, topography, crustal structure and seismic data.
 
-		\paragraph{Trans Hudson Orogen}
-			The Trans Hudson Orogen (THO) is a belt of reworked archaean aged region.
+			This basal deformation would have been enabled by the reheating of the lithosphere due to the passage of the Great Meteor hotpsot \citep[see][]{eaton2007seismic,kaban2015cratonic}.
+			Our observations show that despite the archaean age of the Superior craton, the lithospheric root thickness of a craton is not flat and can be altered as broadly observed within the eastern part of the Superior craton.
+
+			% Consequently, the diamond stability field has been partially or totally destroyed by thermal and/or chemical processes at dif- ferent times and in various areas of the upper mantle.
+The global and permanent deterioration of the diamond stability field through time implies that kimberlites that have passed across the mantle during or after an episode of metasomatism have less chance of intercepting preserved pockets of dia- monds.
+This explains why only 23 percent of the 96 kimber- lites in the southern Superior are diamondiferous and why the proportion of macro diamonds is very low; only seven kimberlites contain at least one macrodiamond.
+			% In contrast, 78 percent of the 55 kimberlites of the Central Superior craton, including the Victor mine at Attawapiskat and the Renard feasibility mine project in the Otish moun- tains, are located in the typical cold and depleted Archean shell, contain abundant diamonds with a high proportion of macro- and gem-quality diamonds and have a strong harzbur- gitic signature (Grütter et al., 2006).
+
+			% This slow anomaly could correspond to an extension the V-shape dent, or divot, observed by \cite{lee1997upper} who explained the feature by the presence of water in the mantle. 
+			% We observe at $250km$ that the slow anomaly go further inland and crosses the Superior Craton northward until the Hudson bay following the GMT.
+
+		\paragraph{Eastern Quebec part}
+			Compared to its western part, the structure here is different than typical cratonic structure as shown in figure ~\ref{cratoncrossedge} 
+			Fast velocities are present down to 150 km marking the shelf of the lithospheric nucleus. 
+			Below, zero to slight positive anomalies caracterize the eastern superior craton (as in the wedded Grenville belt) showing the retractation toward the continent's center of lithospheric nucleus.
+A feature that was not present in our previous less well resolved models \citep{marone2007three,yuan20113,yuan2014multiscale}.
+			Except at its south western part, negative gradient in terms of velocity is present from 50 to 250 km deep. 
+			Whereas at the center of the continent, velocities are usually increasing (positive gradient) between 50 and 150 km, here it monotically decreases.
+			The negative gradient is steeper and the minimum is radially thicker (more than 100 km thick compare to $\sim$50 km beneath the craton nucleus).
+
+			Observed also by Villemaire 2012 with P-wave arrival time. Nice discussion in the end of the paper about lithospheric thinning.
+
+% 		\paragraph{Trans Hudson Orogen}
+% 			The Trans Hudson Orogen (THO) is a belt of reworked archaean aged region.
 It refers to a major Paleo Proterozoic tectonic event (between $2.0$ to $1.8Ga$) that welded the Hearne, Wyoming, and Superior archaean blocks to form core of Laurentia \citep{hoffman1988united,whitmeyer2007tectonic}.
-			% At shallow depths ($ < 60km$) shows lower velocities than archaean surroundings. 
-			It shows typical lithospheric craton signatures with higher than average velocities (See figures \ref{3d-VS} and \ref{cratoncross} associated with negative $\xi$ perturbations. 
-			It is interesting to notice that at shallower depths than $100km$, the THO (including the Sask craton) which welded the Superior and Rae/Hearne cratons, shows lighter (i.e, $+4\%$ compared to $+8\%$) positive anomalies than its archaean surrounding.
-This feature is associated with negative anomalies of $\xi$.
-While most of the proterozoic root start to thin out at $150km$, the THO at the vicinity of the hudson bay shows as high anomalies as Rae/Hearne and Superior areas associated with zero to positive anomalies in term of $\xi$.
+% 			% At shallow depths ($ < 60km$) shows lower velocities than archaean surroundings. 
+% 			It shows typical lithospheric craton signatures with higher than average velocities (See figures \ref{3d-VS} and \ref{cratoncross} associated with negative $\xi$ perturbations. 
+% 			It is interesting to notice that at shallower depths than $100km$, the THO which welded the Superior and Rae/Hearne cratons, shows lighter (i.e, $+4\%$ compared to $+8\%$) positive anomalies than its archaean surrounding.
+% 			This feature is associated with negative anomalies of $\xi$. 
+% 			While most of the proterozoic root start to thin out at $150km$, the THO at the vicinity of the hudson bay shows as high anomalies as Rae/Hearne and Superior areas associated with zero to positive anomalies in term of $\xi$.
+% 			The THO comprises the archaean aged Sask craton (See Figure \ref{mapgeol}. 
+% 			The mantle beneath this exotic craton is interpreted to have been detached and replaced by both the Superior and the Hearne mantle lithosphere during final collision of the orogen \citep{nemeth2005lithospheric}.
+% 			This interpretation, according to \cite{faure2011seismic} can be extrapolated farther to the east and may explain why Trans-Hudson orogen has no imprint in the upper mantle.
 
-	\subsubsection{Yavapai/Mazatzal/Granit-Rhyolite region}
+% 		\paragraph{Rae-Hearne}
+% 			The Rae-Hearne craton shows typical lithospheric craton signatures with higher than average velocities (See figures \ref{3d-VS} and \ref{cratoncross}), and negative perturbations of $\xi$ (See Figure \ref{3d-Xi}).
 
-		Yavapai/Mazatzal/Granit-Rhyolite orogens from 1.8 to 1.6 Ga represent zones of accretion for Proterozoic material and mark the amalgamation of the archean cores. \citep{hoffman1988united} \\
-		West to the RMF, the region shows low velocities influenced by the active tectonic with a radially positive gradient. 
-		The north eastern part shows velocity anomalies slightly lower than the Archaean nucleus. Velocity maxima of $4.8km.s^{-1}$ are located shalower than $100km$. 
-		Below, the radially negative gradient is observed and shows the same characteristic than the eastern Superior craton.
+% 			It shows in the entire region a positive velocity gradient with a maximum of ca.
+$4.7 \: km.s^{-1}$ associated with a maximum of $\xi$ at ca.
+$120 \: km$ deep.
+This gradient is at both top and bottom flat, steep until 170 km and smooth below.
+
+% 			Between 150 km and $\sim$270 km a negative gradient is observed with a minimum of ca. $4.65 \: km.s^{-1}$ at ca. $250 - 270\: km$ deep.
+% 			The nortern part marks the shelf of the lithosphere nucleus by its thinning.
+
+% 	\subsubsection{Yavapai/Mazatzal/Granit-Rhyolite region}
+
+% 		Yavapai/Mazatzal/Granit-Rhyolite orogens from 1.8 to 1.6 Ga represent zones of accretion for Proterozoic material and mark the amalgamation of the archean cores. \citep{hoffman1988united} \\
+% 		West to the RMF, the region shows low velocities influenced by the active tectonic with a radially positive gradient. 
+% 		The north eastern part shows velocity anomalies slightly lower than the Archaean nucleus. Velocity maxima of $4.8km.s^{-1}$ are located shalower than $100km$. 
+% 		Below, the radially negative gradient is observed and shows the same characteristic than the eastern Superior craton.
 It is interesting to note that, beneath Nebraska and Kansas, a high velocity anomaly of about $4.7km.s^{-1}$ is present until ca.
 $170km$ (while it is $120km$ in most of the region) where cretaceous aged kimberlites are sampled.
-		These kimberlites indicate a lithosphere thickness of about $160km$ and have been affected by asthenospheric upwelling related to the Mid-Continent Rift System. \citep{griffin2004lithosphere}
+% 		These kimberlites indicate a lithosphere thickness of about $160km$ and have been affected by asthenospheric upwelling related to the Mid-Continent Rift System. \citep{griffin2004lithosphere}
 
-% Basal drag from Kaban 2015 cannot be visible from temperature sensitive shear waves model, we dont see it but it is normal.
+% % Basal drag from Kaban 2015 cannot be visible from temperature sensitive shear waves model, we dont see it but it is normal.
 
-\subsubsection{Surrounding oceanic} This section needs to be redone...
+\subsubsection{Surrounding oceanic}
 
@@
 \section{Figure}
+\setkeys{Gin}{draft=false}
 
@@
 
+	% % Xsections of the craton
+	% \begin{figure}
+	% 	\begin{minipage}{0.5\linewidth}
+	% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_A.pdf}}
+	% 	\end{minipage}
+	% 	\hfill
+	% 	\begin{minipage}{0.5\linewidth}
+	% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_B.pdf}}
+	% 	\end{minipage}
+
+	% 	\begin{minipage}{0.5\linewidth}
+	% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_gradient_A.pdf}}
+	% 	\end{minipage}
+	% 	\hfill
+	% 	\begin{minipage}{0.5\linewidth}
+	% 		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_gradient_B.pdf}}
+	% 	\end{minipage}
+
+	% 	\caption{\baselineskip 18pt
+	% 	Cross Section of the lithospheric core. 
+	% 	The left side shows cross section of $V_S$ (top) and $\xi$ perturbation from the Isotropic case ($\xi = 1$).
+The right side shows the radial gradient defined as $\frac{dV_S}{dz}$ in the top panel, and as $\frac{d\xi}{dz}$.}
+	% 	\label{cratoncross}
+
+	% \end{figure}
+
 	% Xsections of the craton
 	\begin{figure}
-		\begin{minipage}{0.5\linewidth}
-			\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_A.pdf}}
-		\end{minipage}
-		\hfill
-		\begin{minipage}{0.5\linewidth}
-			\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_B.pdf}}
-		\end{minipage}
 
-		\caption{Cross Section of the lithospheric core}
-		\label{cratoncross}
+		\centerline{\includegraphics[width=\linewidth]{figures/Cross-section-with-geoltest.pdf}}
+
+		\caption{\baselineskip 18pt
+		Cross Section of the lithospheric core at its edges. 
+		Cross Section of the lithospheric core with absolute velocities of $V_S$ in km.s$^{-1}$.}
+		\label{cratoncrossedge}
 
@@
 	\begin{figure}
-		\begin{minipage}{0.5\linewidth}
-			\centerline{\includegraphics[width=\linewidth]{figures/profiles_WE-50.pdf}}
-		\end{minipage}
-		\hfill
-		\begin{minipage}{0.5\linewidth}
-			\centerline{\includegraphics[width=\linewidth]{figures/profiles_SN-100.pdf}}
-		\end{minipage}
 
-		\caption{Cross Section of the lithospheric core showing the core retraction}
-		\label{cratoncrossedge}
+		\centerline{\includegraphics[width=\linewidth]{figures/Cross-gradient.pdf}}
+
+		\caption{\baselineskip 18pt
+		Cross Section of the lithospheric core. 
+		In each panel cross section of $V_S$ (top) and $\xi$ perturbation from the Isotropic case ($\xi = 1$) (bottom).
+		The right side shows cross section of $V_S$ (top) and $\xi$ perturbation from the Isotropic case ($\xi = 1$).
+The left side shows the radial gradient defined as $\frac{dV_S}{dz}$ in the top panel, and as $\frac{d\xi}{dz}$.}
+		\label{cratoncrosscenter}
 
@@
 
+	% \setkeys{Gin}{draft=false}
+		% Xsections of the craton
+	% \begin{figure}
+
+	% 	\fboxsep 0pt
+	% 	\begin{minipage}{0.48\linewidth}
+	% 		\begin{center}
+	% 		(a) \\ \includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxiSN-western-boarder.pdf}
+	% 		\end{center}
+	% 	\end{minipage}
+	% 	\hfill
+	% 	\begin{minipage}{0.48\linewidth}
+	% 		\begin{center}
+	% 		(b) \\ \includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxiSN-esatern-boarder.pdf}
+	% 		\end{center}	
+	% 	\end{minipage}
+
+	% 	\vspace{3mm}
+
+	% 	\begin{minipage}{0.48\linewidth}
+	% 		\begin{center}
+	% 		(c) \\ \includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxiWE-northern-boarder.pdf}
+	% 		\end{center}
+	% 	\end{minipage}
+	% 	\hfill
+	% 	\begin{minipage}{0.48\linewidth}
+	% 		\begin{center}
+	% 		(d) \\ \includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxiWE-southern-boarder.pdf}
+	% 		\end{center}
+	% 	\end{minipage}
+
+	% 	\caption{\baselineskip 18pt
+	% 	Cross Section of the lithospheric core at its edges. 
+	% 	In each panel cross section of $V_S$ (top) and $\xi$ perturbation from the Isotropic case ($\xi = 1$) (bottom).}
+	% 	\label{cratoncrossedge}
+
+	% \end{figure}
+
+% \setkeys{Gin}{draft=false}
 	% Xsection of the GMT track
@@
 	% dIso for Xi
-	\begin{figure}
-		\centerline{\includegraphics[width=\linewidth]{figures/Xi_NASEM-iter3-Xi-iso_100km.png}}
+	% \begin{figure}
+	% 	\begin{minipage}{0.5\linewidth}
+	% 		\centerline{\includegraphics[width=\linewidth]{figures/Xi_NASEM-iter3-Xi-iso_60km.png}}
+	% 	\end{minipage}
+	% 	\hfill
+	% 	\begin{minipage}{0.5\linewidth}
+	% 		\centerline{\includegraphics[width=\linewidth]{figures/Xi_NASEM-iter3-Xi-iso_100km.png}}
+	% 	\end{minipage}
+	% 	% \begin{minipage}{0.5\linewidth}
+	% 	% 	\centerline{\includegraphics[width=\linewidth]{figures/Xi_NASEM-iter3-Xi-iso_150km.png}}
+	% 	% \end{minipage}
 
-		\caption{3D radial anisotropic structure of the continent at 100km compared to the Isotroic case. 
-		Map view shows variations with respect to the Isotropic case as $\frac{\xi - 1}{1}$ in \%. 
-		Map view is saturated to better show where $V_{SV}$ is stronger than $V_{SH}$.
-		Below each map, depth and its regional shear velocity mean is indicated. Minimum and maximum perturbations are also indicated.}
-		\label{3d-xi-iso-100km}
+	% 	\caption{\baselineskip 18pt 
+	% 	3D radial anisotropic structure of the continent at 60 and 100 \: km compared to the Isotropic case. 
+	% 	Map view shows variations with respect to the Isotropic case as $\frac{\xi - 1}{1}$ in \%. 
+	% 	Map view is saturated to better show where $V_{SV}$ is stronger than $V_{SH}$.
+	% 	Below each map, depth and its regional shear velocity mean is indicated. Minimum and maximum perturbations are also indicated.}
+	% 	\label{3d-xi-iso-100km}
 
-	\end{figure}
+	% \end{figure}
+
 \bibliographystyle{apalike}

### 2017-04-27 Paper_NA_br.tex
270417

@@
 	% Gung 2003 Positive anomalies of dlnxi down to 400km in the craton.
-		\paragraph{age of the crust and its corresponding lithosphere}
+		\subsubsection{Lateral edges}
 		Our study confirms the presence of a thick lithospheric root in the craton with faster than average $V_S$ down to 250~km depth, with highest velocities, $\sim$4.7~km.s$^{-1}$, located at depths shallower than  150 km (e.g.
@@
 		Laterally, the cratonic root in $V_S$ coincides roughly with precambrian aged crust (see Figure~\ref{3d-VS} from 60 to 150~km deep and Figure~\ref{cratoncrossedge}.a ).
-		The western edge of the craton follows the RMF (black thick dashed line in Figure~\ref{mapgeol}), instead of the western breakup line (See the red thin line in Figure~\ref{mapgeol} which coincides with RMF in Canada, but is diverging westward in the U.S.).
-		While the eastern/south-eastern coincides with the Llano-Grenville Front (See the blue thin line in Figure~\ref{mapgeol}). 
+		The western edge of the craton follows the RMF (black thick dashed line in Figure~\ref{3d-VS}), instead of the western breakup line (See the red dashed line in Figure~\ref{3d-VS} which coincides with RMF in Canada, but is diverging westward in the U.S.).
+		While the eastern/south-eastern coincides with the Llano-Grenville Front (See the blue thin line in Figure~\ref{3d-VS}) instead of the 650 Ma easter rift margin as observed in \cite{yuan2014multiscale} and \cite{schaeffer2014imaging}.
+		However at 60~km, fast velocities are in better agreement with the 650 Ma easter rift margin. 
 		The northern part ends at the coastline of the Arctic ocean including the arctic archipelago and Greenland.
 
+		\subsubsection{age of the crust and its corresponding lithosphere}
 		These lateral edges, coinciding with precambrian crust, are well observed down to 150 km depth. 
@@
 Rhyolite), Slave, Hearne/Rae and western Superior cratons; while eastern Superior and Wyoming are cast out, as shown in Figure \ref{3d-VS}.
-		There are no obvious correlations between the age of a province and the amplitudes of anomalies and their extension in depth. 
-		For example, archaean eastern Superior craton has $\sim 0\% \: V_S$ anomalies below 200 km (compared to $\sim+$ 4 $\%\:V_S$ beneath the western Superior), similar to the proterozoic Grenville province and archaean Wyoming ( also observed in Gao et al.
+		This is consistent with the global study of \cite{bozdag2016global} but in contrast with recent regional \cite[e.g.][]{yuan2014lithospheric,yuan20113,schaeffer2014imaging,nettles2008radially} and global \cite[e.g][]{french2014whole,french200} studies, where the core is located at the north-east part of the continent.
+
+		There are no obvious correlations between the age of a province with amplitude of anomalies and their extension in depth.
+Such observation is in agreement with several studies ( \cite[e.g for Baltic shield study][]{bruneton2004complex}, \cite[for Australian shield][]{fishwick2005contrasts},\cite[e.g for north American shield study][]{villemaire2012p,darbyshire2013seismic,porritt2015lithospheric} ).
+
+		For example, in~\ref{3d-VS}, archaean eastern Superior craton has $\sim 0\% \: V_S$ anomalies below 200 km (compared to $\sim+$ 4 $\%\:V_S$ beneath the western Superior), similar to the proterozoic Grenville province as observed by \cite{villemaire2012p,bozdag2016global} and archaean Wyoming ( also observed in Gao et al.
 2002, Griffin et al.
@@
 1993, Menzies et al.
-2007); while the proterozoic belt of THO shows as high anomalies at the same depth as within the archaean Slave or western Superior lithosphere.
-		It is interesting to notice that at shallower depths than 100 km, the THO which welded the Superior and Rae/Hearne cratons \citep{hoffman1988united}, shows lighter (i.e, $+4\%$ compared to $+8\%$) positive anomalies than its archaean surrounding.
-		The THO comprises the archaean aged Sask craton (See Figure \ref{mapgeol}).
+2007); while the proterozoic belt of THO at its western part, shows as high anomalies at the same depth as within the archaean Slave or western Superior lithosphere.
+
+		Everywhere within the lithospheric core, from 60 to 100-150 km $V_S$ increases with depth (positive velocity gradient with depth) with a maximum of $\sim$4.7-4.8~km.s$^{-1}$ between 100 and 150 km (See Figure~\ref{cratoncrosscenter} and \ref{cratoncrossedge}).
+		Deeper, a negative gradient of $V_S$ is observed, marking the base of the lithospheric core \citep{eaton2009elusive}, with a minimum of $\sim$4.6 km.s$^{-1}$ that can be reached down $\sim$270 km deep.
+		We identify the cratonic root by the contour of $V_S$ = 4.68~km.s$^{-1}$ in Figure~\ref{cratoncrossedge}, which is chosen for clarity in order to show how variable the depth of the cratonic root is observed in our model.
+		The $V_S$ = 4.68~km.s$^{-1}$ dashed line corresponds roughly to the middle of a zone of decreasing $V_S$ (negative velocity gradient with depth) which reaches a minimum almost uniformly around 250-270 km depth, although there are lateral variations in the strength of the minimum.
+
+		\paragraph{Superior craton}
+		The deepest roots are found in the Western and southwestern Superior craton where velocities in excess of 4.8~km.s$^{-1}$ are found at depths between 250 and 270 km (e.g.
+fig~\ref{cratoncrossedge} profiles BB' and EE').
+		In contrast, the eastern Superior craton lithosphere is much thinner, around 150 km, with fastest velocities correspondingly peaking at depths less than 100 km as shown in fig~\ref{cratoncrossedge} profiles BB' and FF'.
+Such observations show that, the link between the age of the crust and the thickness of the lithospheric upper mantle beneath, is not straighforward.
+		% But implies a late alteration of the Superior craton since its cratonization in archaean times. See next part...
+
+		\paragraph{Slave craton}
+		In the north, the Slave cratonic lithosphere is $\sim$250~km thick (see Fig~\ref{cratoncrossedge}, cross sections AA' and DD'), but does not show as fast velocities as within the southwestern/western superior craton.
+		With shear velocities in the depth range 100-150 km not exceeding 4.7~km.s$^{-1}$.
+A value slower than in \cite{chen2007new}.
+I only have studies from RF so far(Snyder etc...), I am digging for more recent papers...
+
+		\paragraph{THO}
+		At the western part (Canadian part only) of the THO, the lithosphere is 250~km thick (see fig~\ref{cratoncrossedge} profiles AA' and FF' and profile CC' in fig.~\ref{cratoncrosscenter}), while a 100~km thinning is observed at the eastern part (i.e beneath the Hudson bay)
+
+		It is interesting to notice that at shallower depths than 100 km, the western part of the THO which welded the Superior and Rae/Hearne cratons \citep{hoffman1988united}, shows lighter (i.e, $+4\%$ compared to $+8\%$) positive anomalies than its archaean surrounding.
+		The western part of the THO comprises the archaean aged Sask craton (See Figure \ref{mapgeol}).
 		The mantle beneath this exotic craton is interpreted to have been detached and replaced by both the Superior and the Hearne mantle lithosphere during final collision of the orogen \citep{nemeth2005lithospheric}.
 		This interpretation, according to \cite{faure2011seismic} can be extrapolated farther to the east and may explain why Trans-Hudson orogen has no imprint in the upper mantle.
+However, the Hudson bay's local study from \cite{darbyshire2013seismic} , showed that the THO shows a channel of slow velocities compare to its surrounding down to 180~km deep.
 
+		\paragraph{Yavapai/Mazatzal/Granit-Rhyolite orogens}
+		The southernmost extension of the craton (Yavapai/Mazatzal/Granit-Rhyolite orogens) represent zones of accretion for Proterozoic material and mark the amalgamation of the archean cores \citep{hoffman1988united}.
+It shows a weaker but deep root (cross sections EE' and CC' in fig.
+~\ref{cratoncrossedge} and map at 250 km depth in fig.
+\ref{3d-VS}) with velocities of almost 4.7~km.s$^{-1}$.
+There is some indication of incursion from rifting (section EE') separating these younger crustal regions from the Superior craton.
+		It is interesting to note that, beneath Nebraska and Kansas, the high velocity anomaly is located where cretaceous aged kimberlites are sampled. 
+		These kimberlites indicate a lithosphere thickness of about 160~km and have been affected by asthenospheric upwelling related to the Mid-Continent Rift System. \citep{griffin2004lithosphere}
+
+		The minimum in velocity is present everywhere at a roughy constant depth beneath the cratons while the base of this low velocity zone (LVZ) is found everywhere, including outside of the cratons between 300 and 325 km depth.
+That base is very smooth, compared to the structures above, and its minor undulations mirror the thickness of the lithosphere above, with slight upwarping beneath the deepest cratons.
+		The LVZ is very thin and velocity reductions are small (min.
+~4.65~km.s$^{-1}$) beneath regions where the lithosphere is thicker than 200 km, but thicker and more pronounced under the eastern superior (min.
+~4.58~km.s$^{-1}$).
+In the regions bordering the cratons, i.e.
+outside of RMF and Grenville Front, the low velocity minimum is between 100-150~km and very pronounced ($V_S$ < 4.4~km.s$^{-1}$), and the transition out of it is more gradual, extending over ~150 km depth to reach velocities of 4.7~km.s$^{-1}$.
+
+		\subsubsection{Radial anisotropy structure}
 		As observed in \cite{babuvska1998age,gung2003global,yuan2014lithospheric}, there are large lateral negative variations $\xi$ down to $\sim 200km$ depth, with regions of reduced radial anisotropy, in contrast to the more uniform, positive $dln(\xi)$ structure in this depth range in the oceans (See figure \ref{3d-Xi} where $dln(\xi)$ are shown according the Isotropic case ($\xi = 1$)).
@@
 		Except between 60 and 100 km, lower than 1 of $\xi$ ($V_{SV} > V_{SH}$) are observed in Mexico, at the southern part of Florida, the southern part of the Superior craton comprising the Mid Continental rifting, the southern part of the Granit Rhyiolite Province comprising the Failed Rifting area, a large area comprising southern Hearne, Medicine hat, northern Wyoming, B.H.T, Cordillera; and finally beneath Baffin Island.
+
+		Still need to do talk about the periodi upward of xi from depth to link with \ref{cratoncrossedge-xi}
 		% This feature can be observed in Figure \ref{3d-Xi} where perturbations are shown according the Isotropic case ($\xi = 1$)
 
-		\paragraph{Lithospheric layering}
-		Everywhere within the lithospheric core, from 60 to 100-150 km $V_S$ increases with depth (positive radial gradient) with a maximum of $\sim$4.7-4.8 km.s$^{-1}$ between 100 and 150 km (See Figure~\ref{cratoncrosscenter}).
+		% \paragraph{Lithospheric layering}
+		% Everywhere within the lithospheric core, from 60 to 100-150 km $V_S$ increases with depth (positive radial gradient) with a maximum of $\sim$4.7-4.8 km.s$^{-1}$ between 100 and 150 km (See Figure~\ref{cratoncrosscenter}).
 This can be an indication of a Mid lithospheric discontinuity within the lithopsheric core.
-		% However, MLD imaging is beyond the scope of this study.  
-		Deeper, a negative gradient of $V_S$ is observed, marking the base of the lithospheric core \citep{eaton2009elusive}, with a minimum of $\sim$4.6 km.s$^{-1}$ reached at $\sim$270 km depth. 
-		Such structure is very well illustrated at the center of the continent under the Slave, Hearne/Rae, THO and western Superior cratons; which is in contrast with the surrounding of the core, where velocities decrease with depths down to at most 100-150 km and deeper increase down to 400 km (see Figure~\ref{cratoncrosscenter} along the FF' and GG' cross sections of radial $V_S$ gradient showing that within the lithospheric core of NA is characterized by a positive gradient down to $\sim$150~km ,and then, a negative gradient down to at least $\sim$250~km.).
-		Beneath other Archean provinces like Wyoming craton (see Figure~\ref{cratoncrosscenter} along the FF' cross section), west to the RMF, such structure is not present. 
-		Velocities decrease down to 120 km and then increase deeper than 400 km. 
-		East of the RMF in Wyoming craton, velocities are monotically decreasing down to 250 km. 
-		Also, beneath the eastern Superior craton, high velocities are confined at shallower depths than 100 km, and the radial structure is monotically decreasing with depth. 
-		% The lithospheric core in North
-		While the base of the lithosphere (and top of the low velocity zone - LVZ) shows strong topography, the bottom of the LVZ appears largely flat, in particular under the Slave, Hearne/Rae and western Superior cratons (See cross sections in Figure~\ref{cratoncrosscenter} and \ref{cratoncrossedge}.
+		% % However, MLD imaging is beyond the scope of this study.  
+		% Deeper, a negative gradient of $V_S$ is observed, marking the base of the lithospheric core \citep{eaton2009elusive}, with a minimum of $\sim$4.6 km.s$^{-1}$ reached at $\sim$270 km depth. 
+		% Such structure is very well illustrated at the center of the continent under the Slave, Hearne/Rae, THO and western Superior cratons; which is in contrast with the surrounding of the core, where velocities decrease with depths down to at most 100-150 km and deeper increase down to 400 km (see Figure~\ref{cratoncrosscenter} along the FF' and GG' cross sections of radial $V_S$ gradient showing that within the lithospheric core of NA is characterized by a positive gradient down to $\sim$150~km ,and then, a negative gradient down to at least $\sim$250~km.).
+		% Beneath other Archean provinces like Wyoming craton (see Figure~\ref{cratoncrosscenter} along the FF' cross section), west to the RMF, such structure is not present. 
+		% Velocities decrease down to 120 km and then increase deeper than 400 km. 
+		% East of the RMF in Wyoming craton, velocities are monotically decreasing down to 250 km. 
+		% Also, beneath the eastern Superior craton, high velocities are confined at shallower depths than 100 km, and the radial structure is monotically decreasing with depth. 
+		% % The lithospheric core in North
+		% While the base of the lithosphere (and top of the low velocity zone - LVZ) shows strong topography, the bottom of the LVZ appears largely flat, in particular under the Slave, Hearne/Rae and western Superior cratons (See cross sections in Figure~\ref{cratoncrosscenter} and \ref{cratoncrossedge}.
 
-		Regarding the radial evolution of $\xi$, cross section of the $\frac{d\xi}{dz}$ in Figure~\ref{cratoncrosscenter} shows alternances of trends.
-		$\xi$ generally decrease from 50 to $\sim$70-100 km, then increase down to 150 km, decrease again down to $\sim$220-250 km and finally decrease to the Isotropy down to the Transition zone.
+		% Regarding the radial evolution of $\xi$, cross section of the $\frac{d\xi}{dz}$ in Figure~\ref{cratoncrosscenter} shows alternances of trends.
+		% $\xi$ generally decrease from 50 to $\sim$70-100 km, then increase down to 150 km, decrease again down to $\sim$220-250 km and finally decrease to the Isotropy down to the Transition zone.
 Compared to $\frac{dV_S}{dz}$, $\frac{d\xi}{dz}$ in terms of amplitude is 4 order of magnitude below.
@@
 	\subsection{Basal traction beneath the lithosphere}
+
+	Ici faire le lien avec les circle figure, fingers de Scott. Probleme pratique: La carte geol ne va pas jusquau MAR et donc afficher le track n'ira pas jusqau bout...  Avec la figure \ref{Vs-circle}
+
+		As in \cite{schaeffer2014imaging,yuan2014lithospheric,james2014rayleigh,silveira2002anisotropic,mocquet1990three}, we observe along the negative anomaly structure following the Appalachian continental shelf, a stretched positive anomaly structure from the Suwanne terrane in Florida to the Avalonian terranes of New-Scotland.
+		At 60~km, it is made of several blocks of at most $+3\%$ anomalies emdedded in a $+2\%$ structure, associated with zero to positive $\xi$ anomalies.
+Below 100~km the pan african lithosphere \citep{yuan2014lithospheric} is cut in two by a slow velocity anomaly comprising both the Great Meteor Track (GMT) and Bermuda tracks, and further southeast, the Low Velocities Fingers (LVF) observed in \cite{french2013waveform}.
+This is best observed in figure~\ref{Vs-circle}.
+		Further East, the slow anomaly follows the GMT crossing the continental shelf, Appalachian and Grenville orogens down to 200~km.
+This slow anomaly corresponds to the V-shape dent, or divot, observed by \cite{lee1997upper} who explained the feature by the presence of water in the mantle.
+We observe at 250~km that the slow anomaly go further inland and crosses the Superior Craton northward until the Hudson bay following the GMT.
+
+		Our model images a slow velocity anomaly in agreement with the reconstruction of the Great Meteor Track proposed by \cite{heaman2000timing}.
+		In figure \ref{cratoncrosscenter} the cross-section DD' follows the Great Meteor track and shows the thinning of the cratonic lithosphere that may be attributed to the passage of the Great Meteor hotspot during the opening of the atlantic ocean.
+		It is interesting to note that the low velocity channel beneath the base of the lithosphere shows lower velocities than usualy observed beneath the other Archaean cratons.
+
+		% \cite{heaman2000timing}, based on $U-Pb$ perovskite age determinations of kimberlites, proposed a progressive southeastward younging of kimberlite magmatism throughout much of eastern North America.
+Considering it as a strong support for small volume mantle melting that have occurred along the continental portion of the Mesozoic Great Meteor mantle plume hotspot track initiated during the opening of the North Atlantic Ocean at about $200Ma$.
+\citep[See figure 4.
+in][]{heaman2000timing}
+
+		% In figure \ref{cratoncrosscenter} the cross-section DD' follows the Great Meteor track and shows the thinning of the cratonic lithosphere that may be attributed to the passage of the Great Meteor hotspot during the opening of the atlantic ocean.
+
+		After accretion of the Appalachian terranes to eastern Laurentia during the Paleozoic, no further significant episodes of magmatism or tectonism affected the western Superior region until the passage of the Great Meteor hot spot at 200–110~Ma \cite{heaman2000timing,villemaire2012p}.
+		Xenoliths analysis from kimberlites at Kirkaland Lake (south eastern part of the western Superior) exhibit some evidence of diamonds, i.e, the base of the lithosphere was deeper than 170~km at the time of kimberlitic eruption \citep{jones2014electrical} estimated at $160 - 152 \: Ma$ by \cite{heaman2000timing,heaman2003timing}.
+		This can be interpreted as either the lithosphere is today thick, and the lower lithosphere was not sampled by xenolith entrainment in the kimberlitic magmas (but nevertheless diamonds were entrained); or the lowermost lithosphere has been eroded away since eruption, possibly by the movement of the North American plate over the Great Meteor hotspot \citep{heaman2000timing,villemaire2012p,jones2014electrical}.
+In the study of \cite{faure2011seismic}, 23 $\%$ of kimberlite samples from southern Superior region contain diamonds.
+In contrast with 78 $\%$ at the northern Superior.
+This means that the diamond stability field within the mantle has been deteriorated through time and kimberlites that have passed across the mantle during/after such deterioration have less chances to intercept diamonds.
+\citep{faure2011seismic}
+
+		\cite{davies1994thermomechanical} described a process of thermo-mechanical erosion which involves heating-softening of the lithospheric base by a plume tail, followed by mechanical removal of the material due to convective transfer.
+\citep{rondenay2000lithospheric}
+		\cite{kaban2015cratonic} proposed that basal drag induced by plate motion caused a south-western shift of the Superior's cratonic root.
+A feature that was observed by the local inversion of gravity, topography, crustal structure and seismic data.
+This basal deformation would have been enabled by the reheating of the lithosphere due to the passage of the Great Meteor hotpsot \citep[see][]{eaton2007seismic,kaban2015cratonic}.
+
+		Our observations show that despite the archaean age of the Superior craton, the lithospheric root thickness of a craton is not flat and can be altered as broadly observed within the eastern part of the Superior craton.
+
+	% The largest continental keel on Earth is that beneath Laurentia, the cratonic core of North America, where the geological record spans more than 2 billion years of the Precambrian from the Archean formation of the Superior craton to the Proterozoic Grenville orogen [Hoffman, 1988].
+After accretion of the Appalachian terranes to eastern Laurentia during the Paleozoic, no further significant epi- sodes of magmatism or tectonism affected the region until the passage of the Great Meteor hot spot at $200–110 Ma [e.g., Heaman and Kjarsgaard, 2000].
+The hot spot track extends $6000 km from northern Hudson Bay to the northeast US coast, and continues southeast as a chain of Atlantic seamounts.
+Eastern North America thus represents an excellent opportunity to address questions associated with Precambrian tectonics, and the ability of the litho- sphere to resist modification by hot spot tectonism.
+
 	% Its thickness is defined by the limit at which the cooler temperatures of the boundary layer can be resolved from the mantle adiabat (which is pinned to the potential temperature of the mantle, i.e., the temperature of the convecting mantle if it were adiabatically decompressed to the surface of Earth in the solid state).
@@
 
-	Compare studies between ReOs and UPb and what they indicate
-		\paragraph{Superior Craton}
-			In our model, the Superior craton can be split in two parts: The western part showing typical cratonic structures while the eastern part shows slower velocities at all depths.
+	% Is the plate motion slower beneath cratons, if yes maybe basal drag occurs there.
+
+% 	Compare studies between ReOs and UPb and what they indicate
+% 		\paragraph{Superior Craton}
+% 			In our model, the Superior craton can be split in two parts: The western part showing typical cratonic structures while the eastern part shows slower velocities at all depths.
 Below 200 km, the distinction between the eastern and western part correlates the GMT.
 
-		\paragraph{Western Ontario part}
-			The western Superior craton shows typical lithospheric craton signatures with highest velocities as shown in figure ~\ref{3d-VS} and in figure~\ref{cratoncross}.
+% 		\paragraph{Western Ontario part}
+% 			The western Superior craton shows typical lithospheric craton signatures with highest velocities as shown in figure ~\ref{3d-VS} and in figure~\ref{cratoncross}.
 In terms of $\xi$, slight negative to zero perturbations are observed at the very western part, while close to the Mid continental rifting, highest negative perturbations indicate that $V_{SV}$ is higher than $V_{SH}$.
 This last feature is associated with the thinning of the cratonic lithosphere when crossing the Grenville front as observed by \cite{darbyshire2007new}.
-			In the entire region there is a positive velocity gradient with a maximum of ca. $4.7 \: km.s^{-1}$ underneathed by a negative gradient  with a minimum of ca. $4.6km.s^{-1}$ at ca. $250km$ deep.
-% 
-			% associated with a maximum of $\xi$ at ca. $150km$ deep.
+% 			In the entire region there is a positive velocity gradient with a maximum of ca. $4.7 \: km.s^{-1}$ underneathed by a negative gradient  with a minimum of ca. $4.6km.s^{-1}$ at ca. $250km$ deep.
+% % 
+% 			% associated with a maximum of $\xi$ at ca. $150km$ deep.
 
-			Opposed to the Slave craton, the topography of the top of the negative gradient is variable. 
-			From SW to NE, the core of maximum velocity can reach beween 70 and 150km of thickness. 
-			Beneath the maximum velocity region of  $\sim$ 150 km the negative gradient appears stipper than beneath the $70km$ one, while the topography of the deeper positive gradient appears flat.
+% 			Opposed to the Slave craton, the topography of the top of the negative gradient is variable. 
+% 			From SW to NE, the core of maximum velocity can reach beween 70 and 150km of thickness. 
+% 			Beneath the maximum velocity region of  $\sim$ 150 km the negative gradient appears stipper than beneath the $70km$ one, while the topography of the deeper positive gradient appears flat.
 The topography of the bottom of the negative gradient is observed flat {\color{red} It is the same in Slave - One linkk to share with other regions}
 
-			% The band of minimum velocity within the cratonic root has similar amplitudes, in terms of velocity, than beneath the younger lithosphere but at shalower depths between $100$ and $150 \: km$. 
-			Following the Great Meteor track built by \cite{heaman2000timing}, which crosses the Superior craton, it is interesting to note that the low velocity channel beneath the base of the lithosphere shows lower velocties than usualy observed beneath the other Archaean cratons.
-			Xenoliths analysis from kimberlites at Kirkaland Lake (south eastern part of the western Superior) exhibit some evidence of diamonds, i.e, the base of the lithosphere was deeper than 170 km at the time of kimberlitic eruption \citep{jones2014electrical} estimated at $160 - 152 \: Ma$ by \cite{heaman2000timing,heaman2003timing}.
-			This can be interpreted as either the lithosphere is today thick, and the lower lithosphere was not sampled by xenolith entrainment in the kimberlitic magmas (but nevertheless diamonds were entrained); or the lowermost lithosphere has been eroded away since eruption, possibly by the movement of the North American plate over the Great Meteor hotspot \citep{heaman2000timing,jones2014electrical}.
+% 			% The band of minimum velocity within the cratonic root has similar amplitudes, in terms of velocity, than beneath the younger lithosphere but at shalower depths between $100$ and $150 \: km$. 
+% 			Following the Great Meteor track built by \cite{heaman2000timing}, which crosses the Superior craton, it is interesting to note that the low velocity channel beneath the base of the lithosphere shows lower velocties than usualy observed beneath the other Archaean cratons.
+% 			Xenoliths analysis from kimberlites at Kirkaland Lake (south eastern part of the western Superior) exhibit some evidence of diamonds, i.e, the base of the lithosphere was deeper than 170 km at the time of kimberlitic eruption \citep{jones2014electrical} estimated at $160 - 152 \: Ma$ by \cite{heaman2000timing,heaman2003timing}.
+% 			This can be interpreted as either the lithosphere is today thick, and the lower lithosphere was not sampled by xenolith entrainment in the kimberlitic magmas (but nevertheless diamonds were entrained); or the lowermost lithosphere has been eroded away since eruption, possibly by the movement of the North American plate over the Great Meteor hotspot \citep{heaman2000timing,jones2014electrical}.
 In the study of \cite{faure2011seismic}, 23 $\%$ of kimberlite samples from southern Superior region contain diamonds.
@@
 
-			\cite{kaban2015cratonic} proposed that basal drag induced by plate motion caused a south-western shift of the Superior's cratonic root.
+% 			\cite{kaban2015cratonic} proposed that basal drag induced by plate motion caused a south-western shift of the Superior's cratonic root.
 A feature that was observed by the local inversion of gravity, topography, crustal structure and seismic data.
 
-			This basal deformation would have been enabled by the reheating of the lithosphere due to the passage of the Great Meteor hotpsot \citep[see][]{eaton2007seismic,kaban2015cratonic}.
-			Our observations show that despite the archaean age of the Superior craton, the lithospheric root thickness of a craton is not flat and can be altered as broadly observed within the eastern part of the Superior craton.
+% 			This basal deformation would have been enabled by the reheating of the lithosphere due to the passage of the Great Meteor hotpsot \citep[see][]{eaton2007seismic,kaban2015cratonic}.
+% 			Our observations show that despite the archaean age of the Superior craton, the lithospheric root thickness of a craton is not flat and can be altered as broadly observed within the eastern part of the Superior craton.
 
@@
 
-		\paragraph{Eastern Quebec part}
-			Compared to its western part, the structure here is different than typical cratonic structure as shown in figure ~\ref{cratoncrossedge} 
-			Fast velocities are present down to 150 km marking the shelf of the lithospheric nucleus. 
-			Below, zero to slight positive anomalies caracterize the eastern superior craton (as in the wedded Grenville belt) showing the retractation toward the continent's center of lithospheric nucleus.
+		% \paragraph{Eastern Quebec part}
+		% 	Compared to its western part, the structure here is different than typical cratonic structure as shown in figure ~\ref{cratoncrossedge} 
+		% 	Fast velocities are present down to 150 km marking the shelf of the lithospheric nucleus. 
+		% 	Below, zero to slight positive anomalies caracterize the eastern superior craton (as in the wedded Grenville belt) showing the retractation toward the continent's center of lithospheric nucleus.
 A feature that was not present in our previous less well resolved models \citep{marone2007three,yuan20113,yuan2014multiscale}.
-			Except at its south western part, negative gradient in terms of velocity is present from 50 to 250 km deep. 
-			Whereas at the center of the continent, velocities are usually increasing (positive gradient) between 50 and 150 km, here it monotically decreases.
-			The negative gradient is steeper and the minimum is radially thicker (more than 100 km thick compare to $\sim$50 km beneath the craton nucleus).
+		% 	Except at its south western part, negative gradient in terms of velocity is present from 50 to 250 km deep. 
+		% 	Whereas at the center of the continent, velocities are usually increasing (positive gradient) between 50 and 150 km, here it monotically decreases.
+		% 	The negative gradient is steeper and the minimum is radially thicker (more than 100 km thick compare to $\sim$50 km beneath the craton nucleus).
 
-			Observed also by Villemaire 2012 with P-wave arrival time. Nice discussion in the end of the paper about lithospheric thinning.
+		% 	Observed also by Villemaire 2012 with P-wave arrival time. Nice discussion in the end of the paper about lithospheric thinning.
 
@@
 
-\subsubsection{Surrounding oceanic}
+% \subsubsection{Surrounding oceanic}
 
-	Our inversion is focusing on the continental part, as shown in our spatial parameterization (i.e. see figure~\ref{density_coverage}) where the region is sampled by data.
-	Except the oceanic borders of NA, further away, the structure has not been updated from SEMucb\_wm1 \citep{french2014whole}.
+% 	Our inversion is focusing on the continental part, as shown in our spatial parameterization (i.e. see figure~\ref{density_coverage}) where the region is sampled by data.
+% 	Except the oceanic borders of NA, further away, the structure has not been updated from SEMucb\_wm1 \citep{french2014whole}.
 
-	\subsubsection{Eastern Passive margin}
-		As in \cite{schaeffer2014imaging,yuan2014lithospheric,james2014rayleigh,silveira2002anisotropic,mocquet1990three}, we observe along the negative anomaly structure following the Appalachian continental shelf, a stretched positive anomaly structure from the Suwanne terrane in Florida to the Avalonian terranes of New-Scotland.
-		At $60km$ it is made of several blocks of at most $+3\%$ anomalies emdedded in a $+2\%$ structure, associated with zero to positive $\xi$ anomalies.
+% 	\subsubsection{Eastern Passive margin}
+% 		As in \cite{schaeffer2014imaging,yuan2014lithospheric,james2014rayleigh,silveira2002anisotropic,mocquet1990three}, we observe along the negative anomaly structure following the Appalachian continental shelf, a stretched positive anomaly structure from the Suwanne terrane in Florida to the Avalonian terranes of New-Scotland.
+% 		At $60km$ it is made of several blocks of at most $+3\%$ anomalies emdedded in a $+2\%$ structure, associated with zero to positive $\xi$ anomalies.
 At $150km$, the structure is split into 2 blocks by a slow anomaly comprising both the Great Meteor (GMT) and Bermuda tracks, and further southeast, the Low Velocities Fingers (LVF) observed in \cite{french2013waveform}.
 
-		Such structure would correspond to late proterozoic gondwanian continental terranes that formed during the Pan African orogenesis \citep{kennedy1964structural} and remained attached to the North American margin during the opening of the Atlantic ocean.
+% 		Such structure would correspond to late proterozoic gondwanian continental terranes that formed during the Pan African orogenesis \citep{kennedy1964structural} and remained attached to the North American margin during the opening of the Atlantic ocean.
 \citep[e.g.][]{thomas2006tectonic}
-		On the continental shelf, authors \citep[e.g.][]{o1983avalon,nance2002cordilleran} identified the west Avalonian terranes of pan-african affinity as in Florida \citep[see][]{smith1982review} that might have been underlayed by a thin (ca.
+% 		On the continental shelf, authors \citep[e.g.][]{o1983avalon,nance2002cordilleran} identified the west Avalonian terranes of pan-african affinity as in Florida \citep[see][]{smith1982review} that might have been underlayed by a thin (ca.
 $100 \: km$) lithosphere before the Pangaea assembling.
@@
 
-		Fast veolicity associated with typical oceanic radial anisotropic structures are observed beneath the whole gulf of Mexico down to $150 \: km$ deep.
+% 		Fast veolicity associated with typical oceanic radial anisotropic structures are observed beneath the whole gulf of Mexico down to $150 \: km$ deep.
 Despite the proximity with pan-african lithosphere, such high velocity anomalies represent older than $\sim 150 \: Ma$ oceanic lithosphere.
 \citep{muller2008age,pindell2009tectonic}
-		-- No Magnetic anomalies recorded as in Pan African, Oceanic seafloor according to Pindel --
+% 		-- No Magnetic anomalies recorded as in Pan African, Oceanic seafloor according to Pindel --
 
@@
 
-\subsubsection{The Great Meteor track}
+% \subsubsection{The Great Meteor track}
 
-	Our model images a slow velocity anomaly in agreement with the reconstruction of the Great Meteor Track (thin red line in figure ~\ref{3d-VS}) proposed by \cite{heaman2000timing}.
-	Below $100km$ the pan african lithosphere is cut in two by a slow velocity anomaly beneath the New England seamounts (youngest part of the GMT).
+% 	Our model images a slow velocity anomaly in agreement with the reconstruction of the Great Meteor Track (thin red line in figure ~\ref{3d-VS}) proposed by \cite{heaman2000timing}.
+% 	Below $100km$ the pan african lithosphere is cut in two by a slow velocity anomaly beneath the New England seamounts (youngest part of the GMT).
 Further East, the slow anomaly follows the GMT crossing the continental shelf, Appalachian and Grenville orogens down to $200km$.
-	This slow anomaly corresponds to the V-shape dent, or divot, observed by \cite{lee1997upper} who explained the feature by the presence of water in the mantle. 
-	We observe at $250km$ that the slow anomaly go further inland and crosses the Superior Craton northward until the Hudson bay following the GMT.
-	See the cross section along the re-built GMT in figure ~\ref{gmtcross}.
+% 	This slow anomaly corresponds to the V-shape dent, or divot, observed by \cite{lee1997upper} who explained the feature by the presence of water in the mantle. 
+% 	We observe at $250km$ that the slow anomaly go further inland and crosses the Superior Craton northward until the Hudson bay following the GMT.
+% 	See the cross section along the re-built GMT in figure ~\ref{gmtcross}.
 
-	\cite{heaman2000timing}, based on $U-Pb$ perovskite age determinations of kimberlites, proposed a progressive southeastward younging of kimberlite magmatism throughout much of eastern North America.
+% 	\cite{heaman2000timing}, based on $U-Pb$ perovskite age determinations of kimberlites, proposed a progressive southeastward younging of kimberlite magmatism throughout much of eastern North America.
 Considering it as a strong support for small volume mantle melting that have occurred along the continental portion of the Mesozoic Great Meteor mantle plume hotspot track initiated during the opening of the North Atlantic Ocean at about $200Ma$.
@@
 
-	\cite{davies1994thermomechanical} described a process of thermo-mechanical erosion which involves heating-softening of the lithospheric base by a plume tail, followed by mechanical removal of the material due to convective transfer.
+% 	\cite{davies1994thermomechanical} described a process of thermo-mechanical erosion which involves heating-softening of the lithospheric base by a plume tail, followed by mechanical removal of the material due to convective transfer.
 \citep{rondenay2000lithospheric}
@@
 		\centering
-		\includegraphics[width=1\textwidth]{figures/Vs-60-300-for-paper.png}
+		\includegraphics[width=1\textwidth]{figures/Vs-60-300-for-paper-with-borders.png}
 
@@
 
-		\centerline{\includegraphics[width=\linewidth]{figures/Cross-section-with-geoltest.pdf}}
+		\centerline{\includegraphics[width=\linewidth]{figures/xsections-SN-WE-final.png}}
 
@@
 
-		\centerline{\includegraphics[width=\linewidth]{figures/Cross-gradient.pdf}}
+		\centerline{\includegraphics[width=\linewidth]{figures/xsections-craton-cross-gmt-tho.png}}
 
@@
 		Cross Section of the lithospheric core. 
-		In each panel cross section of $V_S$ (top) and $\xi$ perturbation from the Isotropic case ($\xi = 1$) (bottom).
-		The right side shows cross section of $V_S$ (top) and $\xi$ perturbation from the Isotropic case ($\xi = 1$).
-The left side shows the radial gradient defined as $\frac{dV_S}{dz}$ in the top panel, and as $\frac{d\xi}{dz}$.}
+		Cross Section of the lithospheric core with absolute velocities of $V_S$ in km.s$^{-1}$.}
 		\label{cratoncrosscenter}
+
+	\end{figure}
+
+	\begin{figure}
+
+		\centerline{\includegraphics[width=\linewidth]{figures/xsections-SN-WE-final-XI.png}}
+
+		\caption{\baselineskip 18pt
+		Cross Section of the lithospheric core at its edges. 
+		Cross Section of the lithospheric core with perturbations of $\xi$ in \% with respect to the isotropic case (i.e $\xi = 1$).}
+		\label{cratoncrossedge-xi}
 
@@
 	\begin{figure}
-		\centerline{\includegraphics[width=\linewidth]{figures/profiles_NASEM3_vs_dxi_GMTrack.pdf}}
+		\centering
+		\includegraphics[width=1\textwidth]{figures/Vs-60-300-for-paper-circle.png}
 
-		\caption{Cross Section of the lithosphere along the Great Meteor track based on \cite{heaman2000timing} construction}
-		\label{gmtcross}
+		\caption{\baselineskip 18pt
+		3D isotropic shear wave velocity structure of the continent embeded in SEM\_ucbwm1 (starting model).
+Map views are shown from 60 km down to 250 km, as variations with respect to the regional mean as $\frac{dV_S}{V_{S_{O}}}$ .
+		Below each map, depth and its regional shear velocity mean is indicated. Minimum and maximum perturbations are also indicated..
+		}
+
+		\label{Vs-circle}
 
 	\end{figure}
+
+	% \begin{figure}
+	% 	\centerline{\includegraphics[width=\linewidth]{figures/Vs-60-300-for-paper-circle.png}}
+
+	% 	\caption{3D isotropic shear wave velocity structure of the continent embeded in SEM\_ucbwm1 (starting model).
+Map views are shown from 60 km down to 250 km, as variations with respect to the regional mean as $\frac{dV_S}{V_{S_{O}}}$ .
+	% 	Below each map, depth and its regional shear velocity mean is indicated. Minimum and maximum perturbations are also indicated.}
+	% 	\label{Vs-circle}
+
+	% \end{figure}
 
